---
title: "main"
author: "Ashley Blawas"
date: '2024-06-13'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries include=FALSE}
library(tidyverse)
library(readxl)
library(stringr)
library(dplyr)
```

```{r}
#options(repos = "http://cran.rstudio.com/")

#Install tinytex package
#install.packages("tinytex")

#Install tinytex
#tinytex::install_tinytex(force = TRUE)
```

```{r  include=FALSE}
# Where does data live?
#data_path = "C:/Users/ashle/Dropbox/Ashley/Post-Doc/Projects/DQO preconditioning/data"

# What is the file name?
#file_name = "10364_Supplemental_Data_061224.xlsx"

# What are the sheet names
sheet1 = "S1. Unenriched Metadata"
sheet2 = "S2. Uneriched Abundances" # NP = # of precursors, PGQ = protein group quantity
sheet3 = "S3. Phos Metadata"
sheet4 = "S4. Phos localized" # Place where it was on the protein
sheet5 = "S5. Phos not localized" # Not able to place where it was on the protein
sheet6 = "S6. Phos filtered" # 
```

## Load in the metadata
```{r}
my_excel <- read_excel("10364_Supplemental_Data_061224.xlsx")

# Load in the Unenriched Metadata
#unenriched_metadata <- read_excel(paste0(data_path, "/", file_name), sheet = sheet1)
unenriched_metadata <- read_excel(paste0("10364_Supplemental_Data_061224.xlsx"), sheet = sheet1)

# Make the Animal IDs for the controls unique
n_na = sum(is.na(unenriched_metadata$Condition))
unenriched_metadata$Condition[is.na(unenriched_metadata$Condition)] <- paste0("NA_VALUE_", seq_len(n_na))

# Load in the Phosphorylation Metadata
#phos_metadata <- read_excel(paste0(data_path, "/", file_name), sheet = sheet3)
phos_metadata <- read_excel(paste0("10364_Supplemental_Data_061224.xlsx"), sheet = sheet3)

# Make the Animal IDs for the controls unique
n_na = sum(phos_metadata$Replicate == "SPQC_prep")
phos_metadata$Replicate[phos_metadata$Replicate == "SPQC_prep"] <- paste0("SPQC_prep", seq_len(n_na))

n_na = sum(phos_metadata$Replicate == "SPQC_tech")
phos_metadata$Replicate[phos_metadata$Replicate == "SPQC_tech"] <- paste0("SPQC_tech", seq_len(n_na))
```

## Load in the unenriched dataset
```{r}
# Load in the Unenriched Metadata
#unenriched_abundances <- read_excel(paste0(data_path, "/", file_name), sheet = sheet2)
unenriched_abundances <- read_excel(paste0("10364_Supplemental_Data_061224.xlsx"), sheet = sheet2)

for (i in 1:nrow(unenriched_metadata)){
  curr_name = names(unenriched_abundances)[i+3] 
  curr_id = unlist(str_extract_all(curr_name, "ID.+htrms"))
  
  temp_metadata = unenriched_metadata %>%
  filter(str_detect(`File Name`, curr_id))
  
  curr_animal = temp_metadata$Animal
  curr_condition = temp_metadata$Condition
  
  unenriched_abundances <- unenriched_abundances %>% rename(!! paste0(curr_animal, "_", curr_condition, "_", "NP") := curr_name)
  
}


for (i in 1:nrow(unenriched_metadata)){
  curr_name = names(unenriched_abundances)[i+36] 
  curr_id = unlist(str_extract_all(curr_name, "ID.+htrms"))
  
  temp_metadata = unenriched_metadata %>%
  filter(str_detect(`File Name`, curr_id))
  
  curr_animal = temp_metadata$Animal
  curr_condition = temp_metadata$Condition
  
  unenriched_abundances <- unenriched_abundances %>% rename(!! paste0(curr_animal, "_", curr_condition, "_", "PGQ") := curr_name)
  
}
```

## Load in the filtered phosphoprotein dataset
```{r}
# Load in the Unenriched Metadata
#phos_filtered <- read_excel(paste0(data_path, "/", file_name), sheet = sheet6)
phos_filtered <- read_excel(paste0("10364_Supplemental_Data_061224.xlsx"), sheet = sheet6)

# Drop CV columns
phos_filtered <- phos_filtered %>% dplyr::select(!starts_with("CV"))

# Change data frame organization to have names
for (i in 1:nrow(phos_metadata)){
  curr_name = names(phos_filtered)[i+5] # Add 5 because there are 5 columns that are not IDs
  curr_id = unlist(str_extract_all(curr_name, "ID.+htrms"))
  
  temp_metadata = phos_metadata %>%
  filter(str_detect(`File Name`, curr_id))
  
  curr_animal = temp_metadata$Condition
  curr_condition = temp_metadata$Replicate
  
  phos_filtered <- phos_filtered %>% rename(!! paste0(curr_animal, "_", curr_condition, "_", "EG") := curr_name)
  
}
```


# Make data frame for plotting

```{r}
phos_filtered <- phos_filtered %>% select(-PG.ProteinAccessions, -PG.ProteinDescriptions, -PG.ProteinNames, -EG.PrecursorId)

phos_filtered <- phos_filtered %>% pivot_longer(!PG.Genes, names_to = "Sample") %>%
  rename(Value = value) %>% 
  rename(Gene = PG.Genes)

# Mutate new columns
phos_filtered <- phos_filtered %>% mutate(Animal = str_extract(Sample, "^\\D+")) %>%
  mutate(Date = str_extract(Sample, "\\d+")) %>%
  mutate(Condition = sub("^[^_]*_([^_]*).*", "\\1", Sample)) %>%
  dplyr::select(-Sample) %>%
  dplyr::select(Gene, Animal, Condition, Date, Value)
```

# Look at change between time point A & C and C & D
```{r}
#Ashley's initial work below
#upup <- phos_filtered %>% 
#  group_by(Gene, Animal) %>%
#  if

#NEW CECE
#Making a version of the phos_filtered dataset with only the max values for each condition
phos_filtered_max_values <- phos_filtered %>%
  group_by(Gene, Animal, Condition) %>%
  summarize(MaxValue = max(Value)) %>%
  filter(Condition %in% c("CA", "CC", "CD", "PC6A", "PC6C", "PC6D")) %>%
  ungroup()

#Making datasets that calculate and store the differences between CA and CC, CD and CC, PC6C and PC6A, and PC6D and PC6C
phos_filtered_change_CA_CC <- phos_filtered_max_values %>%
  group_by(Gene, Animal) %>%
  summarize(Difference = MaxValue[Condition == "CC"] - MaxValue[Condition == "CA"]) %>%
  mutate(Condition = "CA_CC") %>%
  bind_rows(phos_filtered_max_values)

phos_filtered_change_CC_CD <- phos_filtered_max_values %>%
  group_by(Gene, Animal) %>%
  summarize(Difference = MaxValue[Condition == "CD"] - MaxValue[Condition == "CC"]) %>% mutate(Condition = "CC_CD") %>%
  bind_rows(phos_filtered_max_values)

phos_filtered_change_PC6A_PC6C <- phos_filtered_max_values %>%
  group_by(Gene, Animal) %>%
  summarize(Difference = MaxValue[Condition == "PC6C"] - MaxValue[Condition == "PC6A"]) %>%
  mutate(Condition = "PC6A_PC6C") %>%
  bind_rows(phos_filtered_max_values)

phos_filtered_change_PC6C_PC6D <- phos_filtered_max_values %>%
  group_by(Gene, Animal) %>%
  summarize(Difference = MaxValue[Condition == "PC6D"] - MaxValue[Condition == "PC6C"]) %>%
  mutate(Condition = "PC6C_PC6D") %>%
  bind_rows(phos_filtered_max_values)

#Making an overall dataset combining all of these datasets
phos_filtered_change <- bind_rows(phos_filtered_change_CA_CC, phos_filtered_change_CC_CD, phos_filtered_change_PC6A_PC6C, phos_filtered_change_PC6C_PC6D) %>%
  filter(Condition %in% c("CA_CC", "CC_CD", "PC6A_PC6C", "PC6C_PC6D"))

print(phos_filtered_change)
```

```{r}
#Printing a sample geom_point to visualize these changes for the HBA gene
phos_filtered_change %>%
  filter(Gene == "HBA") %>%
  group_by(Animal, Condition) %>%
  summarize(MaxValue = max(Difference)) %>%
  ungroup() %>%
  ggplot(aes(x=Condition, y = MaxValue)) + facet_wrap(vars(Animal), scales = "free") + geom_point()
```

#Making new dataset to conduct paired t-test w/ protein abundance and phosphorylation
```{r}
# Load in sheet2
sheet2_metadata <- read_excel(paste0("10364_Supplemental_Data_061224.xlsx"), sheet = sheet2)

# Replace NaN values with 0 for all numeric columns
sheet2_metadata[] <- lapply(sheet2_metadata, function(x) {
  if (is.numeric(x)) {
    x[is.nan(x)] <- 0
  }
  return(x)
})

# Function to convert NaN to NA in character columns
sheet2_metadata[] <- lapply(sheet2_metadata, function(x) {
  if (is.character(x)) {
    x[x == "NaN"] <- NA  # Convert "NaN" string to NA
  }
  return(x)
})

# Replace NA values with "0" in the entire data frame
sheet2_metadata[is.na(sheet2_metadata)] <- "0"

# Convert the character columns to numeric columns
sheet2_metadata$"[1] ID112569_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[1] ID112569_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[2] ID112570_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[2] ID112570_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[3] ID112571_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[3] ID112571_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[4] ID112572_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[4] ID112572_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[5] ID112573_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[5] ID112573_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[6] ID112574_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[6] ID112574_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[7] ID112575_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[7] ID112575_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[8] ID112576_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[8] ID112576_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[9] ID112577_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[9] ID112577_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[10] ID112578_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[10] ID112578_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[11] ID112579_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[11] ID112579_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[12] ID112580_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[12] ID112580_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[13] ID112581_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[13] ID112581_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[14] ID112582_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[14] ID112582_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[15] ID112583_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[15] ID112583_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[16] ID112584_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[16] ID112584_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[17] ID112585_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[17] ID112585_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[18] ID112586_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[18] ID112586_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[19] ID112587_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[19] ID112587_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[20] ID112588_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[20] ID112588_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[21] ID112589_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[21] ID112589_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[22] ID112590_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[22] ID112590_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[23] ID112592_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[23] ID112592_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[24] ID112593_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[24] ID112593_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[25] ID112594_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[25] ID112594_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[26] ID112595_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[26] ID112595_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[27] ID112596_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[27] ID112596_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[28] ID112597_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[28] ID112597_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[29] ID112598_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[29] ID112598_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[30] ID112599_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[30] ID112599_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[31] IDSPQC_02_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[31] IDSPQC_02_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[32] IDSPQC_03_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[32] IDSPQC_03_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[33] IDSPQC_04_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[33] IDSPQC_04_MA10140C_10364_050724.htrms.PG.Quantity")

# Log2 normalization applied to all numeric columns
# Added 1 before log transformation to avoid taking log of zero
sheet2_metadata_normalized <- as.data.frame(lapply(sheet2_metadata, function(x) if(is.numeric(x)) log2(x + 1) else x))
```

#Conducting paired T-tests to assess how an individual's phopho levels at a specific residue change between points A, B, C, and D
```{r}
# Generic paired t-test format
#t_test_result <- t.test(data_set$column_before, data_set$column_after, paired = TRUE)

#Kolohe CA to CC
t_test_kolohe_CA_CC <- t.test(sheet2_metadata_normalized$`X.1..ID112569_01_MA10140C_10364_050724.htrms.PG.Quantity`, sheet2_metadata_normalized$`X.2..ID112570_01_MA10140C_10364_050724.htrms.PG.Quantity`, paired = TRUE)
print(t_test_kolohe_CA_CC)
#This approach cannot be correct because it is bunching all of the residues together to compare between these two scenarios

#Instead, try rearranging the dataset to compare CA and CC phosphorylation for a specific residue among all individuals

#other issue is I can't even use Sheet 2 exclusively bc it is by protein, not residue, bc protein amount is done by protein

#therefore, sheet 2 is more protein-wide phosphorylation rather than residue phosphorylation
```

##Making a heatmap for Figure 3 in the paper plan
#Phosphorylation results broad overview
```{r}
install.packages("readxl")
install.packages("dplyr")
install.packages("ggplot2")
install.packages("pheatmap")
install.packages("reshape2")

library(readxl)
library(dplyr)
library(ggplot2)
library(pheatmap)
library(reshape2)
```

```{r}
# Load data from Excel
heatmap_data_ <- read_excel("DQO Heatmap Datasheet.xlsx", sheet = 1)  # Adjust file path and sheet number/name

install.packages("dplyr")
library(dplyr)

# Select columns 2, 16, and 57
heatmap_data_filtered <- heatmap_data_ %>%
  select(2, 36, 37, 38, 39)
```

```{r}
heatmap_data_filtered <- na.omit(heatmap_data_filtered)

heatmap_data_filtered <- heatmap_data_filtered %>%
  mutate_at(vars(2, 3, 4, 5), ~ as.numeric(ifelse(. >= 1, 2, 1)))
```

```{r}
# Multiply Column 2 by Column 3 and store the result in a new column (Column 6)
heatmap_data_filtered <- heatmap_data_filtered %>%
  mutate(Kolohe = `Kolohe_Control_Diff_CA` * `Kolohe_PC6_Diff_CA`)

heatmap_data_filtered <- heatmap_data_filtered %>%
  mutate(Hoku = `Hoku_Control_Diff_CA` * `Hoku_PC6_Diff_CA`)
```

```{r}
# Remove columns 2, 3, 4, and 5 by their indices
heatmap_data_filtered <- heatmap_data_filtered %>%
  select(-c(2, 3, 4, 5))
```

```{r}
library(reshape2) 

# Reshape data into long format
heatmap_data_filtered_long <- melt(heatmap_data_filtered, id.vars = "PG.Genes", variable.name = "Condition", value.name = "Value")
```

```{r}
# Create the heatmap plot
heatmap_data_filtered_long[3434, 1] <- "HNRNPH2;HNRNPH1"

heatmap_data_filtered_long_test <- heatmap_data_filtered_long[c(1, 56, 102, 546), ]

#Creating the heat map
heatmap_plot <- ggplot(heatmap_data_filtered_long_test, aes(x = Condition, y = `PG.Genes`, fill = factor(Value))) +
  geom_tile(color = "white", linewidth = 1) + 
  scale_fill_manual(values = c("1" = "blue", "2" = "white", "4" = "red")) + 
  labs(title = "Gene Expression Heatmap", x = "Condition", y = "Gene", fill = "Value") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(heatmap_plot)
```

```{r}
# Making a heatmap where each gene is different
heatmap_data_filtered_copy <- heatmap_data_filtered

# If your first column is named "Gene", replace df[, 1] with df$Gene
new_names <- character(nrow(heatmap_data_filtered_copy))
counter <- 1

for (i in 1:nrow(heatmap_data_filtered_copy)) {
  if (i == 1 || heatmap_data_filtered_copy$`PG.Genes`[i] != heatmap_data_filtered_copy$`PG.Genes`[i - 1]) {
    counter <- 1
  } else {
    counter <- counter + 1
  }
  new_names[i] <- paste0(heatmap_data_filtered_copy$`PG.Genes`[i], ".", counter)
}

heatmap_data_filtered_copy$New_Gene_Name <- new_names

heatmap_data_filtered_copy$`PG.Genes` <- NULL

# Reshape data into long format
heatmap_data_filtered_copy_long <- melt(heatmap_data_filtered_copy, id.vars = "New_Gene_Name", variable.name = "Condition", value.name = "Value")

heatmap_data_filtered_copy_long_test <- heatmap_data_filtered_copy_long[c(1, 56, 102, 546, 8336, 8391, 13040), ]

#Creating the heat map
heatmap_plot <- ggplot(heatmap_data_filtered_copy_long, aes(x = Condition, y = `New_Gene_Name`, fill = factor(Value))) +
  geom_tile(color = "white", linewidth = 1) + 
  scale_fill_manual(values = c("1" = "blue", "2" = "white", "4" = "red")) + 
  labs(title = "Gene Expression Heatmap", x = "Condition", y = "Gene", fill = "Value") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(heatmap_plot)

ggsave("plot.png", plot = last_plot(), width = 100, height = 150, units = "in", dpi = 300, limitsize=FALSE)
```
#Filtered data to get rid of between 0.5 and 2 & setting up for heat map
```{r}
library(dplyr)

#filtered_heatmap_ <- read_excel("NEW_Filtered_HeatMap.xlsx")
#now loaded in as NEW_Filtered_HeatMap

NEW_Filtered_HeatMap <- NEW_Filtered_HeatMap[, c("PG.Genes", "Differential of Kolohe Control C/A", "Differential of Hoku Control C/A", "Differential of Kolohe PC6 C/A", "Differential of Hoku PC6 C/A")]

NEW_Filtered_HeatMap$`Differential of Kolohe Control C/A` <- as.numeric(NEW_Filtered_HeatMap$`Differential of Kolohe Control C/A`)

NEW_Filtered_HeatMap$`Differential of Kolohe PC6 C/A` <- as.numeric(NEW_Filtered_HeatMap$`Differential of Kolohe PC6 C/A`)

NEW_Filtered_HeatMap$`Differential of Hoku Control C/A` <- as.numeric(NEW_Filtered_HeatMap$`Differential of Hoku Control C/A`)

NEW_Filtered_HeatMap$`Differential of Hoku PC6 C/A` <- as.numeric(NEW_Filtered_HeatMap$`Differential of Hoku PC6 C/A`)

#xtrafiltered_heatmap <- NEW_Filtered_HeatMap %>%
#  filter(`Differential of Kolohe Control C/A` <= 0.5 | `Differential of Kolohe Control C/A` >= 2)

xtrafiltered_heatmap <- NEW_Filtered_HeatMap %>%
  filter((`Differential of Kolohe Control C/A` <= 0.5 | `Differential of Kolohe Control C/A` >= 2) | (`Differential of Kolohe PC6 C/A` <= 0.5 | `Differential of Kolohe PC6 C/A` >= 2) | (`Differential of Hoku Control C/A` <= 0.5 | `Differential of Hoku Control C/A` >= 2) | (`Differential of Hoku PC6 C/A` <= 0.5 | `Differential of Hoku PC6 C/A` >= 2))

# Replace NA values in the PG.Genes column with "PRSS1"
xtrafiltered_heatmap$`PG.Genes`[is.na(xtrafiltered_heatmap$`PG.Genes`)] <- "PRSS1"

# Initialize a counter and a vector to store the modified gene names
counter_ <- 1
modified_genes_ <- character(length(xtrafiltered_heatmap$`PG.Genes`))

# Iterate through the 'gene' column and modify the gene names
for (i in 1:length(xtrafiltered_heatmap$`PG.Genes`)) {
  if (i == 1 || xtrafiltered_heatmap$`PG.Genes`[i] != xtrafiltered_heatmap$`PG.Genes`[i - 1]) {
    counter_ <- 1  # Reset counter when a new gene is encountered
  }
  modified_genes_[i] <- paste(xtrafiltered_heatmap$`PG.Genes`[i], counter_, sep = ".")
  counter_ <- counter_ + 1  # Increment counter for each instance of the same gene
}

# Add the modified gene names to the dataframe
xtrafiltered_heatmap$modified_gene_ <- modified_genes_

xtrafiltered_heatmap$`PG.Genes` <- NULL
```

#Make heatmap for rows 1-200 in dataset
```{r}
# Make a dataset with only the first 200 rows
xtrafiltered_heatmap_1_to_200 <- xtrafiltered_heatmap[1:200, ]

# Reshape data into long format
library(reshape2)
#xtrafiltered_heatmap_long <- melt(xtrafiltered_heatmap, id.vars = "modified_gene_", #variable.name = "Condition", value.name = "Value")
xtrafiltered_heatmap_1_to_200_long <- melt(xtrafiltered_heatmap_1_to_200, id.vars = "modified_gene_", variable.name = "Condition", value.name = "Value")


#Creating the heat map
library(ggplot2)
#heatmap_plot <- ggplot(xtrafiltered_heatmap_long, aes(x = Condition, y = `modified_gene_`, fill = cut(value, breaks = c(-Inf, 0.5, 2, Inf), labels = c("blue", "white", "red")))) +
#  geom_tile(color = "white", linewidth = 1) + 
#  scale_fill_manual(values = c("blue" = "blue", "white" = "white", "red" = "red")) + 
#  labs(title = "Gene Expression Heatmap", x = "Condition", y = "Gene", fill = "Value") +
#  theme_minimal() +
#  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Custom labels for the X-axis
custom_labels <- c("Differential of Kolohe Control C/A" = "Kolohe Control", "Differential of Hoku Control C/A" = "Hoku Control", "Differential of Kolohe PC6 C/A" = "Kolohe PC6", "Differential of Hoku PC6 C/A" = "Hoku PC6")

ggplot(xtrafiltered_heatmap_1_to_200_long, aes(x = Condition, y = modified_gene_, fill = cut(Value, breaks = c(-Inf, 0.5, 2, Inf), 
                                          labels = c("blue", "white", "red")))) +
  geom_tile() +  # geom_tile for heatmap
  scale_fill_manual(values = c("blue" = "blue", "white" = "white", "red" = "red")) +
  theme_minimal() +
  labs(title = "Heatmap with custom colors based on value ranges") +
  scale_x_discrete(labels = custom_labels) +  # Change X-axis labels
  theme(axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1)) + #make x-axis text vertical
  theme(axis.text.y = element_text(size = 5, angle = 45, vjust = 0.5))

#need to cluster
```

#Make heatmap for rows 201-400 in dataset
```{r}
# Make a dataset with only rows 201-400
xtrafiltered_heatmap_201_to_400 <- xtrafiltered_heatmap[201:400, ]

# Perform hierarchical clustering on rows (Y-axis)
# Attempt at clustering
xtrafiltered_heatmap_201_to_400[is.na(xtrafiltered_heatmap_201_to_400)] <- 0 #replacing NA with 0 to run the next 3 lines of code successfully
row_dist <- dist(xtrafiltered_heatmap_201_to_400)  # Compute distance matrix
row_clust <- hclust(row_dist)  # Perform hierarchical clustering
#new_order <- rownames(xtrafiltered_heatmap_201_to_400)[row_clust$order]  # Get new row order
new_order <- as.vector(rownames(xtrafiltered_heatmap_201_to_400)[row_clust$order])

# Reorder factor levels for ggplot
# Attempt at clustering
xtrafiltered_heatmap_201_to_400_long$modified_gene_ <- as.character(xtrafiltered_heatmap_201_to_400_long$modified_gene_) #ensure gene columb is a character before reordering
xtrafiltered_heatmap_201_to_400_long$modified_gene_ <- factor(xtrafiltered_heatmap_201_to_400_long$modified_gene_, levels = new_order)

# Reshape data into long format
library(reshape2)
xtrafiltered_heatmap_201_to_400_long <- melt(xtrafiltered_heatmap_201_to_400, id.vars = "modified_gene_", variable.name = "Condition", value.name = "Value")

#Creating the heat map
library(ggplot2)

# Custom labels for the X-axis
custom_labels <- c("Differential of Kolohe Control C/A" = "Kolohe Control", "Differential of Hoku Control C/A" = "Hoku Control", "Differential of Kolohe PC6 C/A" = "Kolohe PC6", "Differential of Hoku PC6 C/A" = "Hoku PC6")

ggplot(xtrafiltered_heatmap_201_to_400_long, aes(x = Condition, y = modified_gene_, fill = cut(Value, breaks = c(-Inf, 0.5, 2, Inf), 
                                          labels = c("blue", "white", "red")))) +
  geom_tile() +  # geom_tile for heatmap
  scale_fill_manual(values = c("blue" = "blue", "white" = "white", "red" = "red")) +
  theme_minimal() +
  labs(title = "Heatmap with custom colors based on value ranges") +
  scale_x_discrete(labels = custom_labels) +  # Change X-axis labels
  theme(axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1)) + #make x-axis text vertical
  theme(axis.text.y = element_text(size = 5, angle = 45, vjust = 0.5))

#need to cluster
```

#Map heatmap for rows 201-400 in dataset in pheatmap
```{r}
install.packages("pheatmap")
 
# load package
library(pheatmap)

# install DESeq if necessary
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
BiocManager::install("DESeq2")
 
# install and load package
library(DESeq2)

# moving gene names to first column in dataset
#xtrafiltered_heatmap_201_to_400_pheatmap <- xtrafiltered_heatmap_201_to_400[c(ncol(xtrafiltered_heatmap_201_to_400), 1:(ncol(xtrafiltered_heatmap_201_to_400)-1))] #If the column you want moved is the last column
xtrafiltered_heatmap_201_to_400_pheatmap <- xtrafiltered_heatmap_201_to_400_pheatmap[c(2, 1, 3:ncol(xtrafiltered_heatmap_201_to_400_pheatmap))] #If the column you want moved is the 2nd column

# Convert tibble to data frame
xtrafiltered_heatmap_201_to_400_pheatmap <- as.data.frame(xtrafiltered_heatmap_201_to_400_pheatmap)

# setting the first column (gene names) as row names
rownames(xtrafiltered_heatmap_201_to_400_pheatmap) <- xtrafiltered_heatmap_201_to_400_pheatmap[, 1]

# removing the first column from the data frame (since it's now row names)
xtrafiltered_heatmap_201_to_400_pheatmap <- xtrafiltered_heatmap_201_to_400_pheatmap[, -1]

# Convert Data to a Numeric Matrix
xtrafiltered_heatmap_201_to_400_pheatmap_matrix <- as.matrix(xtrafiltered_heatmap_201_to_400_pheatmap)

# Replace NAs/NaNs/Inf with zero (otherwise pheatmap will not work)
xtrafiltered_heatmap_201_to_400_pheatmap_matrix[is.na(xtrafiltered_heatmap_201_to_400_pheatmap_matrix) | is.nan(xtrafiltered_heatmap_201_to_400_pheatmap_matrix) | is.infinite(xtrafiltered_heatmap_201_to_400_pheatmap_matrix)] <- 0

xtrafiltered_heatmap_201_to_400_pheatmap_matrix <- scale(xtrafiltered_heatmap_201_to_400_pheatmap_matrix)

# Generate heatmap
pheatmap(xtrafiltered_heatmap_201_to_400_pheatmap_matrix, scale = "row")
```

#Make heatmap for rows 401-600 in dataset
```{r}
# Make a dataset with only rows 401-600
xtrafiltered_heatmap_401_to_600 <- xtrafiltered_heatmap[401:600, ]

# Reshape data into long format
library(reshape2)
xtrafiltered_heatmap_401_to_600_long <- melt(xtrafiltered_heatmap_401_to_600, id.vars = "modified_gene_", variable.name = "Condition", value.name = "Value")


#Creating the heat map
library(ggplot2)

# Custom labels for the X-axis
custom_labels <- c("Differential of Kolohe Control C/A" = "Kolohe Control", "Differential of Hoku Control C/A" = "Hoku Control", "Differential of Kolohe PC6 C/A" = "Kolohe PC6", "Differential of Hoku PC6 C/A" = "Hoku PC6")

ggplot(xtrafiltered_heatmap_401_to_600_long, aes(x = Condition, y = modified_gene_, fill = cut(Value, breaks = c(-Inf, 0.5, 2, Inf), 
                                          labels = c("blue", "white", "red")))) +
  geom_tile() +  # geom_tile for heatmap
  scale_fill_manual(values = c("blue" = "blue", "white" = "white", "red" = "red")) +
  theme_minimal() +
  labs(title = "Heatmap with custom colors based on value ranges") +
  scale_x_discrete(labels = custom_labels) +  # Change X-axis labels
  theme(axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1)) + #make x-axis text vertical
  theme(axis.text.y = element_text(size = 5, angle = 45, vjust = 0.5))

#need to cluster
```

#Make heatmap for rows 601-800 in dataset
```{r}
# Make a dataset with only rows 601-800
xtrafiltered_heatmap_601_to_800 <- xtrafiltered_heatmap[601:800, ]

# Reshape data into long format
library(reshape2)
xtrafiltered_heatmap_601_to_800_long <- melt(xtrafiltered_heatmap_601_to_800, id.vars = "modified_gene_", variable.name = "Condition", value.name = "Value")


#Creating the heat map
library(ggplot2)

# Custom labels for the X-axis
custom_labels <- c("Differential of Kolohe Control C/A" = "Kolohe Control", "Differential of Hoku Control C/A" = "Hoku Control", "Differential of Kolohe PC6 C/A" = "Kolohe PC6", "Differential of Hoku PC6 C/A" = "Hoku PC6")

ggplot(xtrafiltered_heatmap_601_to_800_long, aes(x = Condition, y = modified_gene_, fill = cut(Value, breaks = c(-Inf, 0.5, 2, Inf), 
                                          labels = c("blue", "white", "red")))) +
  geom_tile() +  # geom_tile for heatmap
  scale_fill_manual(values = c("blue" = "blue", "white" = "white", "red" = "red")) +
  theme_minimal() +
  labs(title = "Heatmap with custom colors based on value ranges") +
  scale_x_discrete(labels = custom_labels) +  # Change X-axis labels
  theme(axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1)) + #make x-axis text vertical
  theme(axis.text.y = element_text(size = 5, angle = 45, vjust = 0.5))

#need to cluster
```

#Make heatmap for rows 801-1000 in dataset
```{r}
# Make a dataset with only rows 801-1000
xtrafiltered_heatmap_801_to_1000 <- xtrafiltered_heatmap[801:1000, ]

# Reshape data into long format
library(reshape2)
xtrafiltered_heatmap_801_to_1000_long <- melt(xtrafiltered_heatmap_801_to_1000, id.vars = "modified_gene_", variable.name = "Condition", value.name = "Value")


#Creating the heat map
library(ggplot2)

# Custom labels for the X-axis
custom_labels <- c("Differential of Kolohe Control C/A" = "Kolohe Control", "Differential of Hoku Control C/A" = "Hoku Control", "Differential of Kolohe PC6 C/A" = "Kolohe PC6", "Differential of Hoku PC6 C/A" = "Hoku PC6")

ggplot(xtrafiltered_heatmap_801_to_1000_long, aes(x = Condition, y = modified_gene_, fill = cut(Value, breaks = c(-Inf, 0.5, 2, Inf), 
                                          labels = c("blue", "white", "red")))) +
  geom_tile() +  # geom_tile for heatmap
  scale_fill_manual(values = c("blue" = "blue", "white" = "white", "red" = "red")) +
  theme_minimal() +
  labs(title = "Heatmap with custom colors based on value ranges") +
  scale_x_discrete(labels = custom_labels) +  # Change X-axis labels
  theme(axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1)) + #make x-axis text vertical
  theme(axis.text.y = element_text(size = 5, angle = 45, vjust = 0.5))

#need to cluster
```

#Make heatmap for rows 1001-1200 in dataset
```{r}
# Make a dataset with only rows 1001-1200
xtrafiltered_heatmap_1001_to_1200 <- xtrafiltered_heatmap[1001:1200, ]

# Reshape data into long format
library(reshape2)
xtrafiltered_heatmap_1001_to_1200_long <- melt(xtrafiltered_heatmap_1001_to_1200, id.vars = "modified_gene_", variable.name = "Condition", value.name = "Value")


#Creating the heat map
library(ggplot2)

# Custom labels for the X-axis
custom_labels <- c("Differential of Kolohe Control C/A" = "Kolohe Control", "Differential of Hoku Control C/A" = "Hoku Control", "Differential of Kolohe PC6 C/A" = "Kolohe PC6", "Differential of Hoku PC6 C/A" = "Hoku PC6")

ggplot(xtrafiltered_heatmap_1001_to_1200_long, aes(x = Condition, y = modified_gene_, fill = cut(Value, breaks = c(-Inf, 0.5, 2, Inf), 
                                          labels = c("blue", "white", "red")))) +
  geom_tile() +  # geom_tile for heatmap
  scale_fill_manual(values = c("blue" = "blue", "white" = "white", "red" = "red")) +
  theme_minimal() +
  labs(title = "Heatmap with custom colors based on value ranges") +
  scale_x_discrete(labels = custom_labels) +  # Change X-axis labels
  theme(axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1)) + #make x-axis text vertical
  theme(axis.text.y = element_text(size = 5, angle = 45, vjust = 0.5))

#need to cluster
```

#Make heatmap for rows 1201-1400 in dataset
```{r}
# Make a dataset with only rows 1201-1400
xtrafiltered_heatmap_1201_to_1400 <- xtrafiltered_heatmap[1201:1400, ]

# Reshape data into long format
library(reshape2)
xtrafiltered_heatmap_1201_to_1400_long <- melt(xtrafiltered_heatmap_1201_to_1400, id.vars = "modified_gene_", variable.name = "Condition", value.name = "Value")


#Creating the heat map
library(ggplot2)

# Custom labels for the X-axis
custom_labels <- c("Differential of Kolohe Control C/A" = "Kolohe Control", "Differential of Hoku Control C/A" = "Hoku Control", "Differential of Kolohe PC6 C/A" = "Kolohe PC6", "Differential of Hoku PC6 C/A" = "Hoku PC6")

ggplot(xtrafiltered_heatmap_1201_to_1400_long, aes(x = Condition, y = modified_gene_, fill = cut(Value, breaks = c(-Inf, 0.5, 2, Inf), 
                                          labels = c("blue", "white", "red")))) +
  geom_tile() +  # geom_tile for heatmap
  scale_fill_manual(values = c("blue" = "blue", "white" = "white", "red" = "red")) +
  theme_minimal() +
  labs(title = "Heatmap with custom colors based on value ranges") +
  scale_x_discrete(labels = custom_labels) +  # Change X-axis labels
  theme(axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1)) + #make x-axis text vertical
  theme(axis.text.y = element_text(size = 5, angle = 45, vjust = 0.5))

#need to cluster
```

#Make heatmap for rows 1401-1600 in dataset
```{r}
# Make a dataset with only rows 1401-1600
xtrafiltered_heatmap_1401_to_1600 <- xtrafiltered_heatmap[1401:1600, ]

# Reshape data into long format
library(reshape2)
xtrafiltered_heatmap_1401_to_1600_long <- melt(xtrafiltered_heatmap_1401_to_1600, id.vars = "modified_gene_", variable.name = "Condition", value.name = "Value")


#Creating the heat map
library(ggplot2)

# Custom labels for the X-axis
custom_labels <- c("Differential of Kolohe Control C/A" = "Kolohe Control", "Differential of Hoku Control C/A" = "Hoku Control", "Differential of Kolohe PC6 C/A" = "Kolohe PC6", "Differential of Hoku PC6 C/A" = "Hoku PC6")

ggplot(xtrafiltered_heatmap_1401_to_1600_long, aes(x = Condition, y = modified_gene_, fill = cut(Value, breaks = c(-Inf, 0.5, 2, Inf), 
                                          labels = c("blue", "white", "red")))) +
  geom_tile() +  # geom_tile for heatmap
  scale_fill_manual(values = c("blue" = "blue", "white" = "white", "red" = "red")) +
  theme_minimal() +
  labs(title = "Heatmap with custom colors based on value ranges") +
  scale_x_discrete(labels = custom_labels) +  # Change X-axis labels
  theme(axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1)) + #make x-axis text vertical
  theme(axis.text.y = element_text(size = 5, angle = 45, vjust = 0.5))

#need to cluster
```

#Make heatmap for rows 1601-1755 in dataset
```{r}
# Make a dataset with only rows 201-400
xtrafiltered_heatmap_1601_to_1755 <- xtrafiltered_heatmap[1601:1755, ]

# Reshape data into long format
library(reshape2)
xtrafiltered_heatmap_1601_to_1755_long <- melt(xtrafiltered_heatmap_1601_to_1755, id.vars = "modified_gene_", variable.name = "Condition", value.name = "Value")


#Creating the heat map
library(ggplot2)

# Custom labels for the X-axis
custom_labels <- c("Differential of Kolohe Control C/A" = "Kolohe Control", "Differential of Hoku Control C/A" = "Hoku Control", "Differential of Kolohe PC6 C/A" = "Kolohe PC6", "Differential of Hoku PC6 C/A" = "Hoku PC6")

ggplot(xtrafiltered_heatmap_1601_to_1755_long, aes(x = Condition, y = modified_gene_, fill = cut(Value, breaks = c(-Inf, 0.5, 2, Inf), 
                                          labels = c("blue", "white", "red")))) +
  geom_tile() +  # geom_tile for heatmap
  scale_fill_manual(values = c("blue" = "blue", "white" = "white", "red" = "red")) +
  theme_minimal() +
  labs(title = "Heatmap with custom colors based on value ranges") +
  scale_x_discrete(labels = custom_labels) +  # Change X-axis labels
  theme(axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1)) + #make x-axis text vertical
  theme(axis.text.y = element_text(size = 5, angle = 45, vjust = 0.5))

#need to cluster
```

#PHEATMAP Heatmap
#TRY RUNNING FOR P-SITES THAT WERE UPREGULATED OR DOWNREGULATED SIMILARILY IN ALL TRIALS!
```{r}
#Loading libraries
library(dplyr)
library(pheatmap)
library(DESeq2)
library(RColorBrewer)

#Loaded Pheatmap_Excel into R

#Making all the variables numeric except gene names
Pheatmap_Excel$`Differential of Kolohe Control C/A` <- as.numeric(Pheatmap_Excel$`Differential of Kolohe Control C/A`)
Pheatmap_Excel$`Differential of Hoku Control C/A` <- as.numeric(Pheatmap_Excel$`Differential of Hoku Control C/A`)
Pheatmap_Excel$`Differential of Liho Control C/A` <- as.numeric(Pheatmap_Excel$`Differential of Liho Control C/A`)
Pheatmap_Excel$`Differential of Nainoa Control C/A` <- as.numeric(Pheatmap_Excel$`Differential of Nainoa Control C/A`)
Pheatmap_Excel$`Differential of Lono Control C/A` <- as.numeric(Pheatmap_Excel$`Differential of Lono Control C/A`)
Pheatmap_Excel$`Differential of Kolohe PC6 C/A` <- as.numeric(Pheatmap_Excel$`Differential of Kolohe PC6 C/A`)
Pheatmap_Excel$`Differential of Hoku PC6 C/A` <- as.numeric(Pheatmap_Excel$`Differential of Hoku PC6 C/A`)

#Filtering out phosphosites that do not have any significant regulation in any sample
Pheatmap_Excel_1.0 <- Pheatmap_Excel %>%
  filter((`Differential of Kolohe Control C/A` <= 0.5 | `Differential of Kolohe Control C/A` >= 2) | (`Differential of Hoku Control C/A` <= 0.5 | `Differential of Hoku Control C/A` >= 2) | (`Differential of Liho Control C/A` <= 0.5 | `Differential of Liho Control C/A` >= 2) | (`Differential of Nainoa Control C/A` <= 0.5 | `Differential of Nainoa Control C/A` >= 2) | (`Differential of Lono Control C/A` <= 0.5 | `Differential of Lono Control C/A` >= 2) | (`Differential of Kolohe PC6 C/A` <= 0.5 | `Differential of Kolohe PC6 C/A` >= 2) | (`Differential of Hoku PC6 C/A` <= 0.5 | `Differential of Hoku PC6 C/A` >= 2))

#Modify the gene names to be unique by adding numerical counts to the end of the gene name
counter_ <- 1 # initialize a counter
gene <- character(length(Pheatmap_Excel_1.0$`PG.Genes`)) # initialize a vector to store the modified gene names

for (i in 1:length(Pheatmap_Excel_1.0$`PG.Genes`)) { # iterate through the 'gene' column and modify the names
  if (i == 1 || Pheatmap_Excel_1.0$`PG.Genes`[i] != Pheatmap_Excel_1.0$`PG.Genes`[i - 1]) {
    counter_ <- 1  # reset counter when a new gene is encountered
  }
  gene[i] <- paste(Pheatmap_Excel_1.0$`PG.Genes`[i], counter_, sep = ".")
  counter_ <- counter_ + 1  # increment counter for each instance of the same gene
}

Pheatmap_Excel_1.0$gene <- gene # add the modified gene names to the dataframe
Pheatmap_Excel_1.0$`PG.Genes` <- NULL # remove the old column

Pheatmap_Excel_1.0 <- Pheatmap_Excel_1.0[c(ncol(Pheatmap_Excel_1.0), 1:(ncol(Pheatmap_Excel_1.0)-1))] # moving gene names to first column in dataset; CANNOT RUN THIS AGAIN OR WILL MESS UP ORDERING

#Setting up data for pheatmap
Pheatmap_Excel_2.0 <- as.data.frame(Pheatmap_Excel_1.0) # convert tibble to data frame
rownames(Pheatmap_Excel_2.0) <- Pheatmap_Excel_2.0[, 1] # setting the first column (gene names) as row names
Pheatmap_Excel_2.0 <- Pheatmap_Excel_2.0[, -1] # removing the first column from the data frame (since it's now row names)
Pheatmap_Excel_3.0 <- as.matrix(Pheatmap_Excel_2.0) # convert Data to a Numeric Matrix
Pheatmap_Excel_3.0[is.na(Pheatmap_Excel_3.0) | is.nan(Pheatmap_Excel_3.0) | is.infinite(Pheatmap_Excel_3.0)] <- 0 # replace NAs/NaNs/Inf with zero (otherwise pheatmap will not work)

#Group row annotations together at top of heatmap - THIS DOES NOT WORK BC THE ROWS ARE REARRANGED BY CLUSTERING WHEN PLOTTING PHEATMAP ANYWAYS
#inflammation_genes <- c("APOA1.1","APOA1.2", "CAMK1D.1", "CAMK1D.2", "CAMK1D.3", "CAMK1D.4", "CAMK1D.5", "CAMK1D.6","CLU.1", "FN1.1", "FN1.2", "FN1.3", "FN1.4", "FN1.5", "FN1.6", "GC.1", "GC.2", "GIT1.1", "GIT1.2", "GIT1.3", "GIT1.4", "GIT1.5", "GIT1.6","GPX1.1", "IL16.1", "IL16.2", "ITIH4.1", "ITIH4.2", "LAT.1", "LAT.2", "LAT.3", "PLCG2.1", "PLCG2.2", "PLCG2.3", "PLCG2.4", "PLCG2.5", "RAB44.1", "RAB44.2", "RAB44.3", "RAB44.4", "RHBDF2.1", "RHBDF2.2", "SELP.1", "SELP.2", "SYK.1", "SYK.2", "TMSB4X.1", "TMSB4X.2", "TMSB4X.3", "TMSB4X.4", "TMSB4X.5", "TMSB4X.6", "TMSB4X.7", "TMSB4X.8", "ZDHHC5.1", "ZDHHC5.2", "ZDHHC5.3", "ZDHHC5.4", "ZDHHC5.5", "ZDHHC5.6", "ZDHHC5.7", "ZDHHC5.8", "ZDHHC5.9", "ZDHHC5.10")
#Pheatmap_Excel_3.0_reordered <- Pheatmap_Excel_3.0[c(inflammation_genes, setdiff(rownames(Pheatmap_Excel_3.0), inflammation_genes)), ]

#Column annotations
annotation_col <- data.frame(
  Condition = factor(c("control", "control", "control", 
                       "control", "control", "preconditioning", "preconditioning")),
  Individual = factor(c("Kolohe", "Hoku", "Liho", "Nainoa", "Lono", "Kolohe", "Hoku"))
)
rownames(annotation_col) <- colnames(Pheatmap_Excel_3.0)

#Row annotations
annotation_row <- data.frame(
  Pathway = case_when(
    rownames(Pheatmap_Excel_3.0) %in% c("APOA1.1","APOA1.2", "C5", "CAMK1D.1", "CAMK1D.2", "CAMK1D.3", "CAMK1D.4", "CAMK1D.5", "CAMK1D.6","CLU.1", "FN1.1", "FN1.2", "FN1.3", "FN1.4", "FN1.5", "FN1.6", "GC.1", "GC.2", "GIT1.1", "GIT1.2", "GIT1.3", "GIT1.4", "GIT1.5", "GIT1.6","GPX1.1", "IL16.1", "IL16.2", "ITIH4.1", "ITIH4.2", "LAT.1", "LAT.2", "LAT.3", "PLAA", "PLCG2.1", "PLCG2.2", "PLCG2.3", "PLCG2.4", "PLCG2.5", "RAB44.1", "RAB44.2", "RAB44.3", "RAB44.4", "RHBDF2.1", "RHBDF2.2", "SELP.1", "SELP.2", "SYK.1", "SYK.2", "TMSB4X.1", "TMSB4X.2", "TMSB4X.3", "TMSB4X.4", "TMSB4X.5", "TMSB4X.6", "TMSB4X.7", "TMSB4X.8", "ZDHHC5.1", "ZDHHC5.2", "ZDHHC5.3", "ZDHHC5.4", "ZDHHC5.5", "ZDHHC5.6", "ZDHHC5.7", "ZDHHC5.8", "ZDHHC5.9", "ZDHHC5.10") ~ "Inflammation Response",
    rownames(Pheatmap_Excel_3.0) %in% c("BAG3.1", "BAG3.2", "BAG3.3", "BAG3.4", "BAG3.5", "BAG3.6", "CCDC88B.1", "CCDC88B.2", "CCDC88B.3", "CCDC88B.4", "CCDC88B.5", "CCDC88B.6", "CCDC88B.7", "CCDC88B.8", "CNP.1", "DOCK2.1", "FECH.1", "FGB.1", "HSF1.1", "HSF1.2", "HSF1.3", "LRCH4.1", "LRCH4.2", "MAPK1.1", "PDE12.1", "PDE12.2", "SERPINC1.1", "SLC2A1.1", "SLC2A1.2", "TREML1.1", "TREML1.2", "TREML1.3", "TREML1.4", ) ~ "Response to External Stimulus",
    TRUE ~ "No"  # Default category for other genes
  )
) # define specific annotations for certain genes
rownames(annotation_row) <- rownames(Pheatmap_Excel_3.0)

# Defining colors for annotations
annotation_colors <- list(
  Condition = c("preconditioning" = "blue", "control" = "red"),
  Pathway = c(`Inflammation Response` = "purple", `Response to External Stimulus` = "black", No = "gray")  
)

#Scaling
#Pheatmap_Excel_4.0 <- scale(Pheatmap_Excel_3.0)
#Add the above line of code and change to Pheatmap_Excel_4.0 in generate heatmap line of code

#Log transforming my data
Pheatmap_Excel_3.0_log <- log2(Pheatmap_Excel_3.0 + 1)

#Changing the coloring of the values
color_palette <- colorRampPalette(c("blue", "white", "red"))(100)  # 100 gradient colors in custom color palette
breaks <- unique(c(seq(0, 0.5, length.out = 30),  # Blue gradient (0 to 0.5)
                   seq(0.51, 2, length.out = 40), # White gradient (0.51 to 2)
                   seq(2.01, max(Pheatmap_Excel_3.0_log, na.rm = TRUE), length.out = 30))) # Red gradient (2.01 to max)
if (length(unique(breaks)) != length(breaks)) {
  warning("Breaks contain duplicate values, adjusting...")
  breaks <- sort(unique(breaks))  # Remove duplicates and sort
}

#Generate heatmap
pheatmap(Pheatmap_Excel_3.0_log, clustering_distance_rows = "euclidean", 
         clustering_distance_cols = "euclidean", color = color_palette, 
         breaks = breaks,
         clustering_method = "complete", show_rownames = FALSE, show_colnames = FALSE, scale = "none", annotation_col = annotation_col, annotation_row = annotation_row, annotation_colors = annotation_colors)

#change color scale
#make longer and skinnier
#add average column to force heat map to be arranged red to blue
#could filter dataset and to be more strict on which genes to include; maybe only filter for phosphosites (rows) that have a significant up/down regulation in both >= 1 control individuals and >= 1 preconditioning individual
```

#CONTROL CLUSTER 1 (PROFILE 7) STEM GRAPH
```{r}
# Libraries
library(ggplot2)
library(dplyr)
library(tidyr)

# STEM Data was imported

# Reformatted data into long format
Profile7 <- melt(DQO_STEM_Results_Control_Trial_Profile_7, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Profile7$Stage <- factor(Profile7$Stage, levels = unique(Profile7$Stage))
### Basic Plot for Profile 7 (Shows the dip at Pre-Met)
ggplot(Profile7, aes(x= Stage, y=n, group=`Protein Names`, color=`Protein Names`)) + ylim(-6000,2000) + #should be min and max of data
  geom_line() +
  labs(title = "Profile 7",
       x = "Time Point",
       y = "Fold Change",
       color = "Gene") +
  theme_minimal()
View(Profile7)

# Highlight AVERAGE in Profile7
highlight_gene <- c("AVERAGE")

#Vector for Group
group_vector_Profile7 <- ifelse(Profile7$`Protein Names` %in% highlight_gene, "Average", "Other")
View(group_vector_Profile7)
df_Profile7 <- as.data.frame(Profile7)
df_Profile7$Color <- c(group_vector_Profile7)
colnames(df_Profile7) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Profile7)
# Ensure that the group levels are factors for colors and opacity
#Vector for Opacity
opacity_values <- c("Average" = 1.0, "Other" = 0.25)
df_Profile7$Group <- factor(df_Profile7$Group, levels = names(opacity_values))
#Vector for Colors
color_values <- c("Average" = "red", "Other" = "grey")
df_Profile7$Group <- factor(df_Profile7$Group, levels = names(color_values))
# Make x-axis values numeric to change spacing
df_Profile7$Stage_num <- recode(df_Profile7$Stage, "A" = 0, "C" = 14, "D" = 19)
# Plot the data with the correct mapping
Profile7_Graph <- ggplot(df_Profile7, aes(x= Stage_num, y= FC, group = `Protein Names`, alpha = Group, color = Group)) + ylim(-7000,10000) +
  geom_line(size = 0.25) +
  geom_line(data = subset(df_Profile7, df_Profile7$Group == "Average"), aes(color = Group), 
            size = 0.75) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .5) +
  geom_vline(xintercept = 0, linetype = 5, color = "black", size = .5) +
  scale_color_manual(values = c("Others" = "grey", "Average" = "red")) +
  scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D")) +  # Uneven spacing
  labs(title = "Profile 7",
       x = "Time Point",
       y = "Fold Change",
       color = "Protein Names") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),  # Center title
        axis.text.x = element_text(size = 14),  # Adjusts the text size of the x-axis tick labels
        axis.text.y = element_text(size = 14),  # Adjusts the text size of the y-axis tick labels
        axis.title.x = element_text(size = 14), # Adjusts the text size of the x-axis title
        axis.title.y = element_text(size = 14), # Adjusts the text size of the y-axis title
        legend.text = element_text(size = 10),
        legend.title = element_blank(),
        panel.grid = element_blank()) # Removes gridlines
Profile7_Graph + guides(alpha = guide_none())
```

#CONTROL CLUSTER 2 (PROFILE 2) STEM GRAPH
```{r}

# STEM Data was imported

# Reformatted data into long format
Profile2 <- melt(DQO_STEM_Results_Control_Trial_Profile_2, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Profile2$Stage <- factor(Profile2$Stage, levels = unique(Profile2$Stage))
### Basic Plot for Profile 7 (Shows the dip at Pre-Met)
ggplot(Profile2, aes(x= Stage, y=n, group=`Protein Names`, color=`Protein Names`)) + ylim(-7000,100) + #should be min and max of data
  geom_line() +
  labs(title = "Profile 2",
       x = "Time Point",
       y = "Fold Change",
       color = "Gene") +
  theme_minimal()
View(Profile2)

# Highlight AVERAGE in Profile7
highlight_gene <- c("AVERAGE")

#Vector for Group
group_vector_Profile2 <- ifelse(Profile2$`Protein Names` %in% highlight_gene, "Average", "Other")
View(group_vector_Profile2)
df_Profile2 <- as.data.frame(Profile2)
df_Profile2$Color <- c(group_vector_Profile2)
colnames(df_Profile2) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Profile2)
# Ensure that the group levels are factors for colors and opacity
#Vector for Opacity
opacity_values <- c("Average" = 1.0, "Other" = 0.5)
df_Profile2$Group <- factor(df_Profile2$Group, levels = names(opacity_values))
#Vector for Colors
color_values <- c("Average" = "red", "Other" = "grey")
df_Profile2$Group <- factor(df_Profile2$Group, levels = names(color_values))
# Make x-axis values numeric to change spacing
df_Profile2$Stage_num <- recode(df_Profile2$Stage, "A" = 0, "C" = 14, "D" = 19)
# Plot the data with the correct mapping
Profile2_Graph <- ggplot(df_Profile2, aes(x= Stage_num, y= FC, group = `Protein Names`, alpha = Group, color = Group)) + ylim(-7000,10000) +
  geom_line(size = 0.25) +
  geom_line(data = subset(df_Profile2, df_Profile2$Group == "Average"), aes(color = Group), 
            size = 0.75) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .5) +
  geom_vline(xintercept = 0, linetype = 5, color = "black", size = .5) +
  scale_color_manual(values = c("Others" = "grey", "Average" = "red")) +
  scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D")) +  # Uneven spacing
  labs(title = "Profile 2",
       x = "Time Point",
       y = "Fold Change",
       color = "Protein Names") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),  # Center title
        axis.text.x = element_text(size = 14),  # Adjusts the text size of the x-axis tick labels
        axis.text.y = element_text(size = 14),  # Adjusts the text size of the y-axis tick labels
        axis.title.x = element_text(size = 14), # Adjusts the text size of the x-axis title
        axis.title.y = element_text(size = 14), # Adjusts the text size of the y-axis title
        legend.text = element_text(size = 10),
        legend.title = element_blank(),
        panel.grid = element_blank()) # Removes gridlines
Profile2_Graph + guides(alpha = guide_none())
```

#CONTROL CLUSTER 3 (PROFILE 3) STEM GRAPH
```{r}

# STEM Data was imported

# Reformatted data into long format
Profile3 <- melt(DQO_STEM_Results_Control_Trial_Profile_3, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Profile3$Stage <- factor(Profile3$Stage, levels = unique(Profile3$Stage))
### Basic Plot for Profile 7 (Shows the dip at Pre-Met)
ggplot(Profile3, aes(x= Stage, y=n, group=`Protein Names`, color=`Protein Names`)) + ylim(-3000,100) + #should be min and max of data
  geom_line() +
  labs(title = "Profile 3",
       x = "Time Point",
       y = "Fold Change",
       color = "Gene") +
  theme_minimal()
View(Profile3)

# Highlight AVERAGE in Profile7
highlight_gene <- c("AVERAGE")

#Vector for Group
group_vector_Profile3 <- ifelse(Profile3$`Protein Names` %in% highlight_gene, "Average", "Other")
View(group_vector_Profile3)
df_Profile3 <- as.data.frame(Profile3)
df_Profile3$Color <- c(group_vector_Profile3)
colnames(df_Profile3) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Profile3)
# Ensure that the group levels are factors for colors and opacity
#Vector for Opacity
opacity_values <- c("Average" = 1.0, "Other" = 0.5)
df_Profile3$Group <- factor(df_Profile3$Group, levels = names(opacity_values))
#Vector for Colors
color_values <- c("Average" = "red", "Other" = "grey")
df_Profile3$Group <- factor(df_Profile3$Group, levels = names(color_values))
# Make x-axis values numeric to change spacing
df_Profile3$Stage_num <- recode(df_Profile3$Stage, "A" = 0, "C" = 14, "D" = 19)
# Plot the data with the correct mapping
Profile3_Graph <- ggplot(df_Profile3, aes(x= Stage_num, y= FC, group = `Protein Names`, alpha = Group, color = Group)) + ylim(-7000,10000) +
  geom_line(size = 0.25) +
  geom_line(data = subset(df_Profile3, df_Profile3$Group == "Average"), aes(color = Group), 
            size = 0.75) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .5) +
  geom_vline(xintercept = 0, linetype = 5, color = "black", size = .5) +
  scale_color_manual(values = c("Others" = "grey", "Average" = "red")) +
  scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D")) +  # Uneven spacing
  labs(title = "Profile 3",
       x = "Time Point",
       y = "Fold Change",
       color = "Protein Names") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),  # Center title
        axis.text.x = element_text(size = 14),  # Adjusts the text size of the x-axis tick labels
        axis.text.y = element_text(size = 14),  # Adjusts the text size of the y-axis tick labels
        axis.title.x = element_text(size = 14), # Adjusts the text size of the x-axis title
        axis.title.y = element_text(size = 14), # Adjusts the text size of the y-axis title
        legend.text = element_text(size = 10),
        legend.title = element_blank(),
        panel.grid = element_blank()) # Removes gridlines
Profile3_Graph + guides(alpha = guide_none())
```

#CONTROL CLUSTER 4 (PROFILE 8) STEM GRAPH
```{r}

# STEM Data was imported

# Reformatted data into long format
Profile8 <- melt(DQO_STEM_Results_Control_Trial_Profile_8, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Profile8$Stage <- factor(Profile8$Stage, levels = unique(Profile8$Stage))
### Basic Plot for Profile 7 (Shows the dip at Pre-Met)
ggplot(Profile8, aes(x= Stage, y=n, group=`Protein Names`, color=`Protein Names`)) + ylim(-1000,7000) + #should be min and max of data
  geom_line() +
  labs(title = "Profile 8",
       x = "Time Point",
       y = "Fold Change",
       color = "Gene") +
  theme_minimal()
View(Profile8)

# Highlight AVERAGE in Profile7
highlight_gene <- c("AVERAGE")

#Vector for Group
group_vector_Profile8 <- ifelse(Profile8$`Protein Names` %in% highlight_gene, "Average", "Other")
View(group_vector_Profile8)
df_Profile8 <- as.data.frame(Profile8)
df_Profile8$Color <- c(group_vector_Profile8)
colnames(df_Profile8) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Profile8)
# Ensure that the group levels are factors for colors and opacity
#Vector for Opacity
opacity_values <- c("Average" = 1.0, "Other" = 0.5)
df_Profile8$Group <- factor(df_Profile8$Group, levels = names(opacity_values))
#Vector for Colors
color_values <- c("Average" = "red", "Other" = "grey")
df_Profile8$Group <- factor(df_Profile8$Group, levels = names(color_values))
# Make x-axis values numeric to change spacing
df_Profile8$Stage_num <- recode(df_Profile8$Stage, "A" = 0, "C" = 14, "D" = 19)
# Plot the data with the correct mapping
Profile8_Graph <- ggplot(df_Profile8, aes(x= Stage_num, y= FC, group = `Protein Names`, alpha = Group, color = Group)) + ylim(-7000,10000) +
  geom_line(size = 0.25) +
  geom_line(data = subset(df_Profile8, df_Profile8$Group == "Average"), aes(color = Group), 
            size = 0.75) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .5) +
  geom_vline(xintercept = 0, linetype = 5, color = "black", size = .5) +
  scale_color_manual(values = c("Others" = "grey", "Average" = "red")) +
  scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D")) +  # Uneven spacing
  labs(title = "Profile 8",
       x = "Time Point",
       y = "Fold Change",
       color = "Protein Names") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),  # Center title
        axis.text.x = element_text(size = 14),  # Adjusts the text size of the x-axis tick labels
        axis.text.y = element_text(size = 14),  # Adjusts the text size of the y-axis tick labels
        axis.title.x = element_text(size = 14), # Adjusts the text size of the x-axis title
        axis.title.y = element_text(size = 14), # Adjusts the text size of the y-axis title
        legend.text = element_text(size = 10),
        legend.title = element_blank(),
        panel.grid = element_blank()) # Removes gridlines
Profile8_Graph + guides(alpha = guide_none())
```

#CONTROL CLUSTER 5 (PROFILE 13) STEM GRAPH
```{r}

# STEM Data was imported

# Reformatted data into long format
Profile13 <- melt(DQO_STEM_Results_Control_Trial_Profile_13, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Profile13$Stage <- factor(Profile13$Stage, levels = unique(Profile13$Stage))
### Basic Plot for Profile 7 (Shows the dip at Pre-Met)
ggplot(Profile13, aes(x= Stage, y=n, group=`Protein Names`, color=`Protein Names`)) + ylim(-100,8000) + #should be min and max of data
  geom_line() +
  labs(title = "Profile 13",
       x = "Time Point",
       y = "Fold Change",
       color = "Gene") +
  theme_minimal()
View(Profile13)

# Highlight AVERAGE in Profile7
highlight_gene <- c("AVERAGE")

#Vector for Group
group_vector_Profile13 <- ifelse(Profile13$`Protein Names` %in% highlight_gene, "Average", "Other")
View(group_vector_Profile13)
df_Profile13 <- as.data.frame(Profile13)
df_Profile13$Color <- c(group_vector_Profile13)
colnames(df_Profile13) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Profile13)
# Ensure that the group levels are factors for colors and opacity
#Vector for Opacity
opacity_values <- c("Average" = 1.0, "Other" = 0.5)
df_Profile13$Group <- factor(df_Profile13$Group, levels = names(opacity_values))
#Vector for Colors
color_values <- c("Average" = "red", "Other" = "grey")
df_Profile13$Group <- factor(df_Profile13$Group, levels = names(color_values))
# Make x-axis values numeric to change spacing
df_Profile13$Stage_num <- recode(df_Profile13$Stage, "A" = 0, "C" = 14, "D" = 19)
# Plot the data with the correct mapping
Profile13_Graph <- ggplot(df_Profile13, aes(x= Stage_num, y= FC, group = `Protein Names`, alpha = Group, color = Group)) + ylim(-7000,10000) +
  geom_line(size = 0.25) +
  geom_line(data = subset(df_Profile13, df_Profile13$Group == "Average"), aes(color = Group), 
            size = 0.5) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .75) +
  geom_vline(xintercept = 0, linetype = 5, color = "black", size = .5) +
  scale_color_manual(values = c("Others" = "grey", "Average" = "red")) +
  scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D")) +  # Uneven spacing
  labs(title = "Profile 13",
       x = "Time Point",
       y = "Fold Change",
       color = "Protein Names") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),  # Center title
        axis.text.x = element_text(size = 14),  # Adjusts the text size of the x-axis tick labels
        axis.text.y = element_text(size = 14),  # Adjusts the text size of the y-axis tick labels
        axis.title.x = element_text(size = 14), # Adjusts the text size of the x-axis title
        axis.title.y = element_text(size = 14), # Adjusts the text size of the y-axis title
        legend.text = element_text(size = 10),
        legend.title = element_blank(),
        panel.grid = element_blank()) # Removes gridlines
Profile13_Graph + guides(alpha = guide_none())
```

#PRECONDITIONING PROFILE 14 STEM GRAPH
```{r}
library(reshape2)

# STEM Data was imported

# Reformatted data into long format
Profile14 <- melt(DQO_STEM_Results_PC6_Trial_Profile_14, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Profile14$Stage <- factor(Profile14$Stage, levels = unique(Profile14$Stage))
### Basic Plot for Profile 7 (Shows the dip at Pre-Met)
ggplot(Profile14, aes(x= Stage, y=n, group=`Protein Names`, color=`Protein Names`)) + ylim(-100,10000) + #should be min and max of data
  geom_line() +
  labs(title = "Profile 14",
       x = "Time Point",
       y = "Fold Change",
       color = "Gene") +
  theme_minimal()
View(Profile14)

# Highlight AVERAGE in Profile7
highlight_gene <- c("AVERAGE")

#Vector for Group
group_vector_Profile14 <- ifelse(Profile14$`Protein Names` %in% highlight_gene, "Average", "Other")
View(group_vector_Profile14)
df_Profile14 <- as.data.frame(Profile14)
df_Profile14$Color <- c(group_vector_Profile14)
colnames(df_Profile14) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Profile14)
# Ensure that the group levels are factors for colors and opacity
#Vector for Opacity
opacity_values <- c("Average" = 1.0, "Other" = 0.5)
df_Profile14$Group <- factor(df_Profile14$Group, levels = names(opacity_values))
#Vector for Colors
color_values <- c("Average" = "blue", "Other" = "grey")
df_Profile14$Group <- factor(df_Profile14$Group, levels = names(color_values))
# Make x-axis values numeric to change spacing
df_Profile14$Stage_num <- recode(df_Profile14$Stage, "A" = 0, "C" = 14, "D" = 19)
# Plot the data with the correct mapping
Profile14_Graph <- ggplot(df_Profile14, aes(x= Stage_num, y= FC, group = `Protein Names`, alpha = Group, color = Group)) + ylim(-7000,10000) +
  geom_line(size = 0.25) +
  geom_line(data = subset(df_Profile14, df_Profile14$Group == "Average"), aes(color = Group), 
            size = 0.75) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .5) +
  geom_vline(xintercept = 0, linetype = 5, color = "black", size = .5) +
  scale_color_manual(values = c("Others" = "grey", "Average" = "blue")) +
  scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D")) +  # Uneven spacing
  labs(title = "Profile 14",
       x = "Time Point",
       y = "Fold Change",
       color = "Gene Symbol") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),  # Center title
        axis.text.x = element_text(size = 14),  # Adjusts the text size of the x-axis tick labels
        axis.text.y = element_text(size = 14),  # Adjusts the text size of the y-axis tick labels
        axis.title.x = element_text(size = 14), # Adjusts the text size of the x-axis title
        axis.title.y = element_text(size = 14), # Adjusts the text size of the y-axis title
        legend.text = element_text(size = 10),
        legend.title = element_blank(),
        panel.grid = element_blank()) # Removes gridlines
Profile14_Graph + guides(alpha = guide_none())
```

#PRECONDITIONING PROFILE 10 STEM GRAPH
```{r}
library(reshape2)

# STEM Data was imported

# Reformatted data into long format
Profile10 <- melt(DQO_STEM_Results_PC6_Trial_Profile_10, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Profile10$Stage <- factor(Profile10$Stage, levels = unique(Profile10$Stage))
### Basic Plot for Profile 7 (Shows the dip at Pre-Met)
ggplot(Profile10, aes(x= Stage, y=n, group=`Protein Names`, color=`Protein Names`)) + ylim(-2000,6000) + #should be min and max of data
  geom_line() +
  labs(title = "Profile 10",
       x = "Time Point",
       y = "Fold Change",
       color = "Gene") +
  theme_minimal()
View(Profile10)

# Highlight AVERAGE in Profile10
highlight_gene <- c("AVERAGE")

#Vector for Group
group_vector_Profile10 <- ifelse(Profile10$`Protein Names` %in% highlight_gene, "Average", "Other")
View(group_vector_Profile10)
df_Profile10 <- as.data.frame(Profile10)
df_Profile10$Color <- c(group_vector_Profile10)
colnames(df_Profile10) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Profile10)
# Ensure that the group levels are factors for colors and opacity
#Vector for Opacity
opacity_values <- c("Average" = 1.0, "Other" = 0.5)
df_Profile10$Group <- factor(df_Profile10$Group, levels = names(opacity_values))
#Vector for Colors
color_values <- c("Average" = "blue", "Other" = "grey")
df_Profile10$Group <- factor(df_Profile10$Group, levels = names(color_values))
# Make x-axis values numeric to change spacing
df_Profile10$Stage_num <- recode(df_Profile10$Stage, "A" = 0, "C" = 14, "D" = 19)
# Plot the data with the correct mapping
Profile10_Graph <- ggplot(df_Profile10, aes(x= Stage_num, y= FC, group = `Protein Names`, alpha = Group, color = Group)) + ylim(-7000,10000) +
  geom_line(size = 0.25) +
  geom_line(data = subset(df_Profile10, df_Profile10$Group == "Average"), aes(color = Group), 
            size = 0.75) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .5) +
  geom_vline(xintercept = 0, linetype = 5, color = "black", size = .5) +
  scale_color_manual(values = c("Others" = "grey", "Average" = "blue")) +
  scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D")) +  # Uneven spacing
  labs(title = "Profile 10",
       x = "Time Point",
       y = "Fold Change",
       color = "Gene Symbol") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),  # Center title
        axis.text.x = element_text(size = 14),  # Adjusts the text size of the x-axis tick labels
        axis.text.y = element_text(size = 14),  # Adjusts the text size of the y-axis tick labels
        axis.title.x = element_text(size = 14), # Adjusts the text size of the x-axis title
        axis.title.y = element_text(size = 14), # Adjusts the text size of the y-axis title
        legend.text = element_text(size = 10),
        legend.title = element_blank(),
        panel.grid = element_blank()) # Removes gridlines
Profile10_Graph + guides(alpha = guide_none())
```

#PRECONDITIONING CLUSTER 1 (PROFILE 14 + 10) STEM GRAPH
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)

# STEM Data was imported

# Reformatted data into long format
Cluster1PC6 <- melt(DQO_STEM_Results_PC6_Trial_Cluster_1_Profile_14_10_, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Cluster1PC6$Stage <- factor(Cluster1PC6$Stage, levels = unique(Cluster1PC6$Stage))
### Basic Plot for Profile 7 (Shows the dip at Pre-Met)
#ggplot(Cluster1PC6, aes(x= Stage, y=n, group=`Protein Names`, color=`Protein Names`)) + ylim(-7000,10000) + #should be min and max of data
#  geom_line() +
#  labs(x = "Time Point",
#       y = "Fold Change",
#       color = "Gene") +
#  theme_minimal()
#View(Cluster1PC6)

# Highlight AVERAGE in Cluster1PC6
highlight_gene <- c("AVERAGE")

#Vector for Group
group_vector_Cluster1PC6 <- ifelse(Cluster1PC6$`Protein Names` %in% highlight_gene, "Average", "Other")
View(group_vector_Cluster1PC6)
df_Cluster1PC6 <- as.data.frame(Cluster1PC6)
df_Cluster1PC6$Color <- c(group_vector_Cluster1PC6)
colnames(df_Cluster1PC6) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Cluster1PC6)
# Ensure that the group levels are factors for colors and opacity
#Vector for Opacity
#opacity_values <- c("Average" = 1.0, "Other" = 0.5)
#df_Cluster1PC6$Group <- factor(df_Cluster1PC6$Group, levels = names(opacity_values))
#Vector for Colors
#color_values <- c("Average" = "blue", "Other" = "grey")
#df_Cluster1PC6$Group <- factor(df_Cluster1PC6$Group, levels = names(color_values))
# Make x-axis values numeric to change spacing
df_Cluster1PC6$Stage_num <- recode(df_Cluster1PC6$Stage, "A" = 0, "C" = 14, "D" = 19)
# Create ranges for 95% ribbon
#df_Cluster1PC6$ymin_Cluster1PC6 <- #ifelse(df_Cluster1PC6$Stage == "A", 0, 
#                  ifelse(df_Cluster1PC6$Stage == "C", #89.41, 
#                         ifelse(df_Cluster1PC6$Stage == #"D", -127.93, NA)))
#
#df_Cluster1PC6$ymax_Cluster1PC6 <- #ifelse(df_Cluster1PC6$Stage == "A", 0, 
#                  ifelse(df_Cluster1PC6$Stage == "C", #2977.67, 
#                         ifelse(df_Cluster1PC6$Stage == #"D", 694.04, NA)))
# Create ranges for 25% ribbon
df_Cluster1PC6$ymin_25_Cluster1PC6 <- ifelse(df_Cluster1PC6$Stage == "A", 0, 
                  ifelse(df_Cluster1PC6$Stage == "C", 441.79, 
                         ifelse(df_Cluster1PC6$Stage == "D", 115.39, NA)))

df_Cluster1PC6$ymax_25_Cluster1PC6 <- ifelse(df_Cluster1PC6$Stage == "A", 0, 
                  ifelse(df_Cluster1PC6$Stage == "C", 736.85, 
                         ifelse(df_Cluster1PC6$Stage == "D", 287.245, NA)))
# Create ranges for 50% ribbon
df_Cluster1PC6$ymin_50_Cluster1PC6 <- ifelse(df_Cluster1PC6$Stage == "A", 0, 
                  ifelse(df_Cluster1PC6$Stage == "C", 310.8, 
                         ifelse(df_Cluster1PC6$Stage == "D", 54.7125, NA)))

df_Cluster1PC6$ymax_50_Cluster1PC6 <- ifelse(df_Cluster1PC6$Stage == "A", 0, 
                  ifelse(df_Cluster1PC6$Stage == "C", 981.29, 
                         ifelse(df_Cluster1PC6$Stage == "D", 440.8975, NA)))
# Create ranges for 75% ribbon
df_Cluster1PC6$ymin_75_Cluster1PC6 <- ifelse(df_Cluster1PC6$Stage == "A", 0, 
                  ifelse(df_Cluster1PC6$Stage == "C", 193.57, 
                         ifelse(df_Cluster1PC6$Stage == "D", 0.6, NA)))

df_Cluster1PC6$ymax_75_Cluster1PC6 <- ifelse(df_Cluster1PC6$Stage == "A", 0, 
                  ifelse(df_Cluster1PC6$Stage == "C", 1504.43, 
                         ifelse(df_Cluster1PC6$Stage == "D", 699.36125, NA)))
# Create ranges for 100% ribbon
df_Cluster1PC6$ymin_100_Cluster1PC6 <- ifelse(df_Cluster1PC6$Stage == "A", 0, 
                  ifelse(df_Cluster1PC6$Stage == "C", 11.18, 
                         ifelse(df_Cluster1PC6$Stage == "D", -1050.14, NA)))

df_Cluster1PC6$ymax_100_Cluster1PC6 <- ifelse(df_Cluster1PC6$Stage == "A", 0, 
                  ifelse(df_Cluster1PC6$Stage == "C", 9770.15, 
                         ifelse(df_Cluster1PC6$Stage == "D", 4952.07, NA)))
# Log transform data
df_Cluster1PC6_log <- df_Cluster1PC6 %>%
  mutate(FC_log = sign(FC) * log(abs(FC) + 1))
# Plot the data with the correct mapping
Cluster1PC6_Graph <- ggplot(df_Cluster1PC6_log, aes(x= Stage_num, y= FC_log, group = `Protein Names`, alpha = Group, color = Group)) + ylim(-10,10) +
  geom_ribbon(aes(ymin = (sign(ymin_100_Cluster1PC6) * log(abs(ymin_100_Cluster1PC6) + 1)), ymax = (sign(ymax_100_Cluster1PC6) * log(abs(ymax_100_Cluster1PC6) + 1))), fill = "#E5E5E5", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_75_Cluster1PC6) * log(abs(ymin_75_Cluster1PC6) + 1)), ymax = (sign(ymax_75_Cluster1PC6) * log(abs(ymax_75_Cluster1PC6) + 1))), fill = "#D3D3D3", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_50_Cluster1PC6) * log(abs(ymin_50_Cluster1PC6) + 1)), ymax = (sign(ymax_50_Cluster1PC6) * log(abs(ymax_50_Cluster1PC6) + 1))), fill = "#A0A0A0", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_25_Cluster1PC6) * log(abs(ymin_25_Cluster1PC6) + 1)), ymax = (sign(ymax_25_Cluster1PC6) * log(abs(ymax_25_Cluster1PC6) + 1))), fill = "#808080", size = 0) + # Shaded region
  geom_line(size = 0.05, color = "black") +
  geom_line(data = subset(df_Cluster1PC6_log, df_Cluster1PC6_log$Group == "Average"), 
            size = 3, alpha = 1) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .5) +
  geom_vline(xintercept = 0, linetype = 1, color = "black", size = .5) +
  geom_hline(yintercept = -10, linetype = 1, color = "black", size = .5) +
  scale_color_manual(values = c("Average" = "blue")) +
  #scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D")) +  # Uneven spacing
  labs(x = "Time Point",
       y = "Log(Fold Change)",
       color = "Gene Symbol") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),  # Center title
        axis.text.x = element_text(size = 14),  # Adjusts the text size of the x-axis tick labels
        axis.text.y = element_text(size = 14),  # Adjusts the text size of the y-axis tick labels
        axis.title.x = element_text(size = 14), # Adjusts the text size of the x-axis title
        axis.title.y = element_text(size = 14), # Adjusts the text size of the y-axis title
        legend.text = element_text(size = 10),
        legend.title = element_blank(),
        panel.grid = element_blank()) # Removes gridlines
Cluster1PC6_Graph + guides(alpha = guide_none())
```

```{r}
#https://www.programiz.com/python-programming/online-compiler/

# Your dataset
data <- c(421.26, 237.59, 179.71, 679.75, 466.11, 433.85, 65.05, 178.22, 384.81, 854.28, 245.78, 601.73, 214.58, 620.41, 452.81, 1657.11, 478.3, 540.85, 727.48, 482.72, 11.18, 97.46, 229.25, 1005.80, 275.69, 192.06, 107.99, 301.55, 348.66, 332.81, 163.22, 101.34, 269.01, 376.76, 149.84, 352.06, 247.74, 189.47, 606.07, 85.37, 162.43, 439.52, 114.74, 1138.96, 761.03, 1701.57, 410.97, 179.57, 151.01, 852.13, 161.97, 514.31, 529.28, 1209.78, 2993.74, 417.32, 224.34, 496.69, 163.54, 2037.34, 604.9, 171.02, 805.21, 871.19, 76.02, 477.2, 789.45, 586.83, 806.04, 623.86, 553.81, 418.23, 242.12, 177.34, 451.39, 327.38, 1107.69, 646.26, 205.75, 269.25, 243.87, 198.42, 938.19, 1017.69, 388.61, 356.39, 480.02, 173.1, 76.71, 145.03, 1164.04, 150.36, 267.31, 106.58, 580.58, 759.35, 147.18, 809.06, 812.37, 260.06, 217.12, 427.11, 800.73, 339.6, 350.79, 526.86, 266.17, 485.21, 245, 477.87, 368.65, 88.9, 283.52, 416.53, 205.67, 621.19, 532.32, 329.04, 223.02, 93.02, 412.86, 413.71, 246.1, 279.22, 404.44, 5501.91, 1168.85, 1304.04, 279.36, 366.56, 404.99, 339.02, 1023.73, 316.31, 778.14, 338.29, 710.19, 2922.32, 3073.64, 629.83, 318.94, 549.54, 108.51, 613.7, 198.91, 193.57, 155.84, 315.47, 100.47, 229.52, 91.15, 284.75, 391.48, 526.01, 760.65, 523.59, 451.04, 626.46, 145.9, 972.08, 2258.59, 254.28, 1561.06, 581.72, 844.04, 152.13, 142.73, 979.25, 1283.42, 731.87, 514.5, 305.32, 2915.09, 741.49, 277.67, 458.06, 171.95, 211.93, 105.24, 279.67, 630.01, 292.22, 687.67, 120.09, 484.37, 952.28, 1778.48, 509.6, 441.87, 183.99, 896.55, 379.49, 1045.39, 305.29, 225.52, 296.03, 1739.79, 866.75, 337.05, 444.36, 521.3, 1602.20, 1325.42, 8656.21, 1222.26, 156.8, 6150.52, 529.15, 1609.76, 402.09, 803.52, 990.81, 1889.69, 1564.52, 3065.03, 263.7, 539.25, 1895.52, 1783.07, 279.27, 1172.95, 183.1, 510.79, 4518.95, 247.27, 215.21, 745.4, 372.17, 453.58, 535.04, 1391.12, 476.48, 174.9, 675.54, 2354.71, 556.71, 1082.22, 837.91, 1465.72, 84.75, 1123.17, 712.7, 1401.77, 1057.32, 511.01, 1151.99, 403.76, 268.32, 1177.79, 725.88, 876.46, 874.76, 201.54, 198.15, 944.25, 1189.49, 352.74, 311.66, 310.8, 220.12, 2269.04, 632.47, 397.01, 472.85, 581.48, 1154.40, 504.6, 134.09, 695, 3532.57, 653.92, 1414.49, 1275.25, 349.92, 436.97, 1759.54, 1432.09, 705.48, 665.18, 497.59, 194.55, 177.57, 350.12, 1559.64, 269.46, 1072.28, 430.58, 238.36, 529.53, 1116.82, 592.77, 753.33, 756.51, 329.32, 878.26, 577.91, 883.48, 323.34, 659.77, 1071.24, 602.75, 836.51, 668.11, 707.46, 4048.87, 7518.83, 3042.31, 413.92, 2386.21, 1003.57, 8236.35, 6486.76, 1695.59, 931.84, 1527.60, 897.38, 416.48, 657.82, 1053.15, 1268.89, 151.97, 164.24, 2801.24, 600.02, 727.54, 2454.98, 966.32, 531.04, 1796.73, 450.88, 231.77, 4019.03, 186.29, 1222.22, 634.13, 1313.66, 569.82, 404.06, 664.12, 958.3, 533.4, 352.94, 476.63, 2177.00, 1548.27, 4148.04, 1110.02, 1894.17, 471.38, 1025.79, 9770.15, 2394.82, 1278.38, 4221.70, 1924.14, 806.92, 572.38, 3633.06, 327.72, 305.67, 743.6, 737.42, 2582.58, 289.96, 331.97, 495.14, 505.77, 728.67, 867.8, 289.81, 420.98, 599.62, 1036.40, 410.11, 406.86, 996.21, 2451.25, 102.3, 546.59, 1628.23, 284.86, 2538.34, 1548.80, 213.53, 289.23, 876.92, 564.91, 342.74, 1200.72, 494.69, 648.32, 451.21, 859.83, 147.34, 811.5, 825.91, 728.94, 169.83, 3169.59, 186.17, 65.47, 773.65, 1350.25, 1327.56, 677.09, 699.25, 4514.35, 443.29, 191.45, 426.63, 1504.43, 1128.74, 420.97, 126.12, 1795.43, 2435.47, 290.22, 465.48, 487.91, 612.47, 441.79, 565.28, 732.27, 616.59, 560.7, 433.5, 46.04, 509.7, 312.04, 981.29, 1136.68, 429.62, 106.97, 26.08, 258.01, 274.52, 1079.01, 287.68, 670.78, 543.44, 709.77, 1377.01, 1968.24, 837.81, 661.8, 2483.07, 2610.86, 986.64, 487.29, 788.62, 732.38, 987.18, 209.19, 157.73, 557.24, 505.73, 751.5, 752.95, 929.26, 1091.89, 513.5, 2066.67, 152.45, 930.57, 952.41, 1227.32, 568.43, 674.01, 1261.83, 455.75, 616.24, 736.85, 701.65, 576.1, 1045.34, 816.47, 567.95, 2888.78, 71.1, 202.82, 360.04, 718.65, 2553.20, 237.45, 816.19, 1055.94, 883.52, 646.24, 552.93, 408.13, 648.81, 715.71, 173.67, 747.12, 936.08, 1219.67, 1090.28, 530.11, 1031.24, 1112.69, 492.59, 291.04, 608.88, 380.09, 1317.77, 830.69, 303.48, 261.68, 423.5, 290.15, 991.47, 362.78, 567.87, 319.27, 945.22, 861.41, 1867.03, 925.1, 623.19, 3022.17, 534.11, 812.58, 822.02, 3386.63, 2983.37, 1203.07, 1152.84, 592.53, 1122.60, 808.16, 123.88, 1502.93, 185.81, 187.51, 912.41, 391.18, 1395.19, 259.05, 6034.12, 537.75, 877.85, 124.87, 820.3, 140.54, 1081.39)

# Calculate Q1 and Q3
q1_25 <- quantile(data, 0.125)
q3_75 <- quantile(data, 0.875)

# Output results
q1_25
q3_75

data_1 <- c(87.96, 22.59, 42.6, 45.98, -20.09, -39.45, 1.44, -8.38, 28.44, 141.08, -87.37, 25.71, 31.41, 35.25, 117.14, 161.17, 40.21, 48.54, 72.65, 123.42, -0.95, -17.21, -80.68, 242.57, -93.52, -14.99, -33.29, -53.13, 91.99, 14.49, 32.99, 10.02, 29.21, -47.96, 14.03, -20.47, 33.86, 16.85, -36.72, 17.22, -54.42, 90.32, -28.45, 280.46, 65.33, 440.96, -148.22, -46.58, 11.03, -91.72, 38.88, 117.33, 32.69, 200.42, 719.01, -136.59, -54.82, -1.07, -5.53, 504.87, -22.27, 1.67, 157.6, -73.06, 11.62, 94.88, 178.08, -10.81, 213.01, 118.37, 19.06, 19.47, 16.98, 11.13, 13.17, -85.41, 65.47, 146.68, -54.95, -83.67, 28.36, -24.39, 187.68, 135.74, -25.44, -80.71, 39.43, -4.17, 16.55, -42.15, 119.38, -1.16, -61.71, 12.54, 21.75, 93.49, 35.35, 24.67, 180.78, 29.67, -1.58, -40.46, 159.64, 33.34, 40.63, 37.72, -16.66, 101.91, 47.81, -48.81, 69.08, 0.48, 23.05, 100.99, -24.53, 96.57, 37.79, 72.25, -7.12, -22.96, 98.01, -13.49, -63.15, -20.61, 103.97, -1050.14, -94.03, 334.38, 19.05, 53.52, 41.63, -29.36, 231.9, 83.24, -32.39, 59.46, 66.13, 770.42, -76.58, 100.71, 73.91, -37.56, 13.39, -21.22, 12.49, 38.8, 15.54, -8.03, -7.97, -4.44, -11.64, -47.46, -132.59, 105.87, 1.5, -4.9, 40.73, -105.95, -6.36, 180.1, 480.95, -24.28, 296.84, 134.6, 125.21, 25.17, 18.88, 144.67, 231.9, -1.64, -166.85, 32.1, 692.59, 89.61, -51.84, 51.84, 34.3, 9.43, 25.09, 29.83, -14.23, 15.39, 155.27, 26.24, -145.51, 249.17, 409.13, 13.27, 88.44, 87.9, 529.93, 235.42, 545.18, 131.28, 80.19, 155.57, 694.35, 260.18, 118.13, 283.91, 231.3, 440.71, 727.68, 2875.88, 614.01, 107.46, 2599.76, 229.98, 1102.74, 204.27, 537.83, 301.5, 804.59, 814.82, 1173.77, 96.02, 346.09, 524.65, 872.16, 111.08, 387.29, 93.2, 151.91, 2810.83, 78.57, 58.29, 269, 110.38, 249.74, 337.56, 975.24, 287.07, 76.29, 355.72, 1341.48, 337.24, 464.41, 567.14, 873.8, 41.05, 554.97, 260.58, 436.03, 704.46, 315.93, 734.95, 285.94, 117.73, 526.88, 402.26, 568.9, 273.22, 103.74, 80.95, 311.2, 619.74, 121.83, 218.41, 114.34, 147.33, 779.73, 441.31, 274.43, 262.08, 301.44, 728.77, 244.34, 62.13, 209.33, 2345.93, 186.54, 536.91, 454.91, 189.31, 165, 935.9, 545.3, 382.22, 451.03, 269.39, 84.23, 61.11, 153.75, 984.42, 155.81, 663.36, 240.27, 70.11, 179.21, 814.31, 296.92, 429.35, 283.56, 180.41, 326.43, 235.01, 494.68, 145.04, 258.64, 378.42, 256.97, 418.26, 442.29, 315.92, 2736.29, 2096.64, 1645.49, 239.38, 775.77, 605, 4708.78, 2944.92, 633.11, 555.48, 1076.84, 370.76, 288.77, 201.34, 485.2, 594.83, 75.07, 107.2, 1279.97, 376.68, 419.71, 960.82, 321.64, 161.51, 640.71, 215.81, 102.17, 2206.73, 68.89, 599.39, 277.83, 777.17, 168.09, 251.92, 396.87, 670.71, 201.72, 106.74, 211.31, 1152.77, 528.05, 2173.22, 695.45, 985.44, 268.12, 298.23, 4952.07, 838.87, 786.96, 2219.31, 1245.53, 270.68, 284.68, 1115.43, 150.25, 149.35, 448.79, 262.38, 917.98, 142.47, 177.15, 230.08, 237.82, 335.2, 606.99, 113.45, 238.83, 346.07, 507.58, 191.78, 126.32, 324.75, 1637.22, 66.55, 395.15, 906.48, 110.79, 1210.63, 922.66, 128.79, 109.98, 263.7, 280.99, 195.05, 358.87, 205.66, 441.65, 282.82, 531.24, 74.8, 309.47, 471.47, 524.56, 106.52, 968.7, 67.83, 23.03, 476.29, 482.56, 540.95, 468.17, 223.16, 1959.74, 216.28, 113.54, 132.18, 575.15, 389.96, 208.66, 86.84, 1067.37, 1029.14, 144.44, 278.03, 314.29, 216.53, 143.53, 173.64, 475.26, 369.41, 287.35, 190.32, 19.74, 200.12, 147.16, 433.94, 534.82, 270.59, 59.85, 17.61, 133.54, 139.19, 502.64, 122.66, 218.81, 181.87, 367.08, 599.77, 1062.09, 384.94, 210.38, 937.25, 1606.00, 721.63, 135.1, 375.11, 361.81, 279.76, 81.11, 58.48, 288.26, 188.89, 403.09, 316.66, 472.88, 399.3, 148.17, 757.01, 109.63, 283.39, 459.95, 606.22, 196.31, 363.26, 671.27, 201.14, 292.13, 313.87, 494.12, 182.17, 718.98, 576.5, 303.48, 1710.89, 33.48, 72.49, 216.9, 336.1, 1556.99, 122.08, 367.15, 707.73, 529.03, 305.35, 259.76, 270.37, 274.22, 288.89, 93.54, 428.58, 518.34, 488.74, 497.68, 321.02, 285.27, 468.35, 357.72, 102.62, 361.02, 269.25, 905.84, 265.09, 168.63, 73.69, 119.15, 93.16, 488.79, 161.73, 405.26, 123.38, 505.9, 600.96, 988.33, 356.3, 360.44, 1752.70, 360.11, 254.07, 269.19, 2275.65, 1084.78, 392.15, 699.92, 209.74, 601.56, 304.19, 69.01, 857.13, 81.87, 100.4, 518.24, 273.44, 728.14, 183.08, 3,887.45, 390.44, 381.06, 60.74, 575.52, 97.94, 610.33)

q25 <- quantile(data_1, 0.25)
q25
```

#PRECONDITIONING CLUSTER 2 (PROFILE 11) STEM GRAPH
```{r}
library(reshape2)

# STEM Data was imported

# Reformatted data into long format
Profile11 <- melt(DQO_STEM_Results_PC6_Trial_Profile_11, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Profile11$Stage <- factor(Profile11$Stage, levels = unique(Profile11$Stage))
### Basic Plot for Profile 7 (Shows the dip at Pre-Met)
ggplot(Profile11, aes(x= Stage, y=n, group=`Protein Names`, color=`Protein Names`)) + ylim(-100,8000) + #should be min and max of data
  geom_line() +
  labs(title = "Profile 11",
       x = "Time Point",
       y = "Fold Change",
       color = "Gene") +
  theme_minimal()
View(Profile11)

# Highlight AVERAGE in Profile11
highlight_gene <- c("AVERAGE")

#Vector for Group
group_vector_Profile11 <- ifelse(Profile11$`Protein Names` %in% highlight_gene, "Average", "Other")
View(group_vector_Profile11)
df_Profile11 <- as.data.frame(Profile11)
df_Profile11$Color <- c(group_vector_Profile11)
colnames(df_Profile11) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Profile11)
# Ensure that the group levels are factors for colors and opacity
#Vector for Opacity
opacity_values <- c("Average" = 1.0, "Other" = 0.5)
df_Profile11$Group <- factor(df_Profile11$Group, levels = names(opacity_values))
#Vector for Colors
color_values <- c("Average" = "blue", "Other" = "grey")
df_Profile11$Group <- factor(df_Profile11$Group, levels = names(color_values))
# Make x-axis values numeric to change spacing
df_Profile11$Stage_num <- recode(df_Profile11$Stage, "A" = 0, "C" = 14, "D" = 19)
# Plot the data with the correct mapping
Profile11_Graph <- ggplot(df_Profile11, aes(x= Stage_num, y= FC, group = `Protein Names`, alpha = Group, color = Group)) + ylim(-7000,10000) +
  geom_line(size = 0.25) +
  geom_line(data = subset(df_Profile11, df_Profile11$Group == "Average"), aes(color = Group), 
            size = 0.75) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .5) +
  geom_vline(xintercept = 0, linetype = 5, color = "black", size = .5) +
  scale_color_manual(values = c("Others" = "grey", "Average" = "blue")) +
  scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D")) +  # Uneven spacing
  labs(title = "Profile 11",
       x = "Time Point",
       y = "Fold Change",
       color = "Gene Symbol") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),  # Center title
        axis.text.x = element_text(size = 14),  # Adjusts the text size of the x-axis tick labels
        axis.text.y = element_text(size = 14),  # Adjusts the text size of the y-axis tick labels
        axis.title.x = element_text(size = 14), # Adjusts the text size of the x-axis title
        axis.title.y = element_text(size = 14), # Adjusts the text size of the y-axis title
        legend.text = element_text(size = 10),
        legend.title = element_blank(),
        panel.grid = element_blank()) # Removes gridlines
Profile11_Graph + guides(alpha = guide_none())
```


```{r}
t.test(DQO_Data_With_and_Without_Lono$`YES LONO CC/CA DIFFERENTIAL`, DQO_Data_With_and_Without_Lono$`NO LONO CC/CA DIFFERENTIAL`, paired = TRUE)

t.test(DQO_Data_With_and_Without_Lono$`YES LONO PC6C/PC6A DIFFERENTIAL`, DQO_Data_With_and_Without_Lono$`NO LONO PC6C/PC6A DIFFERENTIAL`, paired = TRUE)
```

##NO LONO STEM GRAPHS
#Preconditioning Cluster 1 (Profiles 5 and 6)
```{r}
install.packages("khroma")

library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)
library(khroma)

# STEM Data was imported

# Reformatted data into long format
Cluster1P <- melt(NO_LONO_Preconditioning_STEM_Cluster_1, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Cluster1P$Stage <- factor(Cluster1P$Stage, levels = unique(Cluster1P$Stage))

# Highlight AVERAGE in Cluster1PC6
highlight_gene <- c("AVERAGE_PROFILE5", "AVERAGE_PROFILE6")

#Vector for Group
group_vector_Cluster1P <- ifelse(Cluster1P$`Protein Names` %in% highlight_gene, "Average", "Other")
View(group_vector_Cluster1P)
df_Cluster1P <- as.data.frame(Cluster1P)
df_Cluster1P$Color <- c(group_vector_Cluster1P)
colnames(df_Cluster1P) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Cluster1P)

df_Cluster1P$Stage_num <- recode(df_Cluster1P$Stage, "A" = 0, "C" = 14, "D" = 19)

# Create ranges for 25% ribbon
df_Cluster1P$ymin_25_Cluster1P <- ifelse(df_Cluster1P$Stage == "A", 0, 
                  ifelse(df_Cluster1P$Stage == "C", -383.45, 
                         ifelse(df_Cluster1P$Stage == "D", 35.78, NA)))

df_Cluster1P$ymax_25_Cluster1P <- ifelse(df_Cluster1P$Stage == "A", 0, 
                  ifelse(df_Cluster1P$Stage == "C", -219.51, 
                         ifelse(df_Cluster1P$Stage == "D", 134.45, NA)))
# Create ranges for 50% ribbon
df_Cluster1P$ymin_50_Cluster1P <- ifelse(df_Cluster1P$Stage == "A", 0, 
                  ifelse(df_Cluster1P$Stage == "C", -508.97, 
                         ifelse(df_Cluster1P$Stage == "D", 11.6, NA)))

df_Cluster1P$ymax_50_Cluster1P <- ifelse(df_Cluster1P$Stage == "A", 0, 
                  ifelse(df_Cluster1P$Stage == "C", -143.46, 
                         ifelse(df_Cluster1P$Stage == "D", 211, NA)))
# Create ranges for 75% ribbon
df_Cluster1P$ymin_75_Cluster1P <- ifelse(df_Cluster1P$Stage == "A", 0, 
                  ifelse(df_Cluster1P$Stage == "C", -802.73, 
                         ifelse(df_Cluster1P$Stage == "D", -29.65, NA)))

df_Cluster1P$ymax_75_Cluster1P <- ifelse(df_Cluster1P$Stage == "A", 0, 
                  ifelse(df_Cluster1P$Stage == "C", -94.25, 
                         ifelse(df_Cluster1P$Stage == "D", 379.91, NA)))
# Create ranges for 100% ribbon
df_Cluster1P$ymin_100_Cluster1P <- ifelse(df_Cluster1P$Stage == "A", 0, 
                  ifelse(df_Cluster1P$Stage == "C", -14397.39, 
                         ifelse(df_Cluster1P$Stage == "D", -2146, NA)))

df_Cluster1P$ymax_100_Cluster1P <- ifelse(df_Cluster1P$Stage == "A", 0, 
                  ifelse(df_Cluster1P$Stage == "C", -7.36, 
                         ifelse(df_Cluster1P$Stage == "D", 3345.8, NA)))

# Log transform data
df_Cluster1P_log <- df_Cluster1P %>%
  mutate(FC_log = sign(FC) * log(abs(FC) + 1))

# Plot the data with the correct mapping
Cluster1P_Graph <- ggplot(df_Cluster1P_log, aes(x= Stage_num, y= FC_log, group = `Protein Names`, alpha = Group, color = Group)) + ylim(-10,10) +
  geom_ribbon(aes(ymin = (sign(ymin_100_Cluster1P) * log(abs(ymin_100_Cluster1P) + 1)), ymax = (sign(ymax_100_Cluster1P) * log(abs(ymax_100_Cluster1P) + 1))), fill = "#E5E5E5", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_75_Cluster1P) * log(abs(ymin_75_Cluster1P) + 1)), ymax = (sign(ymax_75_Cluster1P) * log(abs(ymax_75_Cluster1P) + 1))), fill = "#D3D3D3", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_50_Cluster1P) * log(abs(ymin_50_Cluster1P) + 1)), ymax = (sign(ymax_50_Cluster1P) * log(abs(ymax_50_Cluster1P) + 1))), fill = "#A0A0A0", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_25_Cluster1P) * log(abs(ymin_25_Cluster1P) + 1)), ymax = (sign(ymax_25_Cluster1P) * log(abs(ymax_25_Cluster1P) + 1))), fill = "#808080", size = 0) + # Shaded region
  #geom_line(size = 0.05, color = "black") +
  geom_line(data = subset(df_Cluster1P_log, df_Cluster1P_log$Group == "Average"), 
            size = 1, alpha = 1) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .5) +
  geom_vline(xintercept = 0, linetype = 1, color = "black", size = .5) +
  geom_hline(yintercept = -10, linetype = 1, color = "black", size = .5) +
  scale_color_manual(values = c("Average" = "#004488")) +
  #scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D")) +  # Uneven spacing
  labs(x = "Time Point",
       y = "Log(Fold Change)",
       color = "Gene Symbol") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),  # Center title
        axis.text.x = element_text(size = 14),  # Adjusts the text size of the x-axis tick labels
        axis.text.y = element_text(size = 14),  # Adjusts the text size of the y-axis tick labels
        axis.title.x = element_text(size = 14), # Adjusts the text size of the x-axis title
        axis.title.y = element_text(size = 14), # Adjusts the text size of the y-axis title
        legend.text = element_text(size = 10),
        legend.title = element_blank(),
        panel.grid = element_blank()) # Removes gridlines
Cluster1P_Graph + guides(alpha = guide_none())
```
#Control Cluster 1 (Profiles 7 and 2)
```{r}
install.packages("khroma")

library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)
library(khroma)

# STEM Data was imported

# Reformatted data into long format
Cluster1C <- melt(NO_LONO_Control_STEM_Cluster_1, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Cluster1C$Stage <- factor(Cluster1C$Stage, levels = unique(Cluster1C$Stage))

# Highlight AVERAGE in Cluster1PC6
highlight_gene <- c("AVERAGE_PROFILE7", "AVERAGE_PROFILE2")

#Vector for Group
group_vector_Cluster1C <- ifelse(Cluster1C$`Protein Names` %in% highlight_gene, "Average", "Other")
View(group_vector_Cluster1C)
df_Cluster1C <- as.data.frame(Cluster1C)
df_Cluster1C$Color <- c(group_vector_Cluster1C)
colnames(df_Cluster1C) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Cluster1C)

df_Cluster1C$Stage_num <- recode(df_Cluster1C$Stage, "A" = 0, "C" = 14, "D" = 19)

# Create ranges for 25% ribbon
df_Cluster1C$ymin_25_Cluster1C <- ifelse(df_Cluster1C$Stage == "A", 0, 
                  ifelse(df_Cluster1C$Stage == "C", -47.8975, 
                         ifelse(df_Cluster1C$Stage == "D", -356.645, NA)))

df_Cluster1C$ymax_25_Cluster1C <- ifelse(df_Cluster1C$Stage == "A", 0, 
                  ifelse(df_Cluster1C$Stage == "C", -7.73625, 
                         ifelse(df_Cluster1C$Stage == "D", -209.21625, NA)))
# Create ranges for 50% ribbon
df_Cluster1C$ymin_50_Cluster1C <- ifelse(df_Cluster1C$Stage == "A", 0, 
                  ifelse(df_Cluster1C$Stage == "C", -76.1125, 
                         ifelse(df_Cluster1C$Stage == "D", -452.8225, NA)))

df_Cluster1C$ymax_50_Cluster1C <- ifelse(df_Cluster1C$Stage == "A", 0, 
                  ifelse(df_Cluster1C$Stage == "C", 17.405, 
                         ifelse(df_Cluster1C$Stage == "D", -159.2125, NA)))
# Create ranges for 75% ribbon
df_Cluster1C$ymin_75_Cluster1C <- ifelse(df_Cluster1C$Stage == "A", 0, 
                  ifelse(df_Cluster1C$Stage == "C", -149.78375, 
                         ifelse(df_Cluster1C$Stage == "D", -743.25625, NA)))

df_Cluster1C$ymax_75_Cluster1C <- ifelse(df_Cluster1C$Stage == "A", 0, 
                  ifelse(df_Cluster1C$Stage == "C", -94.25, 
                         ifelse(df_Cluster1C$Stage == "D", -98.22375, NA)))
# Create ranges for 100% ribbon
df_Cluster1C$ymin_100_Cluster1C <- ifelse(df_Cluster1C$Stage == "A", 0, 
                  ifelse(df_Cluster1C$Stage == "C", -1206.59, 
                         ifelse(df_Cluster1C$Stage == "D", -6149.56, NA)))

df_Cluster1C$ymax_100_Cluster1C <- ifelse(df_Cluster1C$Stage == "A", 0, 
                  ifelse(df_Cluster1C$Stage == "C", 794.88, 
                         ifelse(df_Cluster1C$Stage == "D", -20.38, NA)))

# Log transform data
df_Cluster1C_log <- df_Cluster1C %>%
  mutate(FC_log = sign(FC) * log(abs(FC) + 1))

# Plot the data with the correct mapping
Cluster1C_Graph <- ggplot(df_Cluster1C_log, aes(x= Stage_num, y= FC_log, group = `Protein Names`, alpha = Group, color = Group)) + ylim(-10,10) +
  geom_ribbon(aes(ymin = (sign(ymin_100_Cluster1C) * log(abs(ymin_100_Cluster1C) + 1)), ymax = (sign(ymax_100_Cluster1C) * log(abs(ymax_100_Cluster1C) + 1))), fill = "#E5E5E5", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_75_Cluster1C) * log(abs(ymin_75_Cluster1C) + 1)), ymax = (sign(ymax_75_Cluster1C) * log(abs(ymax_75_Cluster1C) + 1))), fill = "#D3D3D3", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_50_Cluster1C) * log(abs(ymin_50_Cluster1C) + 1)), ymax = (sign(ymax_50_Cluster1C) * log(abs(ymax_50_Cluster1C) + 1))), fill = "#A0A0A0", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_25_Cluster1C) * log(abs(ymin_25_Cluster1C) + 1)), ymax = (sign(ymax_25_Cluster1C) * log(abs(ymax_25_Cluster1C) + 1))), fill = "#808080", size = 0) + # Shaded region
  #geom_line(size = 0.05, color = "black") +
  geom_line(data = subset(df_Cluster1C_log, df_Cluster1C_log$Group == "Average"), 
            size = 1, alpha = 1) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .5) +
  geom_vline(xintercept = 0, linetype = 1, color = "black", size = .5) +
  geom_hline(yintercept = -10, linetype = 1, color = "black", size = .5) +
  scale_color_manual(values = c("Average" = "#6699CC")) +
  #scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D")) +  # Uneven spacing
  labs(x = "Time Point",
       y = "Log(Fold Change)",
       color = "Gene Symbol") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),  # Center title
        axis.text.x = element_text(size = 14),  # Adjusts the text size of the x-axis tick labels
        axis.text.y = element_text(size = 14),  # Adjusts the text size of the y-axis tick labels
        axis.title.x = element_text(size = 14), # Adjusts the text size of the x-axis title
        axis.title.y = element_text(size = 14), # Adjusts the text size of the y-axis title
        legend.text = element_text(size = 10),
        legend.title = element_blank(),
        panel.grid = element_blank()) # Removes gridlines
Cluster1C_Graph + guides(alpha = guide_none())
```
#Control Cluster 2 (Profile 13)
```{r}
install.packages("khroma")

library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)
library(khroma)

# STEM Data was imported

# Reformatted data into long format
Cluster2C <- melt(NO_LONO_Control_STEM_Cluster_2, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Cluster2C$Stage <- factor(Cluster2C$Stage, levels = unique(Cluster2C$Stage))

# Highlight AVERAGE in Cluster1PC6
highlight_gene <- c("AVERAGE_PROFILE13")

#Vector for Group
group_vector_Cluster2C <- ifelse(Cluster2C$`Protein Names` %in% highlight_gene, "Average", "Other")
View(group_vector_Cluster2C)
df_Cluster2C <- as.data.frame(Cluster2C)
df_Cluster2C$Color <- c(group_vector_Cluster2C)
colnames(df_Cluster2C) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Cluster2C)

df_Cluster2C$Stage_num <- recode(df_Cluster2C$Stage, "A" = 0, "C" = 14, "D" = 19)

# Create ranges for 25% ribbon
df_Cluster2C$ymin_25_Cluster2C <- ifelse(df_Cluster2C$Stage == "A", 0, 
                  ifelse(df_Cluster2C$Stage == "C", 78.64625, 
                         ifelse(df_Cluster2C$Stage == "D", 283.4975, NA)))

df_Cluster2C$ymax_25_Cluster2C <- ifelse(df_Cluster2C$Stage == "A", 0, 
                  ifelse(df_Cluster2C$Stage == "C", 126.74625, 
                         ifelse(df_Cluster2C$Stage == "D", 453.855, NA)))
# Create ranges for 50% ribbon
df_Cluster2C$ymin_50_Cluster2C <- ifelse(df_Cluster2C$Stage == "A", 0, 
                  ifelse(df_Cluster2C$Stage == "C", 56.0875, 
                         ifelse(df_Cluster2C$Stage == "D", 207.305, NA)))

df_Cluster2C$ymax_50_Cluster2C <- ifelse(df_Cluster2C$Stage == "A", 0, 
                  ifelse(df_Cluster2C$Stage == "C", 172.765, 
                         ifelse(df_Cluster2C$Stage == "D", 607.7, NA)))
# Create ranges for 75% ribbon
df_Cluster2C$ymin_75_Cluster2C <- ifelse(df_Cluster2C$Stage == "A", 0, 
                  ifelse(df_Cluster2C$Stage == "C", 41.3175, 
                         ifelse(df_Cluster2C$Stage == "D", 138.92375, NA)))

df_Cluster2C$ymax_75_Cluster2C <- ifelse(df_Cluster2C$Stage == "A", 0, 
                  ifelse(df_Cluster2C$Stage == "C", 282.145, 
                         ifelse(df_Cluster2C$Stage == "D", 869.345, NA)))
# Create ranges for 100% ribbon
df_Cluster2C$ymin_100_Cluster2C <- ifelse(df_Cluster2C$Stage == "A", 0, 
                  ifelse(df_Cluster2C$Stage == "C", 8.25, 
                         ifelse(df_Cluster2C$Stage == "D", 21.53, NA)))

df_Cluster2C$ymax_100_Cluster2C <- ifelse(df_Cluster2C$Stage == "A", 0, 
                  ifelse(df_Cluster2C$Stage == "C", 1574.43, 
                         ifelse(df_Cluster2C$Stage == "D", 4906.2, NA)))

# Log transform data
df_Cluster2C_log <- df_Cluster2C %>%
  mutate(FC_log = sign(FC) * log(abs(FC) + 1))

# Plot the data with the correct mapping
Cluster2C_Graph <- ggplot(df_Cluster2C_log, aes(x= Stage_num, y= FC_log, group = `Protein Names`, alpha = Group, color = Group)) + ylim(-10,10) +
  geom_ribbon(aes(ymin = (sign(ymin_100_Cluster2C) * log(abs(ymin_100_Cluster2C) + 1)), ymax = (sign(ymax_100_Cluster2C) * log(abs(ymax_100_Cluster2C) + 1))), fill = "#E5E5E5", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_75_Cluster2C) * log(abs(ymin_75_Cluster2C) + 1)), ymax = (sign(ymax_75_Cluster2C) * log(abs(ymax_75_Cluster2C) + 1))), fill = "#D3D3D3", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_50_Cluster2C) * log(abs(ymin_50_Cluster2C) + 1)), ymax = (sign(ymax_50_Cluster2C) * log(abs(ymax_50_Cluster2C) + 1))), fill = "#A0A0A0", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_25_Cluster2C) * log(abs(ymin_25_Cluster2C) + 1)), ymax = (sign(ymax_25_Cluster2C) * log(abs(ymax_25_Cluster2C) + 1))), fill = "#808080", size = 0) + # Shaded region
  #geom_line(size = 0.05, color = "black") +
  geom_line(data = subset(df_Cluster2C_log, df_Cluster2C_log$Group == "Average"), 
            size = 1, alpha = 1) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .5) +
  geom_vline(xintercept = 0, linetype = 1, color = "black", size = .5) +
  geom_hline(yintercept = -10, linetype = 1, color = "black", size = .5) +
  scale_color_manual(values = c("Average" = "#6699CC")) +
  #scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D")) +  # Uneven spacing
  labs(x = "Time Point",
       y = "Log(Fold Change)",
       color = "Gene Symbol") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),  # Center title
        axis.text.x = element_text(size = 14),  # Adjusts the text size of the x-axis tick labels
        axis.text.y = element_text(size = 14),  # Adjusts the text size of the y-axis tick labels
        axis.title.x = element_text(size = 14), # Adjusts the text size of the x-axis title
        axis.title.y = element_text(size = 14), # Adjusts the text size of the y-axis title
        legend.text = element_text(size = 10),
        legend.title = element_blank(),
        panel.grid = element_blank()) # Removes gridlines
Cluster2C_Graph + guides(alpha = guide_none())
```

##STEM GRAPHS FOR MY THESIS (WITH LONO)
#Control Cluster 1 (Profiles 7, 2, and 3) [WITH LONO]
```{r}
#install.packages("khroma")

library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)
library(khroma)

# STEM Data was imported

# Reformatted data into long format
Cluster1C_YESLONO <- melt(YES_LONO_Control_STEM_Cluster_1, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Cluster1C_YESLONO$Stage <- factor(Cluster1C_YESLONO$Stage, levels = unique(Cluster1C_YESLONO$Stage))

# Highlight AVERAGE in Cluster1PC6
highlight_gene <- c("AVERAGE")

#Vector for Group
group_vector_Cluster1C_YESLONO <- ifelse(Cluster1C_YESLONO$`Protein Names` %in% highlight_gene, "AVERAGE", "Other")
View(group_vector_Cluster1C_YESLONO)
df_Cluster1C_YESLONO <- as.data.frame(Cluster1C_YESLONO)
df_Cluster1C_YESLONO$Color <- c(group_vector_Cluster1C_YESLONO)
colnames(df_Cluster1C_YESLONO) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Cluster1C_YESLONO)

df_Cluster1C_YESLONO$Stage_num <- recode(df_Cluster1C_YESLONO$Stage, "A" = 0, "C" = 14, "D" = 19)

# Create ranges for 25% ribbon
df_Cluster1C_YESLONO$ymin_25_Cluster1C_YESLONO <- ifelse(df_Cluster1C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1C_YESLONO$Stage == "C", -66.73, 
                         ifelse(df_Cluster1C_YESLONO$Stage == "D", -333.01, NA)))

df_Cluster1C_YESLONO$ymax_25_Cluster1C_YESLONO <- ifelse(df_Cluster1C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1C_YESLONO$Stage == "C", -19.72, 
                         ifelse(df_Cluster1C_YESLONO$Stage == "D", -173.93, NA)))
# Create ranges for 50% ribbon
df_Cluster1C_YESLONO$ymin_50_Cluster1C_YESLONO <- ifelse(df_Cluster1C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1C_YESLONO$Stage == "C", -103.21, 
                         ifelse(df_Cluster1C_YESLONO$Stage == "D", -448.88, NA)))

df_Cluster1C_YESLONO$ymax_50_Cluster1C_YESLONO <- ifelse(df_Cluster1C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1C_YESLONO$Stage == "C", -1.89,
                         ifelse(df_Cluster1C_YESLONO$Stage == "D", -131.8, NA)))
# Create ranges for 75% ribbon
df_Cluster1C_YESLONO$ymin_75_Cluster1C_YESLONO <- ifelse(df_Cluster1C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1C_YESLONO$Stage == "C", -196.83, 
                         ifelse(df_Cluster1C_YESLONO$Stage == "D", -723.78, NA)))

df_Cluster1C_YESLONO$ymax_75_Cluster1C_YESLONO <- ifelse(df_Cluster1C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1C_YESLONO$Stage == "C", 28.75, 
                         ifelse(df_Cluster1C_YESLONO$Stage == "D", -97.38, NA)))
# Create ranges for 100% ribbon
df_Cluster1C_YESLONO$ymin_100_Cluster1C_YESLONO <- ifelse(df_Cluster1C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1C_YESLONO$Stage == "C", -2261.59, 
                         ifelse(df_Cluster1C_YESLONO$Stage == "D", -6815.98, NA)))

df_Cluster1C_YESLONO$ymax_100_Cluster1C_YESLONO <- ifelse(df_Cluster1C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1C_YESLONO$Stage == "C", 1462.78, 
                         ifelse(df_Cluster1C_YESLONO$Stage == "D", -4.28, NA)))

# Log transform data
df_Cluster1C_log_YESLONO <- df_Cluster1C_YESLONO %>%
  mutate(FC_log = sign(FC) * log(abs(FC) + 1))

# Plot the data with the correct mapping
Cluster1C_Graph_YESLONO <- ggplot(df_Cluster1C_log_YESLONO, aes(x= Stage_num, y= FC_log, group = `Protein Names`, alpha = Group, color = Group, fill = Group)) +
  geom_ribbon(aes(ymin = (sign(ymin_100_Cluster1C_YESLONO) * log(abs(ymin_100_Cluster1C_YESLONO) + 1)), ymax = (sign(ymax_100_Cluster1C_YESLONO) * log(abs(ymax_100_Cluster1C_YESLONO) + 1))), fill = "#E5E5E5", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_75_Cluster1C_YESLONO) * log(abs(ymin_75_Cluster1C_YESLONO) + 1)), ymax = (sign(ymax_75_Cluster1C_YESLONO) * log(abs(ymax_75_Cluster1C_YESLONO) + 1))), fill = "#D3D3D3", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_50_Cluster1C_YESLONO) * log(abs(ymin_50_Cluster1C_YESLONO) + 1)), ymax = (sign(ymax_50_Cluster1C_YESLONO) * log(abs(ymax_50_Cluster1C_YESLONO) + 1))), fill = "#A0A0A0", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_25_Cluster1C_YESLONO) * log(abs(ymin_25_Cluster1C_YESLONO) + 1)), ymax = (sign(ymax_25_Cluster1C_YESLONO) * log(abs(ymax_25_Cluster1C_YESLONO) + 1))), fill = "#808080", size = 0) + # Shaded region
  #geom_line(size = 0.05, color = "black") +
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 25%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 50%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 75%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 100%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  geom_line(data = subset(df_Cluster1C_log_YESLONO, df_Cluster1C_log_YESLONO$Group == "AVERAGE"), 
            size = 1, alpha = 1) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .25) +
  geom_vline(xintercept = 0, linetype = 1, color = "black", size = .4) +
  geom_hline(yintercept = -10, linetype = 1, color = "black", size = .4) +
  scale_color_manual(values = c("AVERAGE" = "#4FAFFF"),
                     labels = c("AVERAGE" = "Average")) +
  scale_fill_manual( values = c("Middle 25%" = "#808080",
                                "Middle 50%" = "#A0A0A0",
                                "Middle 75%" = "#D3D3D3",
                                "Middle 100%" = "#E5E5E5"),
                     breaks = c("Middle 25%",
                                "Middle 50%",
                                "Middle 75%",
                                "Middle 100%")) +
  #scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D"), 
                     expand = c(0, 0)) +  # X-axis tick marks at positions with custom labels
  scale_y_continuous(limits = c(-10, 10),
                     breaks = seq(-10, 10, by = 5),  # Adding y-axis tick marks at 5 unit intervals
                     expand = c(0, 0)) +  # Ensures no extra space is added around the y-axis
  guides(
    color = guide_legend(
    title = "Group",
    override.aes = list(fill = NA)  # Remove the gray fill box behind the line
  ),
  fill = guide_legend(title = "Confidence Interval")
  ) +
  labs(title = "Cluster 1",
       x = "Time Point",
       y = "Phosphorylation Log Fold Change") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, hjust = 0.5),  # Center title
        panel.grid = element_blank(),       # remove all gridlines
        legend.title = element_blank(),
        legend.position = "bottom",
        axis.ticks = element_line(color = "black"),
        legend.key = element_rect(fill = "transparent", color = NA)
        )
Cluster1C_Graph_YESLONO + guides(alpha = guide_none())
```
#Control Cluster 2 (Profiles 8 and 3) [WITH LONO]
```{r}
#install.packages("khroma")

library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)
library(khroma)

# STEM Data was imported

# Reformatted data into long format
Cluster2C_YESLONO <- melt(YES_LONO_Control_STEM_Cluster_2, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Cluster2C_YESLONO$Stage <- factor(Cluster2C_YESLONO$Stage, levels = unique(Cluster2C_YESLONO$Stage))

# Highlight AVERAGE in Cluster1PC6
highlight_gene <- c("AVERAGE")

#Vector for Group
group_vector_Cluster2C_YESLONO <- ifelse(Cluster2C_YESLONO$`Protein Names` %in% highlight_gene, "AVERAGE", "Other")
View(group_vector_Cluster2C_YESLONO)
df_Cluster2C_YESLONO <- as.data.frame(Cluster2C_YESLONO)
df_Cluster2C_YESLONO$Color <- c(group_vector_Cluster2C_YESLONO)
colnames(df_Cluster2C_YESLONO) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Cluster2C_YESLONO)

df_Cluster2C_YESLONO$Stage_num <- recode(df_Cluster2C_YESLONO$Stage, "A" = 0, "C" = 14, "D" = 19)

# Create ranges for 25% ribbon
df_Cluster2C_YESLONO$ymin_25_Cluster2C_YESLONO <- ifelse(df_Cluster2C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2C_YESLONO$Stage == "C", 3.79, 
                         ifelse(df_Cluster2C_YESLONO$Stage == "D", 207.865, NA)))

df_Cluster2C_YESLONO$ymax_25_Cluster2C_YESLONO <- ifelse(df_Cluster2C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2C_YESLONO$Stage == "C", 34.2675, 
                         ifelse(df_Cluster2C_YESLONO$Stage == "D", 319.9175, NA)))
# Create ranges for 50% ribbon
df_Cluster2C_YESLONO$ymin_50_Cluster2C_YESLONO <- ifelse(df_Cluster2C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2C_YESLONO$Stage == "C", -13.9,
                         ifelse(df_Cluster2C_YESLONO$Stage == "D", 147.92, NA)))

df_Cluster2C_YESLONO$ymax_50_Cluster2C_YESLONO <- ifelse(df_Cluster2C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2C_YESLONO$Stage == "C", 62.485,
                         ifelse(df_Cluster2C_YESLONO$Stage == "D", 428.495, NA)))
# Create ranges for 75% ribbon
df_Cluster2C_YESLONO$ymin_75_Cluster2C_YESLONO <- ifelse(df_Cluster2C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2C_YESLONO$Stage == "C", -40.76, 
                         ifelse(df_Cluster2C_YESLONO$Stage == "D", 90.64, NA)))

df_Cluster2C_YESLONO$ymax_75_Cluster2C_YESLONO <- ifelse(df_Cluster2C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2C_YESLONO$Stage == "C", 111.24, 
                         ifelse(df_Cluster2C_YESLONO$Stage == "D", 707.595, NA)))
# Create ranges for 100% ribbon
df_Cluster2C_YESLONO$ymin_100_Cluster2C_YESLONO <- ifelse(df_Cluster2C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2C_YESLONO$Stage == "C", -514.03, 
                         ifelse(df_Cluster2C_YESLONO$Stage == "D", 15.85, NA)))

df_Cluster2C_YESLONO$ymax_100_Cluster2C_YESLONO <- ifelse(df_Cluster2C_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2C_YESLONO$Stage == "C", 2467.91, 
                         ifelse(df_Cluster2C_YESLONO$Stage == "D", 7504.34, NA)))

# Log transform data
df_Cluster2C_log_YESLONO <- df_Cluster2C_YESLONO %>%
  mutate(FC_log = sign(FC) * log(abs(FC) + 1))

# Plot the data with the correct mapping
Cluster2C_Graph_YESLONO <- ggplot(df_Cluster2C_log_YESLONO, aes(x= Stage_num, y= FC_log, group = `Protein Names`, alpha = Group, color = Group, fill = Group)) +
  geom_ribbon(aes(ymin = (sign(ymin_100_Cluster2C_YESLONO) * log(abs(ymin_100_Cluster2C_YESLONO) + 1)), ymax = (sign(ymax_100_Cluster2C_YESLONO) * log(abs(ymax_100_Cluster2C_YESLONO) + 1))), fill = "#E5E5E5", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_75_Cluster2C_YESLONO) * log(abs(ymin_75_Cluster2C_YESLONO) + 1)), ymax = (sign(ymax_75_Cluster2C_YESLONO) * log(abs(ymax_75_Cluster2C_YESLONO) + 1))), fill = "#D3D3D3", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_50_Cluster2C_YESLONO) * log(abs(ymin_50_Cluster2C_YESLONO) + 1)), ymax = (sign(ymax_50_Cluster2C_YESLONO) * log(abs(ymax_50_Cluster2C_YESLONO) + 1))), fill = "#A0A0A0", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_25_Cluster2C_YESLONO) * log(abs(ymin_25_Cluster2C_YESLONO) + 1)), ymax = (sign(ymax_25_Cluster2C_YESLONO) * log(abs(ymax_25_Cluster2C_YESLONO) + 1))), fill = "#808080", size = 0) + # Shaded region
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 25%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 50%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 75%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 100%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  #geom_line(size = 0.05, color = "black") +
  geom_line(data = subset(df_Cluster2C_log_YESLONO, df_Cluster2C_log_YESLONO$Group == "AVERAGE"), 
            size = 1, alpha = 1) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .25) +
  geom_vline(xintercept = 0, linetype = 1, color = "black", size = .4) +
  geom_hline(yintercept = -10, linetype = 1, color = "black", size = .4) +
  scale_color_manual(values = c("AVERAGE" = "#4FAFFF"),
                     labels = c("AVERAGE" = "Average")) +
  scale_fill_manual( values = c("Middle 25%" = "#808080",
                                "Middle 50%" = "#A0A0A0",
                                "Middle 75%" = "#D3D3D3",
                                "Middle 100%" = "#E5E5E5"),
                     breaks = c("Middle 25%",
                                "Middle 50%",
                                "Middle 75%",
                                "Middle 100%")) +
  #scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D"),
                     expand = c(0,0)) +
  scale_y_continuous(limits = c(-10, 10),
                     breaks = seq(-10, 10, by = 5),
                     expand = c(0,0)) +
  guides(
    color = guide_legend(
      title = "Group",
      override.aes = list(fill = NA)
    ),
    fill = guide_legend(title = "Confidence Interval")
  ) +
  labs(title = "Cluster 2",
       x = "Time Point",
       y = "Phosphorylation Log Fold Change") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, hjust = 0.5),  # Center title
        panel.grid = element_blank(),       # remove all gridlines
        legend.title = element_blank(),
        legend.position = "bottom",
        axis.ticks = element_line(color = "black"),
        legend.key = element_rect(fill = "transparent", color = NA)
        )
Cluster2C_Graph_YESLONO + guides(alpha = guide_none())
```
#Preconditioning Cluster 1 (Profiles 14 and 10) [WITH LONO]
```{r}
#install.packages("khroma")

library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)
library(khroma)

# STEM Data was imported

# Reformatted data into long format
Cluster1P_YESLONO <- melt(YES_LONO_Preconditioning_STEM_Cluster_1, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Cluster1P_YESLONO$Stage <- factor(Cluster1P_YESLONO$Stage, levels = unique(Cluster1P_YESLONO$Stage))

# Highlight AVERAGE in Cluster1PC6
highlight_gene <- c("AVERAGE")

#Vector for Group
group_vector_Cluster1P_YESLONO <- ifelse(Cluster1P_YESLONO$`Protein Names` %in% highlight_gene, "AVERAGE", "Other")
View(group_vector_Cluster1P_YESLONO)
df_Cluster1P_YESLONO <- as.data.frame(Cluster1P_YESLONO)
df_Cluster1P_YESLONO$Color <- c(group_vector_Cluster1P_YESLONO)
colnames(df_Cluster1P_YESLONO) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Cluster1P_YESLONO)

df_Cluster1P_YESLONO$Stage_num <- recode(df_Cluster1P_YESLONO$Stage, "A" = 0, "C" = 14, "D" = 19)

# Create ranges for 25% ribbon
df_Cluster1P_YESLONO$ymin_25_Cluster1P_YESLONO <- ifelse(df_Cluster1P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1P_YESLONO$Stage == "C", 441.79, 
                         ifelse(df_Cluster1P_YESLONO$Stage == "D", 117.14, NA)))

df_Cluster1P_YESLONO$ymax_25_Cluster1P_YESLONO <- ifelse(df_Cluster1P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1P_YESLONO$Stage == "C",736.85, 
                         ifelse(df_Cluster1P_YESLONO$Stage == "D", 287.35, NA)))
# Create ranges for 50% ribbon
df_Cluster1P_YESLONO$ymin_50_Cluster1P_YESLONO <- ifelse(df_Cluster1P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1P_YESLONO$Stage == "C", 310.8, 
                         ifelse(df_Cluster1P_YESLONO$Stage == "D", 58.29, NA)))

df_Cluster1P_YESLONO$ymax_50_Cluster1P_YESLONO <- ifelse(df_Cluster1P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1P_YESLONO$Stage == "C", 981.29,
                         ifelse(df_Cluster1P_YESLONO$Stage == "D", 440.96, NA)))
# Create ranges for 75% ribbon
df_Cluster1P_YESLONO$ymin_75_Cluster1P_YESLONO <- ifelse(df_Cluster1P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1P_YESLONO$Stage == "C", 193.57, 
                         ifelse(df_Cluster1P_YESLONO$Stage == "D", 0.48, NA)))

df_Cluster1P_YESLONO$ymax_75_Cluster1P_YESLONO <- ifelse(df_Cluster1P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1P_YESLONO$Stage == "C", 1504.43, 
                         ifelse(df_Cluster1P_YESLONO$Stage == "D", 699.92, NA)))
# Create ranges for 100% ribbon
df_Cluster1P_YESLONO$ymin_100_Cluster1P_YESLONO <- ifelse(df_Cluster1P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1P_YESLONO$Stage == "C", 11.18, 
                         ifelse(df_Cluster1P_YESLONO$Stage == "D", -1050.14, NA)))

df_Cluster1P_YESLONO$ymax_100_Cluster1P_YESLONO <- ifelse(df_Cluster1P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster1P_YESLONO$Stage == "C", 9770.15, 
                         ifelse(df_Cluster1P_YESLONO$Stage == "D", 4952.07, NA)))

# Log transform data
df_Cluster1P_log_YESLONO <- df_Cluster1P_YESLONO %>%
  mutate(FC_log = sign(FC) * log(abs(FC) + 1))

# Plot the data with the correct mapping
Cluster1P_Graph_YESLONO <- ggplot(df_Cluster1P_log_YESLONO, aes(x= Stage_num, y= FC_log, group = `Protein Names`, alpha = Group, color = Group, fill = Group)) +
  geom_ribbon(aes(ymin = (sign(ymin_100_Cluster1P_YESLONO) * log(abs(ymin_100_Cluster1P_YESLONO) + 1)), ymax = (sign(ymax_100_Cluster1P_YESLONO) * log(abs(ymax_100_Cluster1P_YESLONO) + 1))), fill = "#E5E5E5", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_75_Cluster1P_YESLONO) * log(abs(ymin_75_Cluster1P_YESLONO) + 1)), ymax = (sign(ymax_75_Cluster1P_YESLONO) * log(abs(ymax_75_Cluster1P_YESLONO) + 1))), fill = "#D3D3D3", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_50_Cluster1P_YESLONO) * log(abs(ymin_50_Cluster1P_YESLONO) + 1)), ymax = (sign(ymax_50_Cluster1P_YESLONO) * log(abs(ymax_50_Cluster1P_YESLONO) + 1))), fill = "#A0A0A0", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_25_Cluster1P_YESLONO) * log(abs(ymin_25_Cluster1P_YESLONO) + 1)), ymax = (sign(ymax_25_Cluster1P_YESLONO) * log(abs(ymax_25_Cluster1P_YESLONO) + 1))), fill = "#808080", size = 0) + # Shaded region
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 25%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 50%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 75%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 100%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  #geom_line(size = 0.05, color = "black") +
  geom_line(data = subset(df_Cluster1P_log_YESLONO, df_Cluster1P_log_YESLONO$Group == "AVERAGE"), 
            size = 1, alpha = 1) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .25) +
  geom_vline(xintercept = 0, linetype = 1, color = "black", size = .4) +
  geom_hline(yintercept = -10, linetype = 1, color = "black", size = .4) +
  scale_color_manual(values = c("AVERAGE" = "#0059B2"),
                     labels = c("AVERAGE" = "Average")) +
  scale_fill_manual( values = c("Middle 25%" = "#808080",
                                "Middle 50%" = "#A0A0A0",
                                "Middle 75%" = "#D3D3D3",
                                "Middle 100%" = "#E5E5E5"),
                     breaks = c("Middle 25%",
                                "Middle 50%",
                                "Middle 75%",
                                "Middle 100%")) +
  #scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D"),
                     expand = c(0, 0)) +  # Uneven spacing
  scale_y_continuous(limits = c(-10, 10),
                     breaks = seq(-10, 10, by = 5),
                     expand = c(0, 0)) +
  guides(
    color = guide_legend(
      title = "Group",
      override.aes = list(fill = NA)
    ),
    fill = guide_legend(title = "Confidence Interval")
  ) +
  labs(title = "Cluster 1",
       x = "Time Point",
       y = "Phosphorylation Log Fold Change") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, hjust = 0.5),  # Center title
        panel.grid = element_blank(),       # remove all gridlines
        legend.title = element_blank(),
        legend.position = "bottom",
        axis.ticks = element_line(color = "black"),
        legend.key = element_rect(fill = "transparent", color = NA)
        )
Cluster1P_Graph_YESLONO + guides(alpha = guide_none())
```
#Preconditioning Cluster 2 (Profile 11) [WITH LONO]
```{r}
#install.packages("khroma")

library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)
library(khroma)

# STEM Data was imported

# Reformatted data into long format
Cluster2P_YESLONO <- melt(YES_LONO_Preconditioning_STEM_Cluster_2, id.vars = "Protein Names", variable.name = "Stage", value.name = "n")

# Convert the x-axis variable to a factor with the order as in the data file (Makes A 1st)
Cluster2P_YESLONO$Stage <- factor(Cluster2P_YESLONO$Stage, levels = unique(Cluster2P_YESLONO$Stage))

# Highlight AVERAGE in Cluster1PC6
highlight_gene <- c("AVERAGE")

#Vector for Group
group_vector_Cluster2P_YESLONO <- ifelse(Cluster2P_YESLONO$`Protein Names` %in% highlight_gene, "AVERAGE", "Other")
View(group_vector_Cluster2P_YESLONO)
df_Cluster2P_YESLONO <- as.data.frame(Cluster2P_YESLONO)
df_Cluster2P_YESLONO$Color <- c(group_vector_Cluster2P_YESLONO)
colnames(df_Cluster2P_YESLONO) <- c("Protein Names", "Stage", "FC", "Group")
View(df_Cluster2P_YESLONO)

df_Cluster2P_YESLONO$Stage_num <- recode(df_Cluster2P_YESLONO$Stage, "A" = 0, "C" = 14, "D" = 19)

# Create ranges for 25% ribbon
df_Cluster2P_YESLONO$ymin_25_Cluster2P_YESLONO <- ifelse(df_Cluster2P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2P_YESLONO$Stage == "C", 314.63, 
                         ifelse(df_Cluster2P_YESLONO$Stage == "D", 290.2, NA)))

df_Cluster2P_YESLONO$ymax_25_Cluster2P_YESLONO <- ifelse(df_Cluster2P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2P_YESLONO$Stage == "C",520.9575, 
                         ifelse(df_Cluster2P_YESLONO$Stage == "D", 449.995, NA)))
# Create ranges for 50% ribbon
df_Cluster2P_YESLONO$ymin_50_Cluster2P_YESLONO <- ifelse(df_Cluster2P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2P_YESLONO$Stage == "C", 226.55, 
                         ifelse(df_Cluster2P_YESLONO$Stage == "D", 209.1, NA)))

df_Cluster2P_YESLONO$ymax_50_Cluster2P_YESLONO <- ifelse(df_Cluster2P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2P_YESLONO$Stage == "C", 744.88,
                         ifelse(df_Cluster2P_YESLONO$Stage == "D", 674.625, NA)))
# Create ranges for 75% ribbon
df_Cluster2P_YESLONO$ymin_75_Cluster2P_YESLONO <- ifelse(df_Cluster2P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2P_YESLONO$Stage == "C", 149.4525, 
                         ifelse(df_Cluster2P_YESLONO$Stage == "D", 143.5275, NA)))

df_Cluster2P_YESLONO$ymax_75_Cluster2P_YESLONO <- ifelse(df_Cluster2P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2P_YESLONO$Stage == "C", 1020.6025, 
                         ifelse(df_Cluster2P_YESLONO$Stage == "D", 913.1775, NA)))
# Create ranges for 100% ribbon
df_Cluster2P_YESLONO$ymin_100_Cluster2P_YESLONO <- ifelse(df_Cluster2P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2P_YESLONO$Stage == "C", 40.61, 
                         ifelse(df_Cluster2P_YESLONO$Stage == "D", 35.78, NA)))

df_Cluster2P_YESLONO$ymax_100_Cluster2P_YESLONO <- ifelse(df_Cluster2P_YESLONO$Stage == "A", 0, 
                  ifelse(df_Cluster2P_YESLONO$Stage == "C", 7362.46, 
                         ifelse(df_Cluster2P_YESLONO$Stage == "D", 6697.19, NA)))

# Log transform data
df_Cluster2P_log_YESLONO <- df_Cluster2P_YESLONO %>%
  mutate(FC_log = sign(FC) * log(abs(FC) + 1))

# Plot the data with the correct mapping
Cluster2P_Graph_YESLONO <- ggplot(df_Cluster2P_log_YESLONO, aes(x= Stage_num, y= FC_log, group = `Protein Names`, alpha = Group, color = Group, fill = Group)) +
  geom_ribbon(aes(ymin = (sign(ymin_100_Cluster2P_YESLONO) * log(abs(ymin_100_Cluster2P_YESLONO) + 1)), ymax = (sign(ymax_100_Cluster2P_YESLONO) * log(abs(ymax_100_Cluster2P_YESLONO) + 1))), fill = "#E5E5E5", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_75_Cluster2P_YESLONO) * log(abs(ymin_75_Cluster2P_YESLONO) + 1)), ymax = (sign(ymax_75_Cluster2P_YESLONO) * log(abs(ymax_75_Cluster2P_YESLONO) + 1))), fill = "#D3D3D3", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_50_Cluster2P_YESLONO) * log(abs(ymin_50_Cluster2P_YESLONO) + 1)), ymax = (sign(ymax_50_Cluster2P_YESLONO) * log(abs(ymax_50_Cluster2P_YESLONO) + 1))), fill = "#A0A0A0", size = 0) + # Shaded region
  geom_ribbon(aes(ymin = (sign(ymin_25_Cluster2P_YESLONO) * log(abs(ymin_25_Cluster2P_YESLONO) + 1)), ymax = (sign(ymax_25_Cluster2P_YESLONO) * log(abs(ymax_25_Cluster2P_YESLONO) + 1))), fill = "#808080", size = 0) + # Shaded region
  #geom_line(size = 0.05, color = "black") +
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 25%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 50%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 75%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  geom_ribbon(data = data.frame(x = 1, ymin = 0, ymax = 1, CI_level = "Middle 100%"),
            aes(x = x, ymin = ymin, ymax = ymax, fill = CI_level), inherit.aes = FALSE) +
  geom_line(data = subset(df_Cluster2P_log_YESLONO, df_Cluster2P_log_YESLONO$Group == "AVERAGE"), 
            size = 1, alpha = 1) +
  geom_hline(yintercept = 0, linetype = 5, color = "black", size = .25) +
  geom_vline(xintercept = 0, linetype = 1, color = "black", size = .4) +
  geom_hline(yintercept = -10, linetype = 1, color = "black", size = .4) +
  scale_color_manual(values = c("AVERAGE" = "#0059B2"),
                     labels = c("AVERAGE" = "Average")) +
  scale_fill_manual( values = c("Middle 25%" = "#808080",
                                "Middle 50%" = "#A0A0A0",
                                "Middle 75%" = "#D3D3D3",
                                "Middle 100%" = "#E5E5E5"),
                     breaks = c("Middle 25%",
                                "Middle 50%",
                                "Middle 75%",
                                "Middle 100%")) +
  #scale_alpha_manual(values = opacity_values) +
  scale_x_continuous(breaks = c(0, 14, 19), labels = c("A", "C", "D"),
                     expand = c(0,0)) +
  scale_y_continuous(limits = c(-10, 10),
                     breaks = seq(-10, 10, by = 5),
                     expand = c(0,0)) +
  guides(
    color = guide_legend(
      title = "Group",
      override.aes = list(fill = NA)
    ),
    fill = guide_legend(title = "Confidence Interval")
  ) +
  labs(title = "Cluster 2",
       x = "Time Point",
       y = "Phosphorylation Log Fold Change") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, hjust = 0.5),  # Center title
        panel.grid = element_blank(),       # remove all gridlines
        legend.title = element_blank(),
        legend.position = "bottom",
        axis.ticks = element_line(color = "black"),
        legend.key = element_rect(fill = "transparent", color = NA)
        )
Cluster2P_Graph_YESLONO + guides(alpha = guide_none())
```

#Protein Abundance Analysis for Lono Figure
```{r}
##FOR LONO
# STEM Data was imported

# Install packages if needed
library(dplyr)

# Define your sets of columns
cols_Lono <- c("lono231009_CA_phosphorylation", "lono231009_CC_phosphorylation", "lono231009_CD_phosphorylation", "lono231016_PC6C_phosphorylation", "lono231016_PC6D_phosphorylation")
cols_NotLono <- c("kolohe231011_CA_phosphorylation", "kolohe231011_CC_phosphorylation", "kolohe231011_CD_phosphorylation", "kolohe231016_PC6A_phosphorylation", "kolohe231016_PC6B_phosphorylation", "kolohe231016_PC6C_phosphorylation", "kolohe231016_PC6D_phosphorylation", "liho231012_CA_phosphorylation", "liho231012_CC_phosphorylation", "liho231012_CD_phosphorylation", "liho231018_PC6A_phosphorylation", "liho231018_PC6B_phosphorylation", "liho231018_PC6D_phosphorylation", "nainoa231011_CA_phosphorylation", "nainoa231011_CC_phosphorylation", "nainoa231011_CD_phosphorylation", "nainoa231016_PC6A_phosphorylation", "nainoa231016_PC6B_phosphorylation", "nainoa231016_PC6D_phosphorylation", "hoku231009_CA_phosphorylation", "hoku231009_CC_phosphorylation", "hoku231009_CD_phosphorylation", "hoku231019_PC6A_phosphorylation", "hoku231019_PC6C_phosphorylation", "hoku231019_PC6D_phosphorylation")

# Making sure the columns are numeric
DQO_Abundance_Data[cols_Lono] <- lapply(DQO_Abundance_Data[cols_Lono], as.numeric)
DQO_Abundance_Data[cols_NotLono] <- lapply(DQO_Abundance_Data[cols_NotLono], as.numeric)

# Function to test difference row-by-row (via t-test) while accounting for any NAs
compare_groups <- function(row) {
  lono <- as.numeric(row[cols_Lono])
  notLono <- as.numeric(row[cols_NotLono])
  
  # Remove NAs from both groups
  lono <- lono[!is.na(lono)]
  notLono <- notLono[!is.na(notLono)]
  
  # Ensure both groups have enough data points to perform the test
  if (length(lono) < 2 || length(notLono) < 2) {
    return(NA)  # Return NA if there aren't enough valid data points
  }
  
  # Perform the t-test (Welch’s t-test for unequal variance)
  test <- tryCatch(
    t.test(lono, notLono)$p.value, 
    error = function(e) NA
  )
  
  return(test)
}

# Apply the function to each row
DQO_Abundance_Data$p_value_Lono <- apply(DQO_Abundance_Data, 1, compare_groups)

# Filter rows with significant differences (e.g., p < 0.05)
significant_rows_Lono <- DQO_Abundance_Data %>% filter(p_value_Lono < 0.05)

# Sorting rows by biggest difference between Lono values and non-Lono values
calculate_difference <- function(row) {
  lono <- as.numeric(row[cols_Lono])
  notLono <- as.numeric(row[cols_NotLono])
  
  # Remove NA values if present
  lono <- lono[!is.na(lono)]
  notLono <- notLono[!is.na(notLono)]
  
  if (length(lono) < 1 || length(notLono) < 1) return(NA)
  
  # Calculate the absolute differences between each corresponding value in group1 and group2
  differences <- abs(mean(lono) - mean(notLono))  # Mean difference or sum of abs differences can be used
  
  return(differences)
}
significant_rows_Lono$mean_difference <- apply(significant_rows_Lono, 1, calculate_difference)
significant_rows_Lono_sorted <- significant_rows_Lono[order(-significant_rows_Lono$mean_difference), ]

# View the sorted dataframe
head(significant_rows_Lono_sorted)

# Export the dataframe to an Excel file
install.packages("writexl")
library(writexl)
write_xlsx(significant_rows_Lono_sorted, "significant_rows.xlsx") #in Desktop saved as "DQO Significantly Different Lono Abundance Data"

##FOR KOLOHE
# Define your sets of columns
cols_Kolohe <- c("kolohe231011_CA_phosphorylation", "kolohe231011_CC_phosphorylation", "kolohe231011_CD_phosphorylation", "kolohe231016_PC6A_phosphorylation", "kolohe231016_PC6B_phosphorylation", "kolohe231016_PC6C_phosphorylation", "kolohe231016_PC6D_phosphorylation")
cols_NotKolohe <- c("liho231012_CA_phosphorylation", "liho231012_CC_phosphorylation", "liho231012_CD_phosphorylation", "liho231018_PC6A_phosphorylation", "liho231018_PC6B_phosphorylation", "liho231018_PC6D_phosphorylation", "nainoa231011_CA_phosphorylation", "nainoa231011_CC_phosphorylation", "nainoa231011_CD_phosphorylation", "nainoa231016_PC6A_phosphorylation", "nainoa231016_PC6B_phosphorylation", "nainoa231016_PC6D_phosphorylation", "lono231009_CA_phosphorylation", "lono231009_CC_phosphorylation", "lono231009_CD_phosphorylation", "lono231016_PC6C_phosphorylation", "lono231016_PC6D_phosphorylation", "hoku231009_CA_phosphorylation", "hoku231009_CC_phosphorylation", "hoku231009_CD_phosphorylation", "hoku231019_PC6A_phosphorylation", "hoku231019_PC6C_phosphorylation", "hoku231019_PC6D_phosphorylation")

# Making sure the columns are numeric
DQO_Abundance_Data[cols_Kolohe] <- lapply(DQO_Abundance_Data[cols_Kolohe], as.numeric)
DQO_Abundance_Data[cols_NotKolohe] <- lapply(DQO_Abundance_Data[cols_NotKolohe], as.numeric)

# Function to test difference row-by-row with NA handling
compare_groups <- function(row) {
  kolohe <- as.numeric(row[cols_Kolohe])
  notKolohe <- as.numeric(row[cols_NotKolohe])
  
  # Remove NAs from both groups
  kolohe <- kolohe[!is.na(kolohe)]
  notKolohe <- notKolohe[!is.na(notKolohe)]
  
  # Ensure both groups have enough data points to perform the test
  if (length(kolohe) < 2 || length(notKolohe) < 2) {
    return(NA)  # Return NA if there aren't enough valid data points
  }
  
  # Perform the t-test (Welch’s t-test for unequal variance)
  test <- tryCatch(
    t.test(kolohe, notKolohe)$p.value, 
    error = function(e) NA
  )
  
  return(test)
}

# Apply the function to each row
DQO_Abundance_Data$p_value_Kolohe <- apply(DQO_Abundance_Data, 1, compare_groups)

# Filter rows with significant differences (e.g., p < 0.05)
significant_rows_Kolohe <- DQO_Abundance_Data %>% filter(p_value_Kolohe < 0.05) #There being over 600 rows in this shows there are also significant differences in abundance between other individuals, which is expected given more interindividual variation than intraindividual
```

#Protein Phosphorylation Analysis for Lono Figure
```{r}
##FOR LONO
# STEM Data was imported

# Install packages if needed
library(dplyr)

# Define your sets of columns
cols_Lono_ <- c("lono231009_CA_phosphorylation", "lono231009_CC_phosphorylation", "lono231009_CD_phosphorylation", "lono231016_PC6C_phosphorylation", "lono231016_PC6D_phosphorylation")
cols_NotLono_ <- c("kolohe231011_CA_phosphorylation", "liho231012_CA_phosphorylation", "nainoa231011_CA_phosphorylation", "hoku231009_CA_phosphorylation", "kolohe231011_CC_phosphorylation", "liho231012_CC_phosphorylation", "nainoa231011_CC_phosphorylation", "hoku231009_CC_phosphorylation", "kolohe231011_CD_phosphorylation", "liho231012_CD_phosphorylation", "nainoa231011_CD_phosphorylation", "hoku231009_CD_phosphorylation", "kolohe231016_PC6A_phosphorylation", "liho231018_PC6A_phosphorylation", "nainoa231016_PC6A_phosphorylation", "hoku231019_PC6A_phosphorylation", "kolohe231016_PC6B_phosphorylation", "liho231018_PC6B_phosphorylation", "nainoa231016_PC6B_phosphorylation", "kolohe231016_PC6C_phosphorylation", "hoku231019_PC6C_phosphorylation", "kolohe231016_PC6D_phosphorylation", "liho231018_PC6D_phosphorylation", "nainoa231016_PC6D_phosphorylation", "hoku231019_PC6D_phosphorylation")

# Making sure the columns are numeric
DQO_Phosphorylation_Data[cols_Lono_] <- lapply(DQO_Phosphorylation_Data[cols_Lono_], as.numeric)
DQO_Phosphorylation_Data[cols_NotLono_] <- lapply(DQO_Phosphorylation_Data[cols_NotLono_], as.numeric)

# Function to test difference row-by-row with NA handling
compare_groups <- function(row) {
  lono_ <- as.numeric(row[cols_Lono_])
  notLono_ <- as.numeric(row[cols_NotLono_])
  
  # Remove NAs from both groups
  lono_ <- lono_[!is.na(lono_)]
  notLono_ <- notLono_[!is.na(notLono_)]
  
  # Ensure both groups have enough data points to perform the test
  if (length(lono_) < 2 || length(notLono_) < 2) {
    return(NA)  # Return NA if there aren't enough valid data points
  }
  
  # Perform the t-test (Welch’s t-test for unequal variance)
  test <- tryCatch(
    t.test(lono_, notLono_)$p.value, 
    error = function(e) NA
  )
  
  return(test)
}

# Apply the function to each row
DQO_Phosphorylation_Data$p_value_Lono_ <- apply(DQO_Phosphorylation_Data, 1, compare_groups)

# Filter rows with significant differences (e.g., p < 0.05)
significant_rows_Lono_ <- DQO_Phosphorylation_Data %>% filter(p_value_Lono_ < 0.05)

#ERRORS BELOW HERE WHERE NAS APPEARING IN ROWS
#####Hours: Monday 8:30 - 10:30pm
#####Hours: Tuesday 1:30 - 5:30pm

# Sorting rows by biggest difference between Lono values and non-Lono values
calculate_difference_ <- function(row) {
  lono_ <- as.numeric(row[cols_Lono_])
  notLono_ <- as.numeric(row[cols_NotLono_])
  
  # Remove NA values if present
  lono_ <- lono_[!is.na(lono_)]
  notLono_ <- notLono_[!is.na(notLono_)]
  
  if (length(lono_) < 1 || length(notLono_) < 1) return(NA)
  
  # Calculate the absolute differences between each corresponding value in group1 and group2
  differences_ <- abs(mean(lono_) - mean(notLono_))  # Mean difference or sum of abs differences can be used
  
  return(differences_)
}
significant_rows_Lono_$mean_difference_ <- apply(significant_rows_Lono_, 1, calculate_difference_)
significant_rows_Lono_sorted_ <- significant_rows_Lono_[order(-significant_rows_Lono_$mean_difference_), ]

# View the sorted dataframe
head(significant_rows_Lono_sorted_)

# Export the significant_rows dataframe to an Excel file
install.packages("writexl")
library(writexl)
write_xlsx(significant_rows_Lono_sorted_, "significant_rows_Lono_.xlsx") #in Desktop saved as "DQO Significantly Different Lono Phosphorylation Data"
```

##Thesis Venn Diagrams
```{r}
install.packages("VennDiagram")

library(VennDiagram)
library(grid)
library(gridExtra)

# Calculate scale factor (based on square root of area1 ratio)
scale_factor <- sqrt(8 / 17)

# --- Draw big Venn diagram ---
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))  # 2 columns for side-by-side

pushViewport(viewport(layout.pos.col = 1))  # Left panel
draw.pairwise.venn(
  area1 = 17,
  area2 = 169,
  cross.area = 5,
  category = c("Control", "Preconditioning"),
  fill = c("#4FAFFF", "#0059B2"), #0055FF #0033CC
  lty = "solid",
  lwd = 0.5,
  cex = 1,
  cat.cex = 1,
  cat.pos = c(180, 180),
  cat.fontface = "bold",
  cat.fontfamily = "Arial",
  ext.text = FALSE
)
upViewport()

#Upregulation Plot
#venn1 <- draw.pairwise.venn(
#  area1 = 17,         # Total in Set A
#  area2 = 169,          # Total in Set B
#  cross.area = 5,     # Overlap between A and B
#  category = c("Control", "Preconditioning"),
#  fill = c("#4FAFFF", "#0059B2"),
#  lty = "solid",
#  lwd = 0.5,
#  cex = 1,
#  cat.cex = 1,
#  cat.pos = c(180, 180),
#  cat.fontface = "bold",
#  cat.fontfamily = "Arial",
#  ext.text = FALSE
#)
#grid.draw(venn.plot)

# --- Draw small Venn diagram, scaled down ---
pushViewport(viewport(layout.pos.col = 2))
pushViewport(viewport(width = unit(scale_factor, "npc"),
                      height = unit(scale_factor, "npc"),
                      just = c("center", "center")))

draw.pairwise.venn(
  area1 = 8,
  area2 = 38,
  cross.area = 1,
  category = c("Control", "Preconditioning"),
  fill = c("#4FAFFF", "#0059B2"), #0055FF #0033CC
  lty = "solid",
  lwd = 0.5,
  cex = 1,
  cat.cex = 1,
  cat.pos = c(180, 180),
  cat.fontface = "bold",
  cat.fontfamily = "Arial",
  ext.text = FALSE
)
upViewport(2)  # Return to top level

#Downregulation Plot
#venn2 <- draw.pairwise.venn(
#  area1 = 8,         # Total in Set A
#  area2 = 38,          # Total in Set B
#  cross.area = 1,     # Overlap between A and B
#  category = c("Control", "Preconditioning"),
#  fill = c("#4FAFFF", "#0059B2"),
#  lty = "solid",
#  lwd = 0.5,
#  cex = 1,
#  cat.cex = 1,
#  cat.pos = c(180, 180),
#  cat.fontface = "bold",
#  cat.fontfamily = "Arial",
#  ext.text = FALSE,
#  scaled = FALSE
#)
#grid.draw(venn.plot)
```

##Thesis SAA1 Graph
```{r}
library(ggplot2)

# Create a data frame
df <- data.frame(
  x = c(76, 86, 94, 101, 76, 86, 94, 101, 52, 76, 101, 113, 129, 52, 119),
  y = c(0.99, 1.015, 1.04, 1.407, 4.635, 10.675, 5.083, 11.208, 0, 0, 0, 0, 0, 0, 0),
  group = c("Control", "Control", "Control", "Control", "Preconditioning", "Preconditioning", "Preconditioning", "Preconditioning", "Human", "Human", "Human", "Human", "Human", "Mouse", "Mouse")
)

# Convert 'group' to a factor for easier mapping
df$group <- factor(df$group, levels = c("Control", "Preconditioning", "Human", "Mouse"))

print(levels(df$group))

# Scatterplot with color by group
ggplot(df, aes(x = x, y = y, color = group, shape = group)) +
  geom_point(size = 4) +
  geom_segment(aes(x = 1, xend = 129, y = 0, yend = 0), color = "black") +  # adds a line at y = 0
  geom_segment(aes(x = 1, xend = 1, y = 0, yend = 12), color = "black") +  # moves the y-axis to x = 1
  geom_segment(data = data.frame(x = seq(1, 129, by = 8)),
               aes(x = x, xend = x, y = 0, yend = -0.2),  # short tick lines
               inherit.aes = FALSE, color = "black") +
  geom_segment(data = data.frame(y = seq(0, 12, by = 3)),
             aes(x = -0.5, xend = 1, y = y, yend = y),  # short horizontal tick lines
             inherit.aes = FALSE, color = "black") +
  geom_text(data = data.frame(x = seq(1, 129, by = 8)),
            aes(x = x, y = -0.3, label = x),
            inherit.aes = FALSE, size = 3, vjust = 1, family = "Arial") +
  geom_text(data = data.frame(y = seq(0, 12, by = 3)),
          aes(x = -1, y = y, label = y),
          inherit.aes = FALSE, hjust = 1, size = 3, family = "Arial") +
  scale_x_continuous(limits = c(-2, 129),
                     breaks = seq(1, 129, by = 8)) +  # Set x-axis breaks every 9 units
  scale_y_continuous(limits = c(-1, 12),
                     breaks = seq(0, 12, by = 3)) +
  scale_color_manual(values = c("Control" = "#4FAFFF", 
                                "Preconditioning" = "#0059B2", 
                                "Human" = "black", 
                                "Mouse" = "black")) +
  scale_shape_manual(values = c("Control" = 16, 
                                "Preconditioning" = 15, 
                                "Human" = 4, 
                                "Mouse" = 5)) +
  labs(title = "Serum Amyloid A1",
       x = "Residue Position",
       y = "Phosphorylation Fold Change",
       color = "",
       shape = "") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    panel.grid = element_blank(),       # remove all gridlines
    axis.ticks = element_blank(),       # remove axis ticks
    axis.line = element_blank(),        # remove axis lines
    axis.text = element_blank(),         # remove tick labels
    axis.title.x = element_text(margin = margin(t = -10)),
    axis.title.y = element_text(margin = margin(t = -25)),
    legend.position = "bottom"
  )
```

##Thesis Phosphosites Scatterplot Graph
#Upregulation Scatterplot
```{r}
library(ggplot2)

# Create a data frame
df <- data.frame(
  x = c("SLC16A1", "SLC16A1", "HBA", "HBA", "SAA1", "SAA1", "SPTB", "SPTB", " SLC16A1", " SLC16A1", " SAA1", " SAA1", "IGFBP3", "IGFBP3", "ANK1", "ANK1", " HBA", " HBA", "FGA", "FGA", "PA2G4", "PA2G4", "NAP1L4", "NAP1L4", "EPS15", "EPS15", "TSC22D4", "TSC22D4", " SPTB", " SPTB", " ANK1", " ANK1", " TSC22D4", " TSC22D4", " FGA", " FGA", "MYH10", "MYH10", "MYL4", "MYL4", "  HBA", "  HBA", "CFL1", "CFL1", " EPS15", " EPS15", "   HBA", "   HBA", "CLINT1", "CLINT1", "EPB41", "EPB41", "PRSS1", "PRSS1", "  EPS15", "  EPS15"),
  y = c(7.42938731, 0.62984009, 3.69622316, 0.36960694, 3.30785619, 0, 3.14069742, 0.56508183, 3.10620008, 0.51804794, 3.02813215, 16.3459269, 2.90793876, 0.23961012, 2.76107843, 0, 2.75468129, 1.81760526, 2.68643042, 0.75210154, 2.53002739, 0.32070362, 2.51555905, 0.32679964, 2.41818765, 0.18630192, 2.35404536, 0.25615893, 2.25341681, 0.13156614, 2.25329204, 0, 2.225437, 0.25368351, 2.21226777, 0.63376415, 2.20221013, 2.69347034, 2.13689806, 2.11664257, 2.12588342, 0.20891641, 2.11142403, 0.27770223, 2.07629022, 0.36575316, 2.07560252, 0.19207045, 2.07221873, 0.81491012, 2.06438586, 1.3370225, 2.06406801, 1.06419905, 2.02541472, 0.84647281),
  group = c("Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning")
)

# Convert 'group' to a factor for easier mapping
df$group <- factor(df$group, levels = c("Control", "Preconditioning"))

df$x <- factor(df$x, levels = c("SLC16A1", "HBA", "SAA1", "SPTB", " SLC16A1", " SAA1", "IGFBP3", "ANK1", " HBA", "FGA", "PA2G4", "NAP1L4", "EPS15", "TSC22D4", " SPTB", " ANK1", " TSC22D4", " FGA", "MYH10", "MYL4", "  HBA", "CFL1", " EPS15", "   HBA", "CLINT1", "EPB41", "PRSS1", "  EPS15"))

tick_data <- data.frame(
  x = c("SLC16A1", "SLC16A1", "HBA", "HBA", "SAA1", "SAA1", "SPTB", "SPTB", " SLC16A1", " SLC16A1", " SAA1", " SAA1", "IGFBP3", "IGFBP3", "ANK1", "ANK1", " HBA", " HBA", "FGA", "FGA", "PA2G4", "PA2G4", "NAP1L4", "NAP1L4", "EPS15", "EPS15", "TSC22D4", "TSC22D4", " SPTB", " SPTB", " ANK1", " ANK1", " TSC22D4", " TSC22D4", " FGA", " FGA", "MYH10", "MYH10", "MYL4", "MYL4", "  HBA", "  HBA", "CFL1", "CFL1", " EPS15", " EPS15", "   HBA", "   HBA", "CLINT1", "CLINT1", "EPB41", "EPB41", "PRSS1", "PRSS1", "  EPS15", "  EPS15"),  # your x-axis category labels
  y = 0,
  yend = -0.2
)

# Scatterplot with color by group
ggplot(df, aes(x = x, y = y, color = group, shape = group)) +
  geom_point(size = 3) +
  geom_segment(aes(x = 0, xend = "  EPS15", y = 0, yend = 0), color = "black") +  # adds a line at y = 0
  geom_segment(aes(x = 0, xend = 0, y = 0, yend = 17), color = "black") +  # moves the y-axis to x = 1
  geom_segment(data = tick_data,
               aes(x = x, xend = x, y = 0, yend = -0.4),
               inherit.aes = FALSE,
               color = "black") +
  geom_segment(data = data.frame(y = seq(0, 16, by = 4)),
             aes(x = -0.4, xend = 0, y = y, yend = y),  # short horizontal tick lines
             inherit.aes = FALSE, color = "black") +
  geom_text(data = data.frame(x = c("SLC16A1", "HBA", "SAA1", "SPTB", " SLC16A1", " SAA1", "IGFBP3", "ANK1", " HBA", "FGA", "PA2G4", "NAP1L4", "EPS15", "TSC22D4", " SPTB", " ANK1", " TSC22D4", " FGA", "MYH10", "MYL4", "  HBA", "CFL1", " EPS15", "   HBA", "CLINT1", "EPB41", "PRSS1", "  EPS15")),
          aes(x = x, y = -0.6, label = x),
          inherit.aes = FALSE,
          size = 3, vjust = 1, family = "Arial", angle = 90, hjust = 1) +
  geom_text(data = data.frame(y = seq(0, 16, by = 4)),
          aes(x = -0.75, y = y, label = y),
          inherit.aes = FALSE, hjust = 1, size = 3, family = "Arial") +
  #scale_x_continuous(limits = c(-2, 129),
  #                   breaks = seq(1, 129, by = 8)) +  # Set x-axis breaks every 9 units
  #scale_y_continuous(limits = c(-1, 12),
  #                   breaks = seq(0, 12, by = 3)) +
  scale_x_discrete(expand = expansion(mult = c(0.12, 0.05))) +
  scale_y_discrete(expand = expansion(mult = c(0.25, 0.05))) +
  scale_color_manual(values = c("Control" = "#4FAFFF", 
                                "Preconditioning" = "#0059B2")) +
  scale_shape_manual(values = c("Control" = 16, 
                                "Preconditioning" = 15)) +
  labs(title = "Upregulated Control Phosphosites",
       x = "Phosphosite",
       y = "Phosphorylation Fold Change",
       color = "",
       shape = "") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    panel.grid = element_blank(),       # remove all gridlines
    axis.ticks = element_blank(),       # remove axis ticks
    axis.line = element_blank(),        # remove axis lines
    axis.text = element_blank(),         # remove tick labels
    legend.position = "bottom"
  )
```

#Downregulation Scatterplot
```{r}
library(ggplot2)

# Create a data frame
df <- data.frame(
  x = c("SAA1", "SAA1", "PAK2", "PAK2", "TOM1L1", "TOM1L1", "DNAJA4", "DNAJA4", "ZYX", "ZYX", "BASP1", "BASP1", "CALD1", "CALD1", "HBB", "HBB", " CALD1", " CALD1"),
  y = c(0.38057582, 0, 0.40194081, 2.31296024, 0.45588263, 1.66919687, 0.46005947, 1.10872561, 0.46157609, 0.88343128, 0.46424218, 1.76142088, 0.47385638, 1.00421085, 0.48461013, 0.64952936, 0.49079909, 0.75805743),
  group = c("Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning", "Control", "Preconditioning")
)

# Convert 'group' to a factor for easier mapping
df$group <- factor(df$group, levels = c("Control", "Preconditioning"))

df$x <- factor(df$x, levels = c("SAA1", "PAK2", "TOM1L1", "DNAJA4", "ZYX", "BASP1", "CALD1", "HBB", " CALD1"))

tick_data <- data.frame(
  x = c("SAA1", "SAA1", "PAK2", "PAK2", "TOM1L1", "TOM1L1", "DNAJA4", "DNAJA4", "ZYX", "ZYX", "BASP1", "BASP1", "CALD1", "CALD1", "HBB", "HBB", " CALD1", " CALD1"),  # your x-axis category labels
  y = 0,
  yend = -0.2
)

# Scatterplot with color by group
ggplot(df, aes(x = x, y = y, color = group, shape = group)) +
  geom_point(size = 3) +
  geom_segment(aes(x = 0, xend = " CALD1", y = 0, yend = 0), color = "black") +  # adds a line at y = 0
  geom_segment(aes(x = 0, xend = 0, y = 0, yend = 3), color = "black") +  # moves the y-axis to x = 1
  geom_segment(data = tick_data,
               aes(x = x, xend = x, y = 0, yend = -0.1),
               inherit.aes = FALSE,
               color = "black") +
  geom_segment(data = data.frame(y = seq(0, 3, by = 1)),
             aes(x = -0.1, xend = 0, y = y, yend = y),  # short horizontal tick lines
             inherit.aes = FALSE, color = "black") +
  geom_text(data = data.frame(x = c("SAA1", "PAK2", "TOM1L1", "DNAJA4", "ZYX", "BASP1", "CALD1", "HBB", " CALD1")),
          aes(x = x, y = -0.15, label = x),
          inherit.aes = FALSE,
          size = 3, vjust = 1, family = "Arial", angle = 90, hjust = 1) +
  geom_text(data = data.frame(y = seq(0, 3, by = 1)),
          aes(x = -0.15, y = y, label = y),
          inherit.aes = FALSE, hjust = 1, size = 3, family = "Arial") +
  #scale_x_continuous(limits = c(-2, 129),
  #                   breaks = seq(1, 129, by = 8)) +  # Set x-axis breaks every 9 units
  #scale_y_continuous(limits = c(-1, 12),
  #                   breaks = seq(0, 12, by = 3)) +
  scale_x_discrete(expand = expansion(mult = c(0.2, 0.05))) +
  scale_y_discrete(expand = expansion(mult = c(0.25, 0.05))) +
  scale_color_manual(values = c("Control" = "#4FAFFF", 
                                "Preconditioning" = "#0059B2")) +
  scale_shape_manual(values = c("Control" = 16, 
                                "Preconditioning" = 15)) +
  labs(title = "Downregulated Control Phosphosites",
       x = "Phosphosite",
       y = "Phosphorylation Fold Change",
       color = "",
       shape = "") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    panel.grid = element_blank(),       # remove all gridlines
    axis.ticks = element_blank(),       # remove axis ticks
    axis.line = element_blank(),        # remove axis lines
    axis.text = element_blank(),         # remove tick labels
    legend.position = "bottom"
  )
```

##Thesis Abundance Venn Diagrams
```{r}
install.packages("VennDiagram")

library(VennDiagram)
library(grid)
library(gridExtra)

# Calculate scale factor (based on square root of area1 ratio)
scale_factor <- sqrt(13 / 23)
scale_factor1 <- sqrt(12 / 23)
scale_factor2 <- sqrt(19 / 23)

# --- Set up 2x2 grid ---
grid.newpage()
pushViewport(viewport(layout = grid.layout(2, 2)))  # 2 columns for side-by-side

pushViewport(viewport(layout.pos.row = 1, layout.pos.col = 1))  # Left panel
pushViewport(viewport(width = unit(scale_factor, "npc"),
                      height = unit(scale_factor, "npc"),
                      just = c("center", "center")))
draw.pairwise.venn(
  area1 = 13.001,
  area2 = 17.001,
  cross.area = 0.001,
  category = c("Phosphorylation", "Abundance"),
  fill = c("#4FAFFF", "#E5E5E5"), #0059B2
  lty = "solid",
  lwd = 0.5,
  cex = 0.5,
  cat.cex = 0.5,
  cat.pos = c(0, 0),
  cat.dist = c(0.09, 0.09),
  cat.fontfamily = "Arial",
  ext.text = FALSE,
  label.col = c("black", NA, "black")  # Hide the middle/intersection label
)
upViewport(2)

# --- Draw small Venn diagram, scaled down ---
pushViewport(viewport(layout.pos.row = 1, layout.pos.col = 2))
pushViewport(viewport(width = unit(scale_factor1, "npc"),
                      height = unit(scale_factor1, "npc"),
                      just = c("center", "center")))

draw.pairwise.venn(
  area1 = 12.001,
  area2 = 18.001,
  cross.area = 0.001,
  category = c("Phosphorylation", "Abundance"),
  fill = c("#4FAFFF", "#E5E5E5"), #0059B2
  lty = "solid",
  lwd = 0.5,
  cex = 0.5,
  cat.cex = 0.5,
  cat.pos = c(0, 0),
  cat.dist = c(0.06, 0.06),
  cat.fontfamily = "Arial",
  ext.text = FALSE,
  label.col = c("black", NA, "black")  # Hide the middle/intersection label
)
upViewport(2)

# --- Draw 3rd Venn diagram, scaled down ---
pushViewport(viewport(layout.pos.row = 2, layout.pos.col = 1))
pushViewport(viewport(width = unit(1, "npc"),
                      height = unit(1, "npc"),
                      just = c("center", "center")))
draw.pairwise.venn(
  area1 = 23.001,
  area2 = 169.001,
  cross.area = 0.001,
  category = c("Phosphorylation", "Abundance"),
  fill = c("#0059B2", "#E5E5E5"), #0059B2
  lty = "solid",
  lwd = 0.5,
  cex = 0.5,
  cat.cex = 0.5,
  cat.pos = c(0, 0),
  cat.fontfamily = "Arial",
  ext.text = FALSE,
  label.col = c("black", NA, "black")  # Hide the middle/intersection label
)
upViewport(2)  # Return to top level

# --- Draw 4th Venn diagram, scaled down ---
pushViewport(viewport(layout.pos.row = 2, layout.pos.col = 2))
pushViewport(viewport(width = unit(scale_factor2, "npc"),
                      height = unit(scale_factor2, "npc"),
                      just = c("center", "center")))

draw.pairwise.venn(
  area1 = 19.001,
  area2 = 38.001,
  cross.area = 0.001,
  category = c("Phosphorylation", "Abundance"),
  fill = c("#0059B2", "#E5E5E5"), #0059B2
  lty = "solid",
  lwd = 0.5,
  cex = 0.5,
  cat.cex = 0.5,
  cat.pos = c(0, 0),
  cat.dist = c(0.05, 0.05),
  cat.fontfamily = "Arial",
  ext.text = FALSE,
  label.col = c("black", NA, "black")  # Hide the middle/intersection label
)
upViewport(2)  # Return to top level
```

##Thesis Abundance Scatterplots
#Control Upregulated
```{r}
#install.packages("ggpmisc")

library(ggplot2)
library(ggpmisc)

# Create a data frame
df <- data.frame(
  x = c(7.42938731, 3.69622316, 3.30785619, 3.14069742, 3.10620008, 3.02813215, 2.90793876, 2.76107843, 2.75468129, 2.68643042, 2.53002739, 2.51555905, 2.41818765, 2.35404536, 2.25341681, 2.25329204, 2.225437, 2.21226777, 2.20221013, 2.13689806, 2.12588342, 2.11142403, 2.07629022, 2.07560252, 2.07221873, 2.06438586, 2.06406801, 2.02541472),
  y = c(1.023256, 0.985171, 1, 0.992314, 1.023256, 1, 0.966102, 0.995614, 0.985171, 0.988466, 0.823529, 1, 0.98, 0.93617, 0.992314, 0.995614, 0.93617, 0.988466, 0.989286, 0.963636, 0.985171, 0.983471, 0.98, 0.985171, 0.842105, 1, 0.993695, 0.98),
  group = c("Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control")
)

# Convert 'group' to a factor for easier mapping
df$group <- factor(df$group, levels = c("Control"))

# Scatterplot with color by group
ggplot(df, aes(x = x, y = y, color = group, shape = group)) +
  geom_point(size = 2) +
  geom_segment(aes(x = 0, xend = 8, y = 0, yend = 0), color = "black") +  # adds a line at y = 0
  geom_segment(aes(x = 0, xend = 0, y = 0, yend = 2), color = "black") +  # moves the y-axis to x = 1
  geom_segment(data = data.frame(x = seq(0, 8, by = 2)),
               aes(x = x, xend = x, y = 0, yend = -0.05),  # short tick lines
               inherit.aes = FALSE, color = "black") +
  geom_segment(data = data.frame(y = seq(0, 2, by = 0.5)),
             aes(x = -0.2, xend = 0, y = y, yend = y),  # short horizontal tick lines
             inherit.aes = FALSE, color = "black") +
  geom_text(data = data.frame(x = seq(0, 8, by = 2)),
            aes(x = x, y = -0.1, label = x),
            inherit.aes = FALSE, size = 3, vjust = 1, family = "Arial") +
  geom_text(data = data.frame(y = seq(0, 2, by = 0.5)),
          aes(x = -0.3, y = y, label = y),
          inherit.aes = FALSE, hjust = 1, size = 3, family = "Arial") +
  geom_smooth(method = "lm", se = FALSE, color = "black", size = 0.5) +
  annotate(
    "text",
    x = 6.5, y = 1.8,             # <-- Adjust these manually based on your data
    label = "italic(R)^2 == 0.07774",
    parse = TRUE,
    color = "black",
    size = 3) +
  #scale_x_continuous(limits = c(-2, 10),
  #                   breaks = seq(0, 8, by = 2)) +  # Set x-axis breaks every 9 units
  #scale_y_continuous(limits = c(-1, 2),
  #                   breaks = seq(0, 2, by = 0.5)) +
  scale_color_manual(values = c("Control" = "#4FAFFF")) +
  scale_shape_manual(values = c("Control" = 16)) +
  labs(x = "Phosphorylation Fold Change",
       y = "Protein Abundance Fold Change",
       color = "",
       shape = "") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    panel.grid = element_blank(),       # remove all gridlines
    axis.ticks = element_blank(),       # remove axis ticks
    axis.line = element_blank(),        # remove axis lines
    axis.text = element_blank(),         # remove tick labels
    #axis.title.x = element_text(margin = margin(t = -10)),
    #axis.title.y = element_text(margin = margin(t = -25)),
    legend.position = "bottom"
  )
```

#Control Downregulated
```{r}
#install.packages("ggpmisc")

library(ggplot2)
library(ggpmisc)

# Create a data frame
df <- data.frame(
  x = c(0.38057582, 0.40194081, 0.46005947, 0.46157609, 0.46424218, 0.47385638, 0.48461013, 0.49079909),
  y = c(1, 0.980392, 1, 1.015873, 0.875, 1.031579, 0.984688, 1.031579),
  group = c("Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control"))

# Convert 'group' to a factor for easier mapping
df$group <- factor(df$group, levels = c("Control"))

# Scatterplot with color by group
ggplot(df, aes(x = x, y = y, color = group, shape = group)) +
  geom_point(size = 2) +
  geom_segment(aes(x = 0, xend = 0.5, y = 0, yend = 0), color = "black") +  # adds a line at y = 0
  geom_segment(aes(x = 0, xend = 0, y = 0, yend = 2), color = "black") +  # moves the y-axis to x = 1
  geom_segment(data = data.frame(x = seq(0, 0.5, by = 0.1)),
               aes(x = x, xend = x, y = 0, yend = -0.05),  # short tick lines
               inherit.aes = FALSE, color = "black") +
  geom_segment(data = data.frame(y = seq(0, 2, by = 0.5)),
             aes(x = -0.015, xend = 0, y = y, yend = y),  # short horizontal tick lines
             inherit.aes = FALSE, color = "black") +
  geom_text(data = data.frame(x = seq(0, 0.5, by = 0.1)),
            aes(x = x, y = -0.1, label = x),
            inherit.aes = FALSE, size = 3, vjust = 1, family = "Arial") +
  geom_text(data = data.frame(y = seq(0, 2, by = 0.5)),
          aes(x = -0.02, y = y, label = y),
          inherit.aes = FALSE, hjust = 1, size = 3, family = "Arial") +
  geom_smooth(method = "lm", se = FALSE, color = "black", size = 0.5) +
  annotate(
    "text",
    x = 0.4, y = 1.8,             # <-- Adjust these manually based on your data
    label = "italic(R)^2 == 0.005543",
    parse = TRUE,
    color = "black",
    size = 3) +
  #scale_x_continuous(limits = c(-2, 10),
  #                   breaks = seq(0, 8, by = 2)) +  # Set x-axis breaks every 9 units
  #scale_y_continuous(limits = c(-1, 2),
  #                   breaks = seq(0, 2, by = 0.5)) +
  scale_color_manual(values = c("Control" = "#4FAFFF")) +
  scale_shape_manual(values = c("Control" = 16)) +
  labs(x = "Phosphorylation Fold Change",
       y = "Protein Abundance Fold Change",
       color = "",
       shape = "") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    panel.grid = element_blank(),       # remove all gridlines
    axis.ticks = element_blank(),       # remove axis ticks
    axis.line = element_blank(),        # remove axis lines
    axis.text = element_blank(),         # remove tick labels
    #axis.title.x = element_text(margin = margin(t = -10)),
    #axis.title.y = element_text(margin = margin(t = -25)),
    legend.position = "bottom"
  )
```

#Preconditioning Upregulated
```{r}
#install.packages("ggpmisc")

library(ggplot2)
library(ggpmisc)

# Create a data frame
df <- data.frame(
  x = c(18.5967159, 16.7150515, 16.3459269, 10.710667, 10.1572067, 10.1348594, 7.94080191, 7.56060843, 6.69107234, 5.94366214, 5.70178727, 5.56302898, 5.32404647, 5.24301149, 5.05927489, 4.62545444, 4.25388858, 4.20924069, 3.77473887, 3.41305926, 3.36579152, 3.20495703, 3.16429037, 3.05986237, 3.02427112, 2.90977606, 2.90632634, 2.8559226, 2.81584955, 2.74978313, 2.74927546, 2.74115201, 2.70223005, 2.69347034, 2.67485879, 2.65775488, 2.64148402, 2.634041, 2.62896998, 2.61249843, 2.59751202, 2.58316295, 2.58315838, 2.57740708, 2.50933525, 2.50749474, 2.49976957, 2.49866888, 2.49741802, 2.49253404, 2.49131606, 2.48236687, 2.47488421, 2.46132277, 2.44660881, 2.43777099, 2.43564605, 2.43209636, 2.43136613, 2.40717111, 2.39593722, 2.38225401, 2.38033055, 2.36840258, 2.36274811, 2.35612029, 2.34757856, 2.32309107, 2.32098585, 2.31445986, 2.31296024, 2.30435771, 2.29540032, 2.28095367, 2.27192412, 2.26615973, 2.26572035, 2.2549461, 2.2500197, 2.24062436, 2.23564712, 2.23470921, 2.20313265, 2.19809085, 2.1974391, 2.18718199, 2.17666522, 2.16940519, 2.16919717, 2.16291889, 2.14494634, 2.14142126, 2.137413, 2.13735623, 2.12519967, 2.12007687, 2.11664257, 2.11516053, 2.11391589, 2.11202726, 2.1024141, 2.09854956, 2.08534491, 2.08527426, 2.0848936, 2.08330156, 2.07547169, 2.0739278, 2.07010777, 2.05919193, 2.04919677, 2.04901929, 2.04830575, 2.04829498, 2.04250326, 2.03956244, 2.03825815, 2.03471899, 2.03297128, 2.03293885, 2.03100317, 2.02821745, 2.02589779, 2.02210107, 2.02078217, 2.01743467, 2.01533742, 2.01485805, 2.01163008, 2.01082548, 2.0089481),
  y = c(1.380952, 1.566667, 1.380952, 1.412229, 1.380952, 1.566667, 1.380952, 0.833333, 1.380952, 1.412229, 1.412229, 1.117647, 1.566667, 1.566667, 1.380952, 1.566667, 1.412229, 1.380952, 1.412229, 1.117647, 0.974359, 1.117647, 1.412229, 1.104762, 1.117647, 0.761905, 1.244444, 1, 0.8, 1.019608, 1.412229, 1.079365, 0.761905, 1.412229, 1.104762, 1.117647, 1.412229, 1, 0, 1.052632, 1.212121, 0.966667, 1.037645, 1.045045, 0.989247, 1.079365, 1.044057, 1.007092, 1.117647, 1.041667, 0.983165, 1, 1.412229, 1.117647, 0.985008, 0.666667, 1.006289, 1, 1, 1.006289, 1.079365, 1.412229, 0.952381, 2, 1, 1.005682, 1.130435, 1, 0.955556, 1.055556, 0.952381, 1.122807, 1.333333, 1.005682, 1.019608, 1.049479, 1, 1.044057, 0.952381, 1, 0.933333, 0.982456, 1.333333, 1.045045, 1.055556, 1.047619, 1.117647, 1.666667, 0.761905, 1.122807, 1.333333, 1.066667, 0.952381, 1.005682, 1.006536, 1.037037, 1.128205, 1, 1.005682, 1.212121, 1, 1.111111, 1.212121, 1.333333, 0.998723, 1.047619, 1.244444, 0.8, 1.566667, 0.99115, 1.076923, 1.130435, 0.8, 0.952381, 0.982456, 1, 1.005682, 1.117647, 1.333333, 1.005682, 1, 0.761905, 1.226667, 1.051282, 1.333333, 0.96124, 1.08642, 1.130435, 0.761905, 1, 1.066667),
  group = c("Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control")
)

# Convert 'group' to a factor for easier mapping
df$group <- factor(df$group, levels = c("Control"))

# Scatterplot with color by group
ggplot(df, aes(x = x, y = y, color = group, shape = group)) +
  geom_point(size = 2) +
  geom_segment(aes(x = 0, xend = 20, y = 0, yend = 0), color = "black") +  # adds a line at y = 0
  geom_segment(aes(x = 0, xend = 0, y = 0, yend = 2), color = "black") +  # moves the y-axis to x = 1
  geom_segment(data = data.frame(x = seq(0, 20, by = 4)),
               aes(x = x, xend = x, y = 0, yend = -0.05),  # short tick lines
               inherit.aes = FALSE, color = "black") +
  geom_segment(data = data.frame(y = seq(0, 2, by = 0.5)),
             aes(x = -0.2, xend = 0, y = y, yend = y),  # short horizontal tick lines
             inherit.aes = FALSE, color = "black") +
  geom_text(data = data.frame(x = seq(0, 20, by = 4)),
            aes(x = x, y = -0.1, label = x),
            inherit.aes = FALSE, size = 3, vjust = 1, family = "Arial") +
  geom_text(data = data.frame(y = seq(0, 2, by = 0.5)),
          aes(x = -0.3, y = y, label = y),
          inherit.aes = FALSE, hjust = 1, size = 3, family = "Arial") +
  geom_smooth(method = "lm", se = FALSE, color = "black", size = 0.5) +
  annotate(
    "text",
    x = 6.5, y = 1.8,             # <-- Adjust these manually based on your data
    label = "italic(R)^2 == 0.1464",
    parse = TRUE,
    color = "black",
    size = 3) +
  #scale_x_continuous(limits = c(-2, 10),
  #                   breaks = seq(0, 8, by = 2)) +  # Set x-axis breaks every 9 units
  #scale_y_continuous(limits = c(-1, 2),
  #                   breaks = seq(0, 2, by = 0.5)) +
  scale_color_manual(values = c("Control" = "#0059B2")) +
  scale_shape_manual(values = c("Control" = 15)) +
  labs(x = "Phosphorylation Fold Change",
       y = "Protein Abundance Fold Change",
       color = "",
       shape = "") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    panel.grid = element_blank(),       # remove all gridlines
    axis.ticks = element_blank(),       # remove axis ticks
    axis.line = element_blank(),        # remove axis lines
    axis.text = element_blank(),         # remove tick labels
    #axis.title.x = element_text(margin = margin(t = -10)),
    #axis.title.y = element_text(margin = margin(t = -25)),
    legend.position = "bottom"
  )
```

#Preconditioning Downregulated
```{r}
#install.packages("ggpmisc")

library(ggplot2)
library(ggpmisc)

# Create a data frame
df <- data.frame(
  x = c(0.06101108, 0.13108949, 0.13156614, 0.15591981, 0.16372599, 0.16648762, 0.18630192, 0.19207045, 0.20891641, 0.21029231, 0.23487939, 0.2367399, 0.23774993, 0.23961012, 0.25049901, 0.25368351, 0.25615893, 0.27770223, 0.29912401, 0.32070362, 0.32404736, 0.32679964, 0.330186, 0.33310022, 0.34782921, 0.35811585, 0.36232138, 0.36234299, 0.36241781, 0.36575316, 0.36960694, 0.37126844, 0.38121568, 0.38772374, 0.39404199, 0.42894696, 0.4343351, 0.44585101, 0.46460766, 0.4704672, 0.47579299, 0.48625162, 0.49020639, 0.49393062, 0.49475707, 0.49478146, 0.49887844),
  y = c(0.666667, 0.91954, 0.985008, 0.91954, 0.985008, 0.95405, 1.247312, 0.95405, 0.95405, 0.974359, 0.988201, 0.95405, 1.007092, 0.944444, 0.95405, 0.982456, 0.982456, 1.007092, 1.007092, 0.833333, 0.888889, 0.888889, 0.820513, 1.005682, 1, 0.981779, 1.005682, 0.95405, 0.888889, 1.247312, 0.95405, 1, 0.988201, 0.95405, 0.666667, 0.806202, 0.988201, 0.888889, 0.666667, 0.926554, 0.987342, 1, 1.005682, 1.052632, 0.806202, 0.974565, 0.926554),
  group = c("Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control", "Control"))

# Convert 'group' to a factor for easier mapping
df$group <- factor(df$group, levels = c("Control"))

# Scatterplot with color by group
ggplot(df, aes(x = x, y = y, color = group, shape = group)) +
  geom_point(size = 2) +
  geom_segment(aes(x = 0, xend = 0.5, y = 0, yend = 0), color = "black") +  # adds a line at y = 0
  geom_segment(aes(x = 0, xend = 0, y = 0, yend = 2), color = "black") +  # moves the y-axis to x = 1
  geom_segment(data = data.frame(x = seq(0, 0.5, by = 0.1)),
               aes(x = x, xend = x, y = 0, yend = -0.05),  # short tick lines
               inherit.aes = FALSE, color = "black") +
  geom_segment(data = data.frame(y = seq(0, 2, by = 0.5)),
             aes(x = -0.015, xend = 0, y = y, yend = y),  # short horizontal tick lines
             inherit.aes = FALSE, color = "black") +
  geom_text(data = data.frame(x = seq(0, 0.5, by = 0.1)),
            aes(x = x, y = -0.1, label = x),
            inherit.aes = FALSE, size = 3, vjust = 1, family = "Arial") +
  geom_text(data = data.frame(y = seq(0, 2, by = 0.5)),
          aes(x = -0.02, y = y, label = y),
          inherit.aes = FALSE, hjust = 1, size = 3, family = "Arial") +
  geom_smooth(method = "lm", se = FALSE, color = "black", size = 0.5) +
  annotate(
    "text",
    x = 0.4, y = 1.8,             # <-- Adjust these manually based on your data
    label = "italic(R)^2 == 0.003486",
    parse = TRUE,
    color = "black",
    size = 3) +
  #scale_x_continuous(limits = c(-2, 10),
  #                   breaks = seq(0, 8, by = 2)) +  # Set x-axis breaks every 9 units
  #scale_y_continuous(limits = c(-1, 2),
  #                   breaks = seq(0, 2, by = 0.5)) +
  scale_color_manual(values = c("Control" = "#0059B2")) +
  scale_shape_manual(values = c("Control" = 15)) +
  labs(x = "Phosphorylation Fold Change",
       y = "Protein Abundance Fold Change",
       color = "",
       shape = "") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    panel.grid = element_blank(),       # remove all gridlines
    axis.ticks = element_blank(),       # remove axis ticks
    axis.line = element_blank(),        # remove axis lines
    axis.text = element_blank(),         # remove tick labels
    #axis.title.x = element_text(margin = margin(t = -10)),
    #axis.title.y = element_text(margin = margin(t = -25)),
    legend.position = "bottom"
  )
```

##Thesis Circle Packing Plot for Pathways
#Control Pathways
```{r}
#install.packages("packcircles")
#install.packages("ggforce")
#install.packages("circlize")
#install.packages("scales")

library(packcircles)
library(ggplot2)
library(ggforce)
library(dplyr)
library(circlize)
library(tidyr)
library(scales)
library(ggrepel)  # for handling text labels with automatic wrapping
library(stringr)  # for wrapping text

# Original data
data <- data.frame(
  parent = c("Inflammation", "Inflammation", "Innate Immunity", "Innate Immunity", "Innate Immunity", 
             "Innate Immunity", "Innate Immunity", "Innate Immunity", "Migration", "Migration", 
             "Migration", "Migration"),
  label = c("regulation of inflammatory response", "inflammatory response", "platelet activation", 
            "negative regulation of immune system process", "negative regulation of defense response", 
            "negative regulation of response to external stimulus", 
            "myeloid cell activation involved in immune response", "regulation of immune response", 
            "regulation of actin cytoskeleton organization", "actin filament bundle assembly", 
            "cell chemotaxis", "leukocyte migration"),
  p_value = c(6.10E-04, 9.40E-04, 0.01, 0.01, 2.50E-03, 8.30E-03, 9.70E-03, 0.02, 0.02, 
           3.40E-03, 5.80E-04, 4.60E-04),
  size = c(6.10E-04, 9.40E-04, 0.01, 0.01, 2.50E-03, 8.30E-03, 9.70E-03, 0.02, 0.02, 
              3.40E-03, 5.80E-04, 4.60E-04)
)

# Log scale and size rescaling
data <- data %>%
  mutate(
    log_size = log(size + 1e-6),
    scaled_size = rescale(log_size, to = c(1, 10))
  )

# Layout for each subgroup
nested_layout <- data %>%
  group_by(parent) %>%
  do({
    layout <- circleProgressiveLayout(.$scaled_size, sizetype = "area")
    bind_cols(., layout)
  }) %>%
  ungroup()

# Offset groups horizontally to prevent overlap
nested_layout <- nested_layout %>%
  mutate(parent_id = as.numeric(factor(parent)))

group_spacing <- 15
nested_layout <- nested_layout %>%
  mutate(
    x = x + parent_id * group_spacing,
    y = y
  )

# --- NEW: Compute enclosing circles for each parent group ---
parent_bounds <- nested_layout %>%
  group_by(parent, parent_id) %>%
  summarize(
    center_x = mean(x),
    center_y = mean(y),
    group_radius = max(sqrt((x - mean(x))^2 + (y - mean(y))^2 + radius^2)) + 2,
    .groups = "drop"
  )

# Step 7: Wrap child labels using str_wrap() from stringr
nested_layout <- nested_layout %>%
  mutate(label_wrapped = str_wrap(label, width = 30))  # Adjust width as needed

# Plot nested + enclosing circles
ggplot() +
  # Group enclosing circles
  ggforce::geom_circle(data = parent_bounds, 
                       aes(x0 = center_x, y0 = center_y, r = (group_radius - 1)), 
                       color = "black", fill = NA, linewidth = 0.5, linetype = "solid") +
  # Child circles
  ggforce::geom_circle(data = nested_layout,
                       aes(x0 = x, y0 = y, r = 0.8, alpha = p_value),  # fixed radius
                       fill = "#4FAFFF", color = "white", linewidth = 0.1) +
  # Labels for child pathways
  geom_text_repel(data = nested_layout,
                  aes(x = x, y = y, label = label_wrapped), size = 2.5, max.overlaps = 10) +
  # Parent labels
  geom_text(data = parent_bounds,
            aes(x = center_x, y = center_y + group_radius + 1, label = parent),
            size = 4) +
  # Custom legend with circles (using geom_point instead of geom_circle)
  # Custom legend with circles (using geom_point instead of geom_circle)
  #geom_point(data = nested_layout, 
  #           aes(x = x, y = y, size = radius), 
  #           shape = 16,  # solid circles in the legend
  #           color = "#4FAFFF", 
   #          alpha = 0.7, 
  #           show.legend = TRUE) +  
  # Customizing the legend to ensure size appears properly as circles
  guides(p_value = guide_legend(
    title = "p-value", 
    override.aes = list(shape = 15, fill = "#4FAFFF", color = "#4FAFFF", size = 6)  # solid circle in legend
  )) +
  scale_alpha_continuous(
  name = "p-value",
  range = c(1, 0.3)  # low p = high opacity (dark); high p = low opacity (light)
) +
  coord_fixed() +
  theme_void() +
  theme(legend.position = "bottom")

```

#Preconditioning Pathways
```{r}
#install.packages("packcircles")
#install.packages("ggforce")
#install.packages("circlize")
#install.packages("scales")

library(packcircles)
library(ggplot2)
library(ggforce)
library(dplyr)
library(circlize)
library(tidyr)
library(scales)
library(ggrepel)  # for handling text labels with automatic wrapping
library(stringr)  # for wrapping text

# Original data
data <- data.frame(
  parent = c("GPCR Signaling", "GPCR Signaling", "GPCR Signaling", "GPCR Signaling", "GPCR Signaling", 
             "Coagulation", "Coagulation", "mRNA Processing", "mRNA Processing", "mRNA Processing", 
             "mRNA Processing"),
  label = c("positive regulation of response to stimulus", "G protein-coupled receptor signaling pathway", "GTPase regulator activity", "small GTPase-mediated signal transduction", "regulation of small GTPase mediated signal transduction", "negative regulation of coagulation", "regulation of blood coagulation", "mRNA processing", "RNA biosynthetic process", "RNA splicing", "regulation of RNA splicing"),
  p_value = c(1.00E-03, 1.40E-04, 2.70E-04, 3.60E-05, 1.10E-05, 8.50E-05, 3.90E-04, 6.00E-04, 4.20E-05, 1.20E-04, 3.40E-04),
  size = c(1.00E-03, 1.40E-04, 2.70E-04, 3.60E-05, 1.10E-05, 8.50E-05, 3.90E-04, 6.00E-04, 4.20E-05, 1.20E-04, 3.40E-04)
)

# Log scale and size rescaling
data <- data %>%
  mutate(
    log_size = log(size + 1e-6),
    scaled_size = rescale(log_size, to = c(1, 10))
  )

# Layout for each subgroup
nested_layout <- data %>%
  group_by(parent) %>%
  do({
    layout <- circleProgressiveLayout(.$scaled_size, sizetype = "area")
    bind_cols(., layout)
  }) %>%
  ungroup()

# Offset groups horizontally to prevent overlap
nested_layout <- nested_layout %>%
  mutate(parent_id = as.numeric(factor(parent)))

group_spacing <- 15
nested_layout <- nested_layout %>%
  mutate(
    x = x + parent_id * group_spacing,
    y = y
  )

# --- NEW: Compute enclosing circles for each parent group ---
parent_bounds <- nested_layout %>%
  group_by(parent, parent_id) %>%
  summarize(
    center_x = mean(x),
    center_y = mean(y),
    group_radius = max(sqrt((x - mean(x))^2 + (y - mean(y))^2 + radius^2)) + 2,
    .groups = "drop"
  )

# Step 7: Wrap child labels using str_wrap() from stringr
nested_layout <- nested_layout %>%
  mutate(label_wrapped = str_wrap(label, width = 30))  # Adjust width as needed

# Plot nested + enclosing circles
ggplot() +
  # Group enclosing circles
  ggforce::geom_circle(data = parent_bounds, 
                       aes(x0 = center_x, y0 = center_y, r = (group_radius - 1)), 
                       color = "black", fill = NA, linewidth = 0.5, linetype = "solid") +
  # Child circles
  ggforce::geom_circle(data = nested_layout,
                       aes(x0 = x, y0 = y, r = 0.8, alpha = p_value),  # fixed radius
                       fill = "#0059B2", color = "white", linewidth = 0.1) +
  # Labels for child pathways
  geom_text_repel(data = nested_layout,
                  aes(x = x, y = y, label = label_wrapped), size = 2.5, max.overlaps = 10) +
  # Parent labels
  geom_text(data = parent_bounds,
            aes(x = center_x, y = center_y + group_radius + 1, label = parent),
            size = 4) +
  # Custom legend with circles (using geom_point instead of geom_circle)
  # Custom legend with circles (using geom_point instead of geom_circle)
  #geom_point(data = nested_layout, 
  #           aes(x = x, y = y, size = radius), 
  #           shape = 16,  # solid circles in the legend
  #           color = "#4FAFFF", 
   #          alpha = 0.7, 
  #           show.legend = TRUE) +  
  # Customizing the legend to ensure size appears properly as circles
  guides(p_value = guide_legend(
    title = "p-value", 
    override.aes = list(shape = 15, fill = "#0059B2", color = "#0059B2", size = 6)  # solid circle in legend
  )) +
  scale_alpha_continuous(
  name = "p-value",
  range = c(1, 0.3)  # low p = high opacity (dark); high p = low opacity (light)
) +
  coord_fixed() +
  theme_void() +
  theme(legend.position = "bottom")

```

#NO LONO Lollipop Graph
```{r}
BiocManager::install("org.Hs.eg.db", ask = FALSE)

remotes::install_github("G3viz/g3viz")
library(g3viz)

phospho_data <- data.frame(
  Hugo_Symbol = c("P69905", "P69905"),         # Use UniProt as gene symbol
  AA_Position = c(3, 13),                      # Amino acid positions
  Protein_Change = c("p.S3p", "p.T13p"),
  Mutation_Class = c("Phospho", "Phospho"),
  stringsAsFactors = FALSE
)

# Optional: define plot options if you want to customize colors or labels
options <- g3Lollipop.options()
options$colorSchema <- list(Phospho = "#FF5733")

g3Lollipop(
  mutation.dat = phospho_data,
  gene.symbol = "HBA1",           # Gene name (HGNC symbol)
#  uniprot.id = "P69905",          # UniProt ID if known
  factor.col = "Mutation_Class",
  plot.options = options,
  output.filename = "phospho_plot"
)

```
