---
title: "main"
author: "Ashley Blawas"
date: '2024-06-13'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries include=FALSE}
library(tidyverse)
library(readxl)
library(stringr)
library(dplyr)
library(tidyr)
library(ComplexHeatmap)
library(RColorBrewer)
library(circlize)


# Set home for figures
fig_path = "C:/Users/Ashley Blawas/Dropbox/Ashley/Post-Doc/Projects/DQO preconditioning/figs"

```
# From working with original dataset...
```{r  include=FALSE}

# Where does data live?
data_path = "C:/Users/ashle/Dropbox/Ashley/Post-Doc/Projects/DQO preconditioning/data"

# What is the file name?
file_name = "10364_Supplemental_Data_061224.xlsx"

# What are the sheet names
sheet1 = "S1. Unenriched Metadata"
sheet2 = "S2. Uneriched Abundances" # NP = # of precursors, PGQ = protein group quantity
sheet3 = "S3. Phos Metadata"
sheet4 = "S4. Phos localized" # Place where it was on the protein
sheet5 = "S5. Phos not localized" # Not able to place where it was on the protein
sheet6 = "S6. Phos filtered" # 


```

## Load in the metadata
```{r}
# Load in the Unenriched Metadata
unenriched_metadata <- read_excel(paste0(data_path, "/", file_name), sheet = sheet1)

# Make the Animal IDs for the controls unique
n_na = sum(is.na(unenriched_metadata$Condition))
unenriched_metadata$Condition[is.na(unenriched_metadata$Condition)] <- paste0("NA_VALUE_", seq_len(n_na))

# Load in the Phosphorylation Metadata - had to add an "h" to one of the "tech" in phos metadata
phos_metadata <- read_excel(paste0(data_path, "/", file_name), sheet = sheet3)

# Make the Animal IDs for the controls unique
n_na = sum(phos_metadata$Replicate == "SPQC_prep")
phos_metadata$Replicate[phos_metadata$Replicate == "SPQC_prep"] <- paste0("SPQC_prep", seq_len(n_na))

n_na = sum(phos_metadata$Replicate == "SPQC_tech")
phos_metadata$Replicate[phos_metadata$Replicate == "SPQC_tech"] <- paste0("SPQC_tech", seq_len(n_na))
```

## Load in the unenriched dataset
```{r}
# Load in the Unenriched Metadata
unenriched_abundances <- read_excel(paste0(data_path, "/", file_name), sheet = sheet2)

for (i in 1:nrow(unenriched_metadata)){
  curr_name = names(unenriched_abundances)[i+3] 
  curr_id = unlist(str_extract_all(curr_name, "ID.+htrms"))
  
  temp_metadata = unenriched_metadata %>%
  filter(str_detect(`File Name`, curr_id))
  
  curr_animal = temp_metadata$Animal
  curr_condition = temp_metadata$Condition
  
  unenriched_abundances <- unenriched_abundances %>% rename(!! paste0(curr_animal, "_", curr_condition, "_", "NP") := curr_name)
  
}


for (i in 1:nrow(unenriched_metadata)){
  curr_name = names(unenriched_abundances)[i+36] 
  curr_id = unlist(str_extract_all(curr_name, "ID.+htrms"))
  
  temp_metadata = unenriched_metadata %>%
  filter(str_detect(`File Name`, curr_id))
  
  curr_animal = temp_metadata$Animal
  curr_condition = temp_metadata$Condition
  
  unenriched_abundances <- unenriched_abundances %>% rename(!! paste0(curr_animal, "_", curr_condition, "_", "PGQ") := curr_name)
  
}

```

## PCA of unenriched abundances
```{r}

# Step 1: Make one column (e.g., 'ID') the new header
new_header <- unenriched_abundances$PG.ProteinNames # Take the values of the 'ID' column
unenriched_abundances_t <- unenriched_abundances[, -c(1, 2, 3)]  # Remove the 'ID' column from the original data frame

# Select only columns that contain "PGQ" in the name
unenriched_abundances_t <- unenriched_abundances_t[, grep("PGQ", colnames(unenriched_abundances_t))]

# Step 2: Transpose the data frame
unenriched_abundances_t <- t(unenriched_abundances_t)

# Step 3: Set the new header
colnames(unenriched_abundances_t) <- new_header

# Step 4: Convert the transposed matrix back to a data frame & to numeric entries
unenriched_abundances_t <- as.data.frame(unenriched_abundances_t)
unenriched_abundances_t[] <- lapply(unenriched_abundances_t, as.numeric)

# Step 5: Impute NaNs with column medians
# Impute with column medians
unenriched_abundances_t <- unenriched_abundances_t
for (i in 1:ncol(unenriched_abundances_t)) {
  unenriched_abundances_t[is.na(unenriched_abundances_t[, i]), i] <- 0
}

# Step 6: Run PCA on the numeric abundance data
pca_result <- prcomp(unenriched_abundances_t, scale. = TRUE)

# Step 7: Check the summary of PCA result to view the variance explained by each principal component
#summary(pca_result)

# Step 8: Visualize the first two principal components in a biplot
# This plot shows how each sample (row) is positioned based on the first two principal components
#biplot(pca_result)

# Step 9: Create a PCA plot with points colored by the Protein Group if available
# First, add the PG.ProteinGroups as a factor variable to the PCA results
pca_df <- data.frame(pca_result$x)  # Extract the PCA scores for the samples

# Extract the rownames and add them as a variable to the PCA output
row_names <- rownames(unenriched_abundances_t)
pca_df <- pca_df %>%
  mutate(animal = str_extract(row_names, "^[^0-9]*")) %>%
  mutate(animalday = sub("([^_]+).*", "\\1", row_names)) %>%
  mutate(trialtype = str_extract(row_names, "^[^_]+_([^_]+)")) %>%
  mutate(trialtype = sub("^[^_]+_", "", trialtype)) %>%
  filter(trialtype != "PC6B")

# Filter out rows where 'group' contains 'lono'
pca_df <- pca_df %>% filter(!grepl("lono", animal))

# Add in animal codes
pca_df$animal <- gsub("hoku", "63H4", pca_df$animal)
pca_df$animal <- gsub("kolohe", "6JK5", pca_df$animal)
pca_df$animal <- gsub("liho", "01L5", pca_df$animal)
pca_df$animal <- gsub("nainoa", "9ON6", pca_df$animal)

# Replace weird codes for QC
pca_df$trialtype <- gsub("NA", "QC", pca_df$trialtype)

# Now, use ggplot2 for a more customized PCA plot
ggplot(pca_df, aes(x = PC1, y = PC2, color = animal, shape = trialtype)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(title = "PCA of Protein Abundance Data", x = "PC1", y = "PC2") +
  scale_color_manual(values = c('63H4' = '#CC6677',
                                #'hoku231019' = '#CC6699',
                                '6JK5' = '#DDCC77',
                                #'kolohe231016' = '#DDCC22',
                                '01L5' = '#117733',
                                #'liho231018' = '#116655',
                                #'lono231009' = '#6699CC',
                                #'lono231016' = '#6688CC',
                                '9ON6' = '#EE8866',
                                #'nainoa231016' = '#EE8888',
                                'SPQC' = 'black'), name = "Animal") +  
    scale_shape_manual(values = c(15, 16, 17, 0, 1, 2, 3, 4, 5, 6), name = "Trial Type") +  # Manually assign the shapes
  theme_minimal() +
   geom_hline(yintercept = 0, linetype = "dotted", color = "black", size = 1) +  # Dotted horizontal line at y = 0
  geom_vline(xintercept = 0, linetype = "dotted", color = "black", size = 1) + 
  theme(
    # Adding black box around the plot
    #plot.background = element_rect(color = "black", fill = NA, size = 1), # border size and color
    panel.border = element_rect(color = "black", fill = NA, size = 1), # border around the plot area (panel)
    legend.spacing = unit(0.05, "lines"),  # Adjust spacing between legend entries
    legend.key.size = unit(0.4, "cm"),   # Adjust the size of the legend key
    legend.title = element_text(size = 10), # Adjust the legend title size
    legend.text = element_text(size = 8)   # Adjust the legend text size
  )

filename <- "PCA_unenrichedabundances"
  
 # Save the plot (for example, as a PNG)
ggsave(paste0(fig_path, "/", filename, ".png"), units="in", width = 5, height = 3, dpi = 600, bg = "white")
cat(paste("File", filename, "has been saved.\n"))

```

## Load in the filtered phosphoprotein dataset
```{r}
# Load in the Unenriched Metadata
phos_filtered <- read_excel(paste0(data_path, "/", file_name), sheet = sheet6)

# Drop CV columns
phos_filtered <- phos_filtered %>% dplyr::select(!starts_with("CV"))

# Change data frame organization to have names
for (i in 1:nrow(phos_metadata)){
  curr_name = names(phos_filtered)[i+5] # Add 5 because there are 5 columns that are not IDs
  curr_id = unlist(str_extract_all(curr_name, "ID.+htrms"))
  
  temp_metadata = phos_metadata %>%
  filter(str_detect(`File Name`, curr_id))
  
  curr_animal = temp_metadata$Condition
  curr_condition = temp_metadata$Replicate
  
  phos_filtered <- phos_filtered %>% rename(!! paste0(curr_animal, "_", curr_condition, "_", "EG") := curr_name)
  
}

```


# Make data frame for plotting

```{r}

phos_filtered <- phos_filtered %>% select(-PG.ProteinAccessions, -PG.ProteinDescriptions, -PG.ProteinNames, -EG.PrecursorId)

phos_filtered <- phos_filtered %>% pivot_longer(!PG.Genes, names_to = "Sample") %>%
  rename(Value = value) %>% 
  rename(Gene = PG.Genes)

# Mutate new columns
phos_filtered <- phos_filtered %>% mutate(Animal = str_extract(Sample, "^\\D+")) %>%
  mutate(Date = str_extract(Sample, "\\d+")) %>%
  mutate(Condition = sub("^[^_]*_([^_]*).*", "\\1", Sample)) %>%
  dplyr::select(-Sample) %>%
  dplyr::select(Gene, Animal, Condition, Date, Value)
  

```

## PCA of phosphoenriched abundances
```{r}

# Step 1: Make one column (e.g., 'ID') the new header
new_header <- phos_filtered$PG.ProteinAccessions # Take the values of the 'ID' column
phos_filtered_t <- phos_filtered[, -c(1, 2, 3, 4, 5)]  # Remove the 'ID' column from the original data frame

# Step 2: Transpose the data frame
phos_filtered_t <- t(phos_filtered_t)

# Step 3: Set the new header
colnames(phos_filtered_t) <- new_header

# Step 4: Convert the transposed matrix back to a data frame & to numeric entries
phos_filtered_t <- as.data.frame(phos_filtered_t)
phos_filtered_t[] <- lapply(phos_filtered_t, as.numeric)

# Step 5: Impute NaNs with column medians
# Impute with column medians
phos_filtered_t <- phos_filtered_t
for (i in 1:ncol(phos_filtered_t)) {
  phos_filtered_t[is.na(phos_filtered_t[, i]), i] <- 0
}

# Step 6: Run PCA on the numeric abundance data
pca_result <- prcomp(phos_filtered_t, scale. = TRUE)

# Step 7: Check the summary of PCA result to view the variance explained by each principal component
#summary(pca_result)

# Step 8: Visualize the first two principal components in a biplot
# This plot shows how each sample (row) is positioned based on the first two principal components
#biplot(pca_result)

# Step 9: Create a PCA plot with points colored by the Protein Group if available
# First, add the PG.ProteinGroups as a factor variable to the PCA results
pca_df <- data.frame(pca_result$x)  # Extract the PCA scores for the samples

# Extract the rownames and add them as a variable to the PCA output
row_names <- rownames(phos_filtered_t)
pca_df <- pca_df %>%
  mutate(animal = str_extract(row_names, "^[^0-9]*")) %>%
  mutate(animalday = sub("([^_]+).*", "\\1", row_names)) %>%
  mutate(trialtype = str_extract(row_names, "^[^_]+_([^_]+)")) %>%
  mutate(trialtype = sub("^[^_]+_", "", trialtype))%>%
  filter(trialtype != "PC6B")

# Filter out rows where 'group' contains 'lono'
pca_df <- pca_df %>% filter(!grepl("lono", animal))

# Add in animal codes
pca_df$animal <- gsub("hoku", "63H4", pca_df$animal)
pca_df$animal <- gsub("kolohe", "6JK5", pca_df$animal)
pca_df$animal <- gsub("liho", "01L5", pca_df$animal)
pca_df$animal <- gsub("nainoa", "9ON6", pca_df$animal)

# Replace weird codes for QC
pca_df$trialtype <- gsub("prep", "QC1", pca_df$trialtype)
pca_df$trialtype <- gsub("tech", "QC2", pca_df$trialtype)


# Now, use ggplot2 for a more customized PCA plot
ggplot(pca_df, aes(x = PC1, y = PC2, color = animal, shape = trialtype)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(title = "PCA of Phospho-Enriched Data", x = "PC1", y = "PC2") +
  scale_color_manual(values = c('63H4' = '#CC6677',
                                #'hoku231019' = '#CC6699',
                                '6JK5' = '#DDCC77',
                                #'kolohe231016' = '#DDCC22',
                                '01L5' = '#117733',
                                #'liho231018' = '#116655',
                                #'lono231009' = '#6699CC',
                                #'lono231016' = '#6688CC',
                                '9ON6' = '#EE8866',
                                #'nainoa231016' = '#EE8888',
                                'SPQC' = 'black'), name = "Animal") +  
  scale_shape_manual(values = c(15, 16, 17, 0, 1, 2, 3, 4, 5, 6), name = "Trial Type") +  # Manually assign the shapes
  theme_minimal() +
   geom_hline(yintercept = 0, linetype = "dotted", color = "black", size = 1) +  # Dotted horizontal line at y = 0
  geom_vline(xintercept = 0, linetype = "dotted", color = "black", size = 1) + 
  theme(
    # Adding black box around the plot
    #plot.background = element_rect(color = "black", fill = NA, size = 1), # border size and color
    
    panel.border = element_rect(color = "black", fill = NA, size = 1), # border around the plot area (panel)
    legend.spacing = unit(0.05, "lines"),  # Adjust spacing between legend entries
    legend.key.size = unit(0.4, "cm"),   # Adjust the size of the legend key
    legend.title = element_text(size = 10), # Adjust the legend title size
    legend.text = element_text(size = 8)   # Adjust the legend text size
  )

filename <- "PCA_phosphoenriched"
  
 # Save the plot (for example, as a PNG)
ggsave(paste0(fig_path, "/", filename, ".png"), units="in", width = 5, height = 3, dpi = 600, bg = "white")
cat(paste("File", filename, "has been saved.\n"))

```

## From working with corrected dataset...
```{r}

df_abundance <- read_excel("C:/Users/Ashley Blawas/Dropbox/Ashley/Post-Doc/Projects/DQO preconditioning/data/combined_df.xlsx", 
    sheet = "abundance")

df_phospho <- read_excel("C:/Users/Ashley Blawas/Dropbox/Ashley/Post-Doc/Projects/DQO preconditioning/data/combined_df.xlsx", 
    sheet = "phospho")

```

## Make ABUNDANCE data frame actually useful...
```{r}
# Reshape the abundance data
df_abundance_long <- df_abundance %>%
  pivot_longer(
    cols = 3:32,  # Adjust this if needed
    names_to = c("AnimalName", "Condition", ".value"),
    names_sep = "_"
  )

# Rename the coumns
df_abundance_long <- df_abundance_long  %>%
  rename(
    ProteinCode = PG.ProteinGroups,
    ProteinName = PG.ProteinDescriptions.x
  )

```

## Make PHOSPHO data frame actually useful...
```{r}
# Reshape the abundance data
df_phospho <- df_phospho %>%
  rename_with(~ gsub("^(PTM\\.|PG\\.)", "", .), 
              .cols = starts_with("PTM.") | starts_with("PG.")) %>%
  rename(
    SiteLocation = SiteLocation.x,
ProteinName = ProteinDescriptions.y)


df_phospho_long <- df_phospho %>%
  pivot_longer(
    cols = 9:98,
    names_to = "temp",
    values_to = "Value"
  ) %>%
  # Separate the combined column name into 3 parts
  mutate(
    MeasurementType = case_when(
      str_ends(temp, "Log2_Normalized_Phosphorylation") ~ "Log2NormalizedPhosphorylation",
      str_ends(temp, "Normalized_Phosphorylation") ~ "NormalizedPhosphorylation",
      str_ends(temp, "Phosphorylation") ~ "Phosphorylation",
      TRUE ~ "Other"
    ),
    # Remove the measurement suffix to isolate AnimalName and Condition
    temp_clean = case_when(
      MeasurementType == "Log2NormalizedPhosphorylation" ~ str_remove(temp, "_Log2_Normalized_Phosphorylation"),
      MeasurementType == "NormalizedPhosphorylation" ~ str_remove(temp, "_Normalized_Phosphorylation"),
      MeasurementType == "Phosphorylation" ~ str_remove(temp, "_Phosphorylation"),
      TRUE ~ temp
    )
  ) %>%
  separate(temp_clean, into = c("AnimalName", "Condition"), sep = "_") %>%
  select(-temp)

```

# Heatmap for abundance
```{r}


# Step 1: Filter to Condition == "CA", ensure numeric Abundance
df_ca <- df_abundance_long %>%
  filter(Condition == "CA", !is.na(Abundance)) %>%
  mutate(Abundance = as.numeric(Abundance))

# Step 2: Identify top quartile proteins by total abundance
top_proteins <- df_ca %>%
  group_by(ProteinName) %>%
  summarize(total_abundance = sum(Abundance), .groups = "drop") %>%
  arrange(desc(total_abundance)) %>%
  mutate(rank = row_number()) %>%
  filter(rank <= ceiling(n() * 0.25)) %>% # This is where we set the top X percentile
  pull(ProteinName)

# Step 3: Log2-transform and pivot with aggregation
df_matrix <- df_ca %>%
  filter(ProteinName %in% top_proteins) %>%
  mutate(logAbundance = log2(Abundance + 1)) %>%
  select(ProteinName, AnimalName, logAbundance) %>%
  pivot_wider(
    names_from = AnimalName,
    values_from = logAbundance,
    values_fill = 0,
    values_fn = mean  # ✅ This resolves duplicates
  )

# Step 4: Convert to matrix
df_matrix <- as.data.frame(df_matrix)

# Now it's safe to set rownames
rownames(df_matrix) <- df_matrix$ProteinName

abundance_mat <- df_matrix %>%
  select(-ProteinName) %>%
  mutate(across(everything(), as.numeric)) %>%
  as.matrix()

# Create short protein names (first 20 chars)
full_names <- rownames(abundance_mat)
short_names <- substr(full_names, 1, 20)


# Now plot heatmap with these shortened names
abundance_hm <- Heatmap(
  t(abundance_mat),
  name = "log2(Abundance + 1)",
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  col = colorRamp2(
    breaks = seq(10, 30, length.out = 5),
    colors = brewer.pal(5, "Reds")
  ),
  column_names_gp = grid::gpar(fontsize = 8),
  column_names_rot = 45,  
  heatmap_legend_param = list(title = "log2(Abundance+1)"),
  column_labels = short_names
)

abundance_hm

filename <- "heatmap_abundance"

#  # Save the plot (for example, as a PNG) - 1%
# ggsave(paste0(fig_path, "/", filename, "_top1.png"), units="in", width = 3, height = 2, dpi = 600)
# cat(paste("File", filename, "has been saved.\n"))
# # 
#  # Save the plot (for example, as a PNG) - 5%
# ggsave(paste0(fig_path, "/", filename, "_top5.png"), units="in", width = 9, height = 2, dpi = 600)
# cat(paste("File", filename, "has been saved.\n"))
# 
#  # Save the plot (for example, as a PNG) - 25%
# ggsave(paste0(fig_path, "/", filename, "_top25.png"), units="in", width = 9, height = 2, dpi = 600)
# cat(paste("File", filename, "has been saved.\n"))

png(paste0(fig_path, "/heatmap_abundance_top25.png"), width = 2400, res = 150)
draw(abundance_hm, 
  heatmap_height = unit(6, "cm"))
dev.off()

```

## PCA - Abundance
```{r}

abundance_df <- df_abundance %>%
  select(contains("_Abundance")) %>%
  distinct() %>%
  mutate(across(everything(), ~ log2(as.numeric(.) + 1))) %>%
  rowwise() %>%
  mutate(across(everything(), ~ ifelse(is.na(.), median(c_across(everything()), na.rm = TRUE), .))) %>%
  ungroup()

  
# Transpose: columns become rows (samples)
transposed <- t(abundance_df)

# Get sample names from original colnames
sample_names <- colnames(df_abundance)[grepl("_Abundance", colnames(df_abundance))]

# Extract metadata: Animal and Condition
metadata <- tibble(
  sample = sample_names,
  Animal = str_extract(sample_names, "^[^_]+"),
  Condition = str_extract(sample_names, "(?<=_)[^_]+")
)

# Set row names for transposed data
rownames(transposed) <- sample_names

# Remove constant columns (zero variance proteins)
transposed_clean <- transposed[, apply(transposed, 2, function(x) sd(x, na.rm = TRUE) != 0)]

# Run PCA
pca_result <- prcomp(transposed_clean, scale. = TRUE)

# Get PCA scores (coordinates of samples on principal components)
pca_scores <- as.data.frame(pca_result$x)

# Add sample names as a column for joining metadata
pca_scores <- pca_scores %>%
  mutate(sample = rownames(pca_scores))

# Join metadata
pca_plot_df <- bind_cols(pca_scores, metadata)

# Add in animal codes
pca_plot_df$Animal <- gsub("Hoku", "63H4", pca_plot_df$Animal)
pca_plot_df$Animal <- gsub("Kolohe", "6JK5", pca_plot_df$Animal)
pca_plot_df$Animal <- gsub("Liho", "01L5", pca_plot_df$Animal)
pca_plot_df$Animal <- gsub("Nainoa", "9ON6", pca_plot_df$Animal)
pca_plot_df$Animal <- gsub("Lono", "9FL3", pca_plot_df$Animal)

# Drop Lono PC6C and PC6D
pca_plot_df <- pca_plot_df %>%
  filter(!(Animal == "9FL3" & Condition %in% c("PC6C", "PC6D")))

# Basic ggplot PCA scatter plot
abundance_pca <- ggplot(pca_plot_df, aes(x = PC1, y = PC2, color = as.factor(Animal), shape = Condition)) +
  geom_point(size = 4, alpha = 0.8) +
  labs(title = "Protein Abundance PCA",
       x = paste0("PC1 (", round(summary(pca_result)$importance[2,1]*100, 1), "% variance)"),
       y = paste0("PC2 (", round(summary(pca_result)$importance[2,2]*100, 1), "% variance)")) +
  scale_color_manual(values = c('63H4' = '#CC6677',
                                #'hoku231019' = '#CC6699',
                                '6JK5' = '#DDCC77',
                                #'kolohe231016' = '#DDCC22',
                                '01L5' = '#117733',
                                #'liho231018' = '#116655',
                                #'lono231009' = '#6699CC',
                                '9FL3' = '#6688CC',
                                '9ON6' = '#EE8866',
                                #'nainoa231016' = '#EE8888',
                                'SPQC' = 'black'), name = "Animal") +  
  scale_shape_manual(values = c(15, 16, 17, 0, 1, 2, 3), name = "Trial Type") +  # Manually assign the shapes
  theme_minimal() +
   geom_hline(yintercept = 0, linetype = "dotted", color = "black", size = 1) +  # Dotted horizontal line at y = 0
  geom_vline(xintercept = 0, linetype = "dotted", color = "black", size = 1) + 
  theme(
    # Adding black box around the plot
    #plot.background = element_rect(color = "black", fill = NA, size = 1), # border size and color
    panel.border = element_rect(color = "black", fill = NA, size = 1), # border around the plot area (panel)
    legend.spacing = unit(0.05, "lines"),  # Adjust spacing between legend entries
    legend.key.size = unit(0.4, "cm"),   # Adjust the size of the legend key
    legend.title = element_text(size = 10), # Adjust the legend title size
    legend.text = element_text(size = 8)   # Adjust the legend text size
  ) +
    theme(plot.title = element_text(hjust = 0.5),
    legend.title = element_text(face = "bold")+
    theme_minimal()

  )


ggsave(paste0(fig_path, "/", "pca_abundance", ".png"), units="in", width = 5, height = 3.5, dpi = 600, bg = "white")


```


## PCA phospho
```{r}
# Assuming your data frame is called 'df_phospho'
# Remove non-numeric columns for PCA

phospho_df <- df_phospho %>%
  select(matches("^(?!.*Log2).*_Normalized_Phosphorylation", perl = TRUE)) %>%
  distinct() %>%
  mutate(across(everything(), ~ as.numeric(.))) %>%
  rowwise() %>%
  filter(!if_all(everything(), is.na)) %>%
  mutate(across(everything(), ~ ifelse(is.na(.), median(c_across(everything()), na.rm = TRUE), .))) %>%
  ungroup()

# Transpose: columns become rows (samples)
transposed <- t(phospho_df)

# Get sample names from original colnames
sample_names <- colnames(df_phospho)[grepl("_Log2_Normalized_Phosphorylation", colnames(df_phospho))]

# Extract metadata: Animal and Condition
metadata <- tibble(
  sample = sample_names,
  Animal = str_extract(sample_names, "^[^_]+"),
  Condition = str_extract(sample_names, "(?<=_)[^_]+")
)

# Set row names for transposed data
rownames(transposed) <- sample_names

# Remove constant columns (zero variance proteins)
transposed_clean <- transposed[, apply(transposed, 2, function(x) sd(x, na.rm = TRUE) != 0)]

# Run PCA
pca_result <- prcomp(transposed_clean, scale. = TRUE)

# Get PCA scores (coordinates of samples on principal components)
pca_scores <- as.data.frame(pca_result$x)

# Add sample names as a column for joining metadata
pca_scores <- pca_scores %>%
  mutate(sample = rownames(pca_scores))

# Join metadata
pca_plot_df <- bind_cols(pca_scores, metadata)

# Add in animal codes
pca_plot_df$Animal <- gsub("Hoku", "63H4", pca_plot_df$Animal)
pca_plot_df$Animal <- gsub("Kolohe", "6JK5", pca_plot_df$Animal)
pca_plot_df$Animal <- gsub("Liho", "01L5", pca_plot_df$Animal)
pca_plot_df$Animal <- gsub("Nainoa", "9ON6", pca_plot_df$Animal)
pca_plot_df$Animal <- gsub("Lono", "9FL3", pca_plot_df$Animal)

# Drop Lono PC6C and PC6D
pca_plot_df <- pca_plot_df %>%
  filter(!(Animal == "9FL3" & Condition %in% c("PC6C", "PC6D")))

# Basic ggplot PCA scatter plot
phospho_pca <- ggplot(pca_plot_df, aes(x = PC1, y = PC2, color = as.factor(Animal), shape = Condition)) +
  geom_point(size = 4, alpha = 0.8) +
  labs(title = "Normalized Phosphoprotein PCA",
       x = paste0("PC1 (", round(summary(pca_result)$importance[2,1]*100, 1), "% variance)"),
       y = paste0("PC2 (", round(summary(pca_result)$importance[2,2]*100, 1), "% variance)")) +
  scale_color_manual(values = c('63H4' = '#CC6677',
                                #'hoku231019' = '#CC6699',
                                '6JK5' = '#DDCC77',
                                #'kolohe231016' = '#DDCC22',
                                '01L5' = '#117733',
                                #'liho231018' = '#116655',
                                #'lono231009' = '#6699CC',
                                '9FL3' = '#6688CC',
                                '9ON6' = '#EE8866',
                                #'nainoa231016' = '#EE8888',
                                'SPQC' = 'black'), name = "Animal") +  
  scale_shape_manual(values = c(15, 16, 17, 0, 1, 2, 3), name = "Trial Type") +  # Manually assign the shapes
  theme_minimal() +
  geom_hline(yintercept = 0, linetype = "dotted", color = "black", size = 1) +  # Dotted horizontal line at y = 0
  geom_vline(xintercept = 0, linetype = "dotted", color = "black", size = 1) + 
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1), # border around the plot area
    legend.spacing = unit(0.05, "lines"),  # Adjust spacing between legend entries
    legend.key.size = unit(0.4, "cm"),   # Adjust the size of the legend key
    legend.title = element_text(size = 10, face = "bold"), # Adjust the legend title size
    legend.text = element_text(size = 8),   # Adjust the legend text size
     plot.title = element_text(hjust = 0.5)  )

ggsave(paste0(fig_path, "/", "pca_phospho", ".png"), units = "in", width = 5, height = 3.5, dpi = 600, bg = "white")

```


# Heatmap for phospho
```{r}

# Step 1: Filter to Condition == "CA", ensure numeric Value
df_ca <- df_phospho_long %>%
  filter(Condition == "CA", !is.na(Value)) %>%
  mutate(Value = as.numeric(Value)) %>%
  filter(MeasurementType == "NormalizedPhosphorylation")

# Step 2: Identify top quartile proteins by total abundance
top_proteins <- df_ca %>%
  group_by(ProteinName) %>%
  summarize(total_phospho = sum(Value), .groups = "drop") %>%
  arrange(desc(total_phospho)) %>%
  mutate(rank = row_number()) %>%
  filter(rank <= ceiling(n() * 0.05)) %>%
  pull(ProteinName)

# Step 3: Log2-transform and pivot with aggregation
df_matrix <- df_ca %>%
  filter(ProteinName %in% top_proteins) %>%
  mutate(logValue = log2(Value + 1)) %>%
  select(ProteinName, AnimalName, logValue) %>%
  pivot_wider(
    names_from = AnimalName,
    values_from = logValue,
    values_fill = 0,
    values_fn = mean  # ✅ This resolves duplicates
  )

# Step 4: Convert to matrix
df_matrix <- as.data.frame(df_matrix)

# Now it's safe to set rownames
rownames(df_matrix) <- df_matrix$ProteinName

phospho_mat <- df_matrix %>%
  select(-ProteinName) %>%
  mutate(across(everything(), as.numeric)) %>%
  as.matrix()

# Create short protein names (first 20 chars)
full_names <- rownames(phospho_mat)
short_names <- substr(full_names, 1, 20)


# Now plot heatmap with these shortened names
phospho_hm <- Heatmap(
  t(phospho_mat),
  name = "log2(NormPhospho + 1)",
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  col = colorRamp2(
    breaks = seq(0, 8, length.out = 5),
    colors = brewer.pal(5, "Reds")
  ),
  column_names_gp = grid::gpar(fontsize = 7),
  column_names_rot = 45,  
  heatmap_legend_param = list(title = "log2(NormPhospho+1)"),
  column_labels = short_names
)


png(paste0(fig_path, "/heatmap_phospho_top5.png"), width = 1200, res = 150)
draw(phospho_hm, 
  heatmap_height = unit(6, "cm"))
dev.off()


```
# Heatmap for abundance - 1st 5th and 25th percentiles togehter
```{r}

# Step 1: Filter to Condition == "CA", ensure numeric Abundance
df_ca <- df_abundance_long %>%
  filter((Condition == "CA" | Condition == "PC6A"), !is.na(Abundance)) %>%
  mutate(Abundance = as.numeric(Abundance))


# Function to get top percentile proteins by total abundance
get_top_proteins <- function(df, percentile) {
  df %>%
    group_by(ProteinName) %>%
    summarize(total_abundance = sum(Abundance, na.rm = TRUE), .groups = "drop") %>%
    arrange(desc(total_abundance)) %>%
    mutate(rank = row_number()) %>%
    filter(rank <= ceiling(n() * percentile)) %>%
    pull(ProteinName)
}

# Prepare matrices for each percentile
prepare_matrix <- function(df, top_proteins) {
  df %>%
    filter(ProteinName %in% top_proteins) %>%
    mutate(logAbundance = log2(Abundance + 1)) %>%
    select(ProteinName, AnimalName, logAbundance) %>%
    pivot_wider(
      names_from = AnimalName,
      values_from = logAbundance,
      values_fill = 0,
      values_fn = mean
    ) -> wide_df

  mat <- as.matrix(wide_df %>% select(-ProteinName))
  rownames(mat) <- wide_df$ProteinName
  mat
}

# Load your abundance data here: df_abundance_long
# Assuming df_abundance_long has columns: ProteinName, AnimalName, Abundance, Condition

# Filter to relevant Condition if needed, e.g.:
df_abundance_long <- df_abundance_long %>%
  filter(!is.na(Abundance), Condition %in% c("CA", "PC6A")) %>%
  mutate(Abundance = as.numeric(Abundance))

# Get protein sets
top_1 <- get_top_proteins(df_abundance_long, 0.01)
top_5 <- get_top_proteins(df_abundance_long, 0.05)
top_25 <- get_top_proteins(df_abundance_long, 0.25)

# Prepare matrices
mat_25 <- prepare_matrix(df_abundance_long, top_25)
mat_5 <- prepare_matrix(df_abundance_long, top_5)
mat_1 <- prepare_matrix(df_abundance_long, top_1)

# Truncate rownames for plotting (max 30 chars)
truncate_names <- function(x) substr(x, 1, 30)
rownames(mat_1) <- truncate_names(rownames(mat_1))
rownames(mat_5) <- truncate_names(rownames(mat_5))
# For 25 and 5 percentile, we will NOT show row names

# Color function for heatmaps
col_fun <- colorRamp2(
  breaks = seq(0, max(c(mat_25, mat_5, mat_1), na.rm = TRUE), length.out = 5),
  colors = brewer.pal(5, "Reds")
)

# Set a consistent heatmap width
heatmap_width <- unit(3, "cm")

# Replacing Nas with zero
mat_25[is.na(mat_25)] <- 0
mat_5[is.na(mat_5)] <- 0
mat_1[is.na(mat_1)] <- 0

# Create heatmaps:

hm_25 <- Heatmap(
  mat_25,
  name = "Top 25%",
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  show_row_names = FALSE,
  show_column_names = TRUE,
  column_names_gp = gpar(fontsize = 7),
  column_names_rot = 45,
  col = col_fun,
  column_title = "Top 25% Proteins by Abundance",
  heatmap_legend_param = list(title = "log2(Abundance+1)"),
  width = heatmap_width
)

hm_5 <- Heatmap(
  mat_5,
  name = "Top 5%",
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  show_row_names = FALSE,
  show_column_names = TRUE,
  column_names_gp = gpar(fontsize = 7),
  column_names_rot = 45,
  col = col_fun,
  column_title = "Top 5% Proteins by Abundance",
  heatmap_legend_param = list(title = "log2(Abundance+1)"),
  width = heatmap_width
)

hm_1 <- Heatmap(
  mat_1,
  name = "Top 1%",
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  row_names_gp = gpar(fontsize = 8),
  column_names_gp = gpar(fontsize = 7),
  column_names_rot = 45,
  col = col_fun,
  column_title = "Top 1% Proteins by Abundance",
  heatmap_legend_param = list(title = "log2(Abundance+1)",
    legend_width = unit(6, "cm"),
    title_gp = gpar(fontsize = 8),
    labels_gp = gpar(fontsize = 8)),
  width = heatmap_width
)

# Capture grobs, and draw blue box around first 1% heatmap rows
# hm_25 with box around top 76 rows
grob_25 <- grid.grabExpr({
  draw(hm_25, show_heatmap_legend = FALSE)
  decorate_heatmap_body("Top 25%", {
    n <- 76
    total_rows <- nrow(hm_25@matrix)
    row_height <- 1 / total_rows
    bottom_y <- 1 - n * row_height
    
    grid.rect(
      x = unit(0, "npc"),
      y = unit(bottom_y + (1 - bottom_y) / 2, "npc"),
      width = unit(1, "npc"),
      height = unit(n * row_height, "npc"),
      just = c("left", "center"),
      gp = gpar(col = "blue", fill = NA, lwd = 2)
    )
  })
})

# hm_5 with box around top 16 rows
grob_5 <- grid.grabExpr({
  draw(hm_5, show_heatmap_legend = FALSE)
  decorate_heatmap_body("Top 5%", {
    n <- 16
    total_rows <- nrow(hm_5@matrix)
    row_height <- 1 / total_rows
    bottom_y <- 1 - n * row_height
    
    grid.rect(
      x = unit(0, "npc"),
      y = unit(bottom_y + (1 - bottom_y) / 2, "npc"),
      width = unit(1, "npc"),
      height = unit(n * row_height, "npc"),
      just = c("left", "center"),
      gp = gpar(col = "blue", fill = NA, lwd = 2)
    )
  })
})

# hm_1 with NO box
grob_1 <- grid.grabExpr({
  draw(hm_1, show_heatmap_legend = TRUE)
  # no decorate_heatmap_body call here
})

# Combine the grobs side-by-side
final_plot <- cowplot::plot_grid(
  grob_25, grob_5, grob_1,
  nrow = 1,
  align = "v",
  rel_widths = c(0.8, 1, 1.6)
)

final_plot

# Save plot if you want
# Save
ggsave(paste0(fig_path, "/adbundance_heatmaps.png"), final_plot, bg = "white", width = 9.5, height = 6, dpi = 300)


```


# Heatmap for phospho - 5th and 25th percentiles together
```{r}
library(tidyverse)
library(ComplexHeatmap)
library(circlize)
library(RColorBrewer)
library(grid)

# --- Step 1: Filter CA condition and select numeric values ---
df_ca <- df_phospho_long %>%
  filter((Condition == "CA" | Condition == "PC6A"), !is.na(Value)) %>%
  mutate(Value = as.numeric(Value)) %>%
  filter(MeasurementType == "NormalizedPhosphorylation")

# --- Step 2: Get 25th and 5th percentile protein lists ---
protein_ranks <- df_ca %>%
  group_by(ProteinName) %>%
  summarize(total_phospho = sum(Value), .groups = "drop") %>%
  arrange(desc(total_phospho)) %>%
  mutate(rank = row_number())

top_25_percent <- protein_ranks %>%
  filter(rank <= ceiling(n() * 0.25)) %>%
  pull(ProteinName)

top_5_percent <- protein_ranks %>%
  filter(rank <= ceiling(n() * 0.05)) %>%
  pull(ProteinName)

# --- Step 3: Pivot data and log-transform ---
make_matrix <- function(protein_list) {
  df_ca %>%
    filter(ProteinName %in% protein_list) %>%
    mutate(logValue = log2(Value + 1)) %>%
    select(ProteinName, AnimalName, logValue) %>%
    pivot_wider(
      names_from = AnimalName,
      values_from = logValue,
      values_fill = 0,
      values_fn = mean
    ) %>%
    column_to_rownames("ProteinName") %>%
    as.matrix()
}

mat_25 <- make_matrix(top_25_percent)
mat_5 <- mat_25[rownames(mat_25) %in% top_5_percent, , drop = FALSE]

# Create color palette
col_fun <- colorRamp2(
  breaks = seq(0, max(c(mat_25, mat_5), na.rm = TRUE), length.out = 5),
  colors = brewer.pal(5, "Reds")
)

# Short row names for display
short_names_5 <- substr(rownames(mat_5), 1, 20)

# --- Step 4: Create the 25% heatmap (no row labels) with box around 5% ---
# First get indices for top 5% in 25% heatmap
row_order_25 <- match(rownames(mat_25), rownames(mat_25))
rows_5_in_25 <- which(rownames(mat_25) %in% rownames(mat_5))

# Set equal widths for both heatmaps' core area
heatmap_width <- unit(6, "cm")

# Original rownames
original_names <- rownames(mat_5)

# Truncate to first 30 characters
truncated_names <- substr(original_names, 1, 30)

# Assign truncated names back as rownames for display only
rownames(mat_5) <- truncated_names

# Heatmap for top 25% (no protein names)
hm_25 <- Heatmap(
  mat_25,
  name = "Top 25%",
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  show_row_names = FALSE,     # Don't show protein labels
  show_column_names = TRUE,
  column_names_gp = gpar(fontsize = 8),
  row_names_gp = gpar(fontsize = 6),
  column_title = "Top 25% of Proteins by Phosphorylation",
  column_title_gp = gpar(fontsize = 10),
  col = col_fun,
  width = heatmap_width  # ✅ enforce equal width
)

# Heatmap for top 5% (with protein labels)
hm_5 <- Heatmap(
  mat_5,
  name = "Top 5%",
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  show_row_names = TRUE,      # Show protein labels here
  show_column_names = TRUE,
  column_names_gp = gpar(fontsize = 8),
  row_names_gp = gpar(fontsize = 8),
  column_title = "Top 5% of Proteins by Phosphorylation",
  column_title_gp = gpar(fontsize = 10),
  col = col_fun,
  heatmap_legend_param = list(
    title = "log2(NormPhospho+1)",
    legend_width = unit(6, "cm"),
    title_gp = gpar(fontsize = 8),
    labels_gp = gpar(fontsize = 8)),
  width = heatmap_width  # ✅ enforce equal width
)

# Combine
# Render each heatmap individually and capture as grobs
grob_25 <- grid.grabExpr({
  draw(hm_25, show_heatmap_legend = FALSE)
  decorate_heatmap_body("Top 25%", {
    n <- 23
    total_rows <- nrow(hm_25@matrix)  # number of rows in heatmap matrix
    
    row_height <- 1 / total_rows
    
    bottom_y <- 1 - n * row_height
    
    grid.rect(
      x = unit(0, "npc"),
      y = unit(bottom_y + (1 - bottom_y) / 2, "npc"),
      width = unit(1, "npc"),
      height = unit(n * row_height, "npc"),
      just = c("left", "center"),
      gp = gpar(col = "blue", fill = NA, lwd = 2)
    )
  })
})

grob_5 <- grid.grabExpr(draw(hm_5, heatmap_legend_side = "right", show_heatmap_legend = TRUE))

final_plot <- cowplot::plot_grid(grob_25, grob_5, nrow = 1, rel_widths = c(1, 1.6), align = "v", axis = "tb", greedy = TRUE)


# Save
ggsave(paste0(fig_path, "/phospho_heatmaps.png"), final_plot, bg = "white", width = 9.5, height = 6, dpi = 300)


```

# Dot plot for abundnace
```{r}

# Step 2: Compute mean per Protein per Type
mean_abundance_df <- df_abundance_long %>%
  filter(Condition == "CA") %>%
  group_by(ProteinName) %>%
  summarise(mean_abun = mean(as.numeric(Abundance), na.rm = TRUE), sd_abun= sd(as.numeric(Abundance), na.rm = TRUE), .groups = "drop") 
  

# Step 3: Order by mean value (descending)
mean_abundance_df <- mean_abundance_df %>%
  arrange(desc(mean_abun)) %>%
  mutate(log2_abun = log2(mean_abun)) %>%
  mutate(log2_sd= log2(sd_abun)) %>%
  mutate(ProteinName = factor(ProteinName, levels = unique(ProteinName)))

# Step 4: Plot
dot_abundance <- ggplot(mean_abundance_df, aes(x = ProteinName, y = log2_abun)) +
  geom_point(size = 0.5, alpha = 0.5) +
  theme_minimal() +
  labs(
    title = "Mean Abundance per Protein ",
    x = "Protein (ordered)",
    y = "Mean Value"
  ) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank()
  )

# # Add in animal codes
# df_abundance_long$AnimalName <- gsub("Hoku", "63H4", df_abundance_long$AnimalName)
# df_abundance_long$AnimalName <- gsub("Kolohe", "6JK5", df_abundance_long$AnimalName)
# df_abundance_long$AnimalName <- gsub("Liho", "01L5", df_abundance_long$AnimalName)
# df_abundance_long$AnimalName <- gsub("Nainoa", "9ON6", df_abundance_long$AnimalName)
# df_abundance_long$AnimalName<- gsub("Lono", "9FL3", df_abundance_long$AnimalName)
# 
# 
# df_abundance_long %>%
#   filter(Condition == "CA") %>%
#   mutate(logAbundance = log2(as.numeric(Abundance))) %>%
# ggplot(aes(x = ProteinName, y = logAbundance, color = AnimalName)) +
#   geom_jitter(width = 0.2, alpha = 0.7, size = 2) +
#   #facet_wrap(~ ProteinName, scales = "free_y") +
#   theme_minimal() +
#   scale_color_manual(values = c('63H4' = '#CC6677',
#                                 #'hoku231019' = '#CC6699',
#                                 '6JK5' = '#DDCC77',
#                                 #'kolohe231016' = '#DDCC22',
#                                 '01L5' = '#117733',
#                                 #'liho231018' = '#116655',
#                                 #'lono231009' = '#6699CC',
#                                 '9FL3' = '#6688CC',
#                                 '9ON6' = '#EE8866',
#                                 #'nainoa231016' = '#EE8888',
#                                 'SPQC' = 'black'), name = "Animal")+
#   labs(
#     title = "Abundance per Protein",
#     x = "Measurement Type",
#     y = "Abundance"
#   ) +
#   theme(
#     axis.text.x = element_blank(),
#     axis.ticks.x = element_blank(),
#     panel.grid = element_blank()
#   )
```

# Dot plot for phospho
```{r}

# Add in animal codes
df_phospho_long$AnimalName <- gsub("Hoku", "63H4", df_phospho_long$AnimalName)
df_phospho_long$AnimalName <- gsub("Kolohe", "6JK5", df_phospho_long$AnimalName)
df_phospho_long$AnimalName <- gsub("Liho", "01L5", df_phospho_long$AnimalName)
df_phospho_long$AnimalName <- gsub("Nainoa", "9ON6", df_phospho_long$AnimalName)
df_phospho_long$AnimalName<- gsub("Lono", "9FL3", df_phospho_long$AnimalName)


df_phospho_long %>%
  filter(Condition == "CA") %>%
  filter(MeasurementType == "NormalizedPhosphorylation") %>%
ggplot(aes(x = ProteinName, y = as.numeric(Value), color = AnimalName)) +
  geom_jitter(width = 0.2, alpha = 0.7, size = 2) +
  #facet_wrap(~ ProteinName, scales = "free_y") +
  theme_minimal() +
  scale_color_manual(values = c('63H4' = '#CC6677',
                                #'hoku231019' = '#CC6699',
                                '6JK5' = '#DDCC77',
                                #'kolohe231016' = '#DDCC22',
                                '01L5' = '#117733',
                                #'liho231018' = '#116655',
                                #'lono231009' = '#6699CC',
                                '9FL3' = '#6688CC',
                                '9ON6' = '#EE8866',
                                #'nainoa231016' = '#EE8888',
                                'SPQC' = 'black'), name = "Animal")+
  labs(
    title = "Normalized Phosphorylation per Protein",
    x = "Measurement Type",
    y = "Value"
  ) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank()
  )
```

# DOt plots togehters
```{r}
# Step 1: Filter both long datasets
abund_ca <- df_abundance_long %>%
  filter(Condition == "CA") %>%
  mutate(Type = "Abundance") %>%
  mutate(Value = Abundance)

phos_ca <- df_phospho_long %>%
  filter(Condition == "CA" & MeasurementType == "Phosphorylation") %>%
  mutate(Type = "Phospho") 

# Step 2: Combine both
combined_long <- bind_rows(abund_ca, phos_ca) %>%
  mutate(Log2Value = log2(as.numeric(Value)+1))

# Step 3: Summarize mean and SD per Protein and Type
summary_df <- combined_long %>%
  group_by(ProteinName, Type) %>%
  summarise(
    mean_value = mean(as.numeric(Log2Value), na.rm = TRUE),
    sd_value = sd(as.numeric(Log2Value), na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_value)) %>%
  mutate(ProteinName = factor(ProteinName, levels = unique(ProteinName)))

# Step 4: Plot
ggplot(summary_df, aes(x = ProteinName, y = mean_value, color = Type)) +
  geom_point(size = 0.25, alpha = 0.5) +
  geom_errorbar(aes(ymin = mean_value - sd_value,
                    ymax = mean_value + sd_value),
                width = 0.2, alpha = 0.5, size = 0.2) +
  scale_y_continuous(expand = expansion(mult = c(1, 1))) +
  scale_color_manual(
    name = "",
    values = c("Abundance" = "darkred", "Phospho" = "tomato"),
    labels = c("Abundance" = "Abundance", "Phospho" = "Phosphorylation")
  )+
  ylim(0, 30) +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Protein (ordered by mean Abundance)",
    y = "log2(Mean Value+1)"
  ) +
  theme(
    panel.grid = element_blank(),
    axis.text.y = element_blank(),  # You can tweak size for better readability
    legend.position = "top", legend.direction = "horizontal", legend.justification = "center"
  )

ggsave(paste0(fig_path, "/", "dot_abun_phospho", ".png"), units = "in", width = 3, height = 6, dpi = 600, bg = "white")

```


# Compare post breath-hold control vs PC (upreg)
```{r}

# Step 2: Compute mean abundance per protein per timepoint
abund_summary <- df_abundance_long %>%
  group_by(Protein = ProteinName, Condition) %>%
  summarise(mean_abund = mean(as.numeric(Abundance), na.rm = TRUE), .groups = "drop")

# Step 3: Pivot wider to align timepoints
fold_abundance_df <- abund_summary %>%
  pivot_wider(names_from = Condition, values_from = mean_abund) %>%
  # Step 4: Compute log2 fold changes
  mutate(
    log2_CC_vs_CA   = log2(CC / CA),
    log2_PC6C_vs_PC6A = log2(PC6C / PC6A),
    log2_diff = log2_PC6C_vs_PC6A - log2_CC_vs_CA 
  ) %>%
  # Step 5: Filter top 100 upregulated (most positive log2_diff)
  filter(!is.na(log2_diff)) %>%
  arrange(desc(log2_diff)) %>%
  slice_head(n = 100) %>%
  #mutate(Protein = factor(Protein, levels = Protein[order(log2_diff)])) %>% # preserve order for plot 
  # Step 6 Truncate protein labels and order for plotting
 mutate(
  Protein_label = str_trunc(Protein, width = 20),
  Protein_label = make.unique(Protein_label),
  Protein_label = factor(Protein_label, levels = rev(Protein_label[order(log2_diff)]))
)

# Step 6: Plot
ggplot(fold_abundance_df, aes(x = Protein_label, y = log2_diff)) +
  geom_point(size = 2, color= "steelblue") +
  theme_minimal() +
  labs(
    title = "Top 100 Upregulated Proteins (PC6 relative to Control)",
    x = "Protein",
    y = "Relative log2 Fold Change"
  ) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 4)
  )

ggsave(paste0(fig_path, "/", "upreg_PC6", ".png"), units="in", width = 8, height = 3, dpi = 600, bg = "white")


```

# Compare post breath-hold control vs PC (downreg)
```{r}

# Step 2: Compute mean abundance per protein per timepoint
abund_summary <- df_abundance_long %>%
  group_by(Protein = ProteinName, Condition) %>%
  summarise(mean_abund = mean(as.numeric(Abundance), na.rm = TRUE), .groups = "drop")

# Step 3: Pivot wider to align timepoints
fold_abundance_df <- abund_summary %>%
  pivot_wider(names_from = Condition, values_from = mean_abund) %>%
  # Step 4: Compute log2 fold changes
  mutate(
    log2_CC_vs_CA   = log2(CC / CA),
    log2_PC6C_vs_PC6A = log2(PC6C / PC6A),
    log2_diff = log2_PC6C_vs_PC6A - log2_CC_vs_CA 
  ) %>%
  # Step 5: Filter top 100 upregulated (most positive log2_diff)
  filter(!is.na(log2_diff)) %>%
  arrange(log2_diff) %>%
  slice_head(n = 100) %>%
  #mutate(Protein = factor(Protein, levels = Protein[order(log2_diff)])) %>% # preserve order for plot 
  # Step 6 Truncate protein labels and order for plotting
 mutate(
  Protein_label = str_trunc(Protein, width = 20),
  Protein_label = make.unique(Protein_label),
  Protein_label = factor(Protein_label, levels = Protein_label[order(log2_diff)])
)

# Step 6: Plot
ggplot(fold_abundance_df, aes(x = Protein_label, y = log2_diff)) +
  geom_point(size = 2, color= "steelblue") +
  theme_minimal() +
  labs(
    title = "Top 100 Downregulated Proteins (PC6 relative to Control)",
    x = "Protein",
    y = "Relative log2 Fold Change"
  ) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 4)
  )

ggsave(paste0(fig_path, "/", "downreg_PC6", ".png"), units="in", width = 8, height = 3, dpi = 600, bg = "white")


```
