---
title: "main"
author: "Ashley Blawas"
date: '2024-06-13'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries include=FALSE}
library(tidyverse)
library(readxl)
library(stringr)
library(dplyr)
```

```{r}
#options(repos = "http://cran.rstudio.com/")

#Install tinytex package
#install.packages("tinytex")

#Install tinytex
#tinytex::install_tinytex(force = TRUE)
```

```{r  include=FALSE}
# Where does data live?
#data_path = "C:/Users/ashle/Dropbox/Ashley/Post-Doc/Projects/DQO preconditioning/data"

# What is the file name?
#file_name = "10364_Supplemental_Data_061224.xlsx"

# What are the sheet names
sheet1 = "S1. Unenriched Metadata"
sheet2 = "S2. Uneriched Abundances" # NP = # of precursors, PGQ = protein group quantity
sheet3 = "S3. Phos Metadata"
sheet4 = "S4. Phos localized" # Place where it was on the protein
sheet5 = "S5. Phos not localized" # Not able to place where it was on the protein
sheet6 = "S6. Phos filtered" # 
```

## Load in the metadata
```{r}
my_excel <- read_excel("10364_Supplemental_Data_061224.xlsx")

# Load in the Unenriched Metadata
#unenriched_metadata <- read_excel(paste0(data_path, "/", file_name), sheet = sheet1)
unenriched_metadata <- read_excel(paste0("10364_Supplemental_Data_061224.xlsx"), sheet = sheet1)

# Make the Animal IDs for the controls unique
n_na = sum(is.na(unenriched_metadata$Condition))
unenriched_metadata$Condition[is.na(unenriched_metadata$Condition)] <- paste0("NA_VALUE_", seq_len(n_na))

# Load in the Phosphorylation Metadata
#phos_metadata <- read_excel(paste0(data_path, "/", file_name), sheet = sheet3)
phos_metadata <- read_excel(paste0("10364_Supplemental_Data_061224.xlsx"), sheet = sheet3)

# Make the Animal IDs for the controls unique
n_na = sum(phos_metadata$Replicate == "SPQC_prep")
phos_metadata$Replicate[phos_metadata$Replicate == "SPQC_prep"] <- paste0("SPQC_prep", seq_len(n_na))

n_na = sum(phos_metadata$Replicate == "SPQC_tech")
phos_metadata$Replicate[phos_metadata$Replicate == "SPQC_tech"] <- paste0("SPQC_tech", seq_len(n_na))
```

## Load in the unenriched dataset
```{r}
# Load in the Unenriched Metadata
#unenriched_abundances <- read_excel(paste0(data_path, "/", file_name), sheet = sheet2)
unenriched_abundances <- read_excel(paste0("10364_Supplemental_Data_061224.xlsx"), sheet = sheet2)

for (i in 1:nrow(unenriched_metadata)){
  curr_name = names(unenriched_abundances)[i+3] 
  curr_id = unlist(str_extract_all(curr_name, "ID.+htrms"))
  
  temp_metadata = unenriched_metadata %>%
  filter(str_detect(`File Name`, curr_id))
  
  curr_animal = temp_metadata$Animal
  curr_condition = temp_metadata$Condition
  
  unenriched_abundances <- unenriched_abundances %>% rename(!! paste0(curr_animal, "_", curr_condition, "_", "NP") := curr_name)
  
}


for (i in 1:nrow(unenriched_metadata)){
  curr_name = names(unenriched_abundances)[i+36] 
  curr_id = unlist(str_extract_all(curr_name, "ID.+htrms"))
  
  temp_metadata = unenriched_metadata %>%
  filter(str_detect(`File Name`, curr_id))
  
  curr_animal = temp_metadata$Animal
  curr_condition = temp_metadata$Condition
  
  unenriched_abundances <- unenriched_abundances %>% rename(!! paste0(curr_animal, "_", curr_condition, "_", "PGQ") := curr_name)
  
}
```

## Load in the filtered phosphoprotein dataset
```{r}
# Load in the Unenriched Metadata
#phos_filtered <- read_excel(paste0(data_path, "/", file_name), sheet = sheet6)
phos_filtered <- read_excel(paste0("10364_Supplemental_Data_061224.xlsx"), sheet = sheet6)

# Drop CV columns
phos_filtered <- phos_filtered %>% dplyr::select(!starts_with("CV"))

# Change data frame organization to have names
for (i in 1:nrow(phos_metadata)){
  curr_name = names(phos_filtered)[i+5] # Add 5 because there are 5 columns that are not IDs
  curr_id = unlist(str_extract_all(curr_name, "ID.+htrms"))
  
  temp_metadata = phos_metadata %>%
  filter(str_detect(`File Name`, curr_id))
  
  curr_animal = temp_metadata$Condition
  curr_condition = temp_metadata$Replicate
  
  phos_filtered <- phos_filtered %>% rename(!! paste0(curr_animal, "_", curr_condition, "_", "EG") := curr_name)
  
}
```


# Make data frame for plotting

```{r}
phos_filtered <- phos_filtered %>% select(-PG.ProteinAccessions, -PG.ProteinDescriptions, -PG.ProteinNames, -EG.PrecursorId)

phos_filtered <- phos_filtered %>% pivot_longer(!PG.Genes, names_to = "Sample") %>%
  rename(Value = value) %>% 
  rename(Gene = PG.Genes)

# Mutate new columns
phos_filtered <- phos_filtered %>% mutate(Animal = str_extract(Sample, "^\\D+")) %>%
  mutate(Date = str_extract(Sample, "\\d+")) %>%
  mutate(Condition = sub("^[^_]*_([^_]*).*", "\\1", Sample)) %>%
  dplyr::select(-Sample) %>%
  dplyr::select(Gene, Animal, Condition, Date, Value)
```

# Choose a gene to test
```{r}
#Testing HBA
phos_filtered %>% filter(Gene == "HBA") %>%
  group_by(Animal, Condition) %>%
  summarize(MaxValue = max(Value)) %>%
  ungroup() %>%
  ggplot(aes(x=Condition, y = MaxValue)) +
  facet_wrap(vars(Animal), scales = "free") +
  geom_point()
```

```{r}
#NEW CECE
#Testing FGA
phos_filtered %>% filter(Gene == "FGA") %>%
  group_by(Animal, Condition) %>%
  summarize(MaxValue = max(Value)) %>%
  ungroup() %>%
  ggplot(aes(x=Condition, y = MaxValue)) +
  facet_wrap(vars(Animal), scales = "free") +
  geom_point()
```

```{r}
#NEW CECE
#Testing KNG1
phos_filtered %>% filter(Gene == "KNG1") %>%
  group_by(Animal, Condition) %>%
  summarize(MaxValue = max(Value)) %>%
  ungroup() %>%
  ggplot(aes(x=Condition, y = MaxValue)) +
  facet_wrap(vars(Animal), scales = "free") +
  geom_point()
```

```{r}
#NEW CECE
#Testing RHAG
phos_filtered %>% filter(Gene == "RHAG") %>%
  group_by(Animal, Condition) %>%
  summarize(MaxValue = max(Value)) %>%
  ungroup() %>%
  ggplot(aes(x=Condition, y = MaxValue)) +
  facet_wrap(vars(Animal), scales = "free") +
  geom_point()
```

```{r}
#NEW CECE
#Testing FN1
phos_filtered %>% filter(Gene == "FN1") %>%
  group_by(Animal, Condition) %>%
  summarize(MaxValue = max(Value)) %>%
  ungroup() %>%
  ggplot(aes(x=Condition, y = MaxValue)) +
  facet_wrap(vars(Animal), scales = "free") +
  geom_point()
```

```{r}
#NEW CECE
#Testing CFL1
phos_filtered %>% filter(Gene == "CFL1") %>%
  group_by(Animal, Condition) %>%
  summarize(MaxValue = max(Value)) %>%
  ungroup() %>%
  ggplot(aes(x=Condition, y = MaxValue)) +
  facet_wrap(vars(Animal), scales = "free") +
  geom_point()
```

```{r}
#NEW CECE
#Testing SERPINF2
phos_filtered %>% filter(Gene == "SERPINF2") %>%
  group_by(Animal, Condition) %>%
  summarize(MaxValue = max(Value)) %>%
  ungroup() %>%
  ggplot(aes(x=Condition, y = MaxValue)) +
  facet_wrap(vars(Animal), scales = "free") +
  geom_point()
```

```{r}
#NEW CECE
#Testing SPTB
phos_filtered %>% filter(Gene == "SPTB") %>%
  group_by(Animal, Condition) %>%
  summarize(MaxValue = max(Value)) %>%
  ungroup() %>%
  ggplot(aes(x=Condition, y = MaxValue)) +
  facet_wrap(vars(Animal), scales = "free") +
  geom_point()
```

```{r}
#NEW CECE
#Testing SLC43A1
phos_filtered %>% filter(Gene == "SLC43A1") %>%
  group_by(Animal, Condition) %>%
  summarize(MaxValue = max(Value)) %>%
  ungroup() %>%
  ggplot(aes(x=Condition, y = MaxValue)) +
  facet_wrap(vars(Animal), scales = "free") +
  geom_point()
```

```{r}
#NEW CECE
#Testing SPTB
phos_filtered %>% filter(Gene == "SPTB") %>%
  group_by(Animal, Condition) %>%
  summarize(MaxValue = max(Value)) %>%
  ungroup() %>%
  ggplot(aes(x=Condition, y = MaxValue)) +
  facet_wrap(vars(Animal), scales = "free") +
  geom_point()
```

# Look at change between time point A & C and C & D
```{r}
#Ashley's initial work below
#upup <- phos_filtered %>% 
#  group_by(Gene, Animal) %>%
#  if

#NEW CECE
#Making a version of the phos_filtered dataset with only the max values for each condition
phos_filtered_max_values <- phos_filtered %>%
  group_by(Gene, Animal, Condition) %>%
  summarize(MaxValue = max(Value)) %>%
  filter(Condition %in% c("CA", "CC", "CD", "PC6A", "PC6C", "PC6D")) %>%
  ungroup()

#Making datasets that calculate and store the differences between CA and CC, CD and CC, PC6C and PC6A, and PC6D and PC6C
phos_filtered_change_CA_CC <- phos_filtered_max_values %>%
  group_by(Gene, Animal) %>%
  summarize(Difference = MaxValue[Condition == "CC"] - MaxValue[Condition == "CA"]) %>%
  mutate(Condition = "CA_CC") %>%
  bind_rows(phos_filtered_max_values)

phos_filtered_change_CC_CD <- phos_filtered_max_values %>%
  group_by(Gene, Animal) %>%
  summarize(Difference = MaxValue[Condition == "CD"] - MaxValue[Condition == "CC"]) %>% mutate(Condition = "CC_CD") %>%
  bind_rows(phos_filtered_max_values)

phos_filtered_change_PC6A_PC6C <- phos_filtered_max_values %>%
  group_by(Gene, Animal) %>%
  summarize(Difference = MaxValue[Condition == "PC6C"] - MaxValue[Condition == "PC6A"]) %>%
  mutate(Condition = "PC6A_PC6C") %>%
  bind_rows(phos_filtered_max_values)

phos_filtered_change_PC6C_PC6D <- phos_filtered_max_values %>%
  group_by(Gene, Animal) %>%
  summarize(Difference = MaxValue[Condition == "PC6D"] - MaxValue[Condition == "PC6C"]) %>%
  mutate(Condition = "PC6C_PC6D") %>%
  bind_rows(phos_filtered_max_values)

#Making an overall dataset combining all of these datasets
phos_filtered_change <- bind_rows(phos_filtered_change_CA_CC, phos_filtered_change_CC_CD, phos_filtered_change_PC6A_PC6C, phos_filtered_change_PC6C_PC6D) %>%
  filter(Condition %in% c("CA_CC", "CC_CD", "PC6A_PC6C", "PC6C_PC6D"))

print(phos_filtered_change)
```

```{r}
#Printing a sample geom_point to visualize these changes for the HBA gene
phos_filtered_change %>%
  filter(Gene == "HBA") %>%
  group_by(Animal, Condition) %>%
  summarize(MaxValue = max(Difference)) %>%
  ungroup() %>%
  ggplot(aes(x=Condition, y = MaxValue)) + facet_wrap(vars(Animal), scales = "free") + geom_point()
```

#Making new dataset to conduct paired t-test w/ protein abundance and phosphorylation
```{r}
# Load in sheet2
sheet2_metadata <- read_excel(paste0("10364_Supplemental_Data_061224.xlsx"), sheet = sheet2)

# Replace NaN values with 0 for all numeric columns
sheet2_metadata[] <- lapply(sheet2_metadata, function(x) {
  if (is.numeric(x)) {
    x[is.nan(x)] <- 0
  }
  return(x)
})

# Function to convert NaN to NA in character columns
sheet2_metadata[] <- lapply(sheet2_metadata, function(x) {
  if (is.character(x)) {
    x[x == "NaN"] <- NA  # Convert "NaN" string to NA
  }
  return(x)
})

# Replace NA values with "0" in the entire data frame
sheet2_metadata[is.na(sheet2_metadata)] <- "0"

# Convert the character columns to numeric columns
sheet2_metadata$"[1] ID112569_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[1] ID112569_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[2] ID112570_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[2] ID112570_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[3] ID112571_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[3] ID112571_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[4] ID112572_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[4] ID112572_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[5] ID112573_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[5] ID112573_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[6] ID112574_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[6] ID112574_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[7] ID112575_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[7] ID112575_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[8] ID112576_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[8] ID112576_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[9] ID112577_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[9] ID112577_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[10] ID112578_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[10] ID112578_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[11] ID112579_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[11] ID112579_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[12] ID112580_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[12] ID112580_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[13] ID112581_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[13] ID112581_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[14] ID112582_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[14] ID112582_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[15] ID112583_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[15] ID112583_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[16] ID112584_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[16] ID112584_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[17] ID112585_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[17] ID112585_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[18] ID112586_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[18] ID112586_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[19] ID112587_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[19] ID112587_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[20] ID112588_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[20] ID112588_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[21] ID112589_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[21] ID112589_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[22] ID112590_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[22] ID112590_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[23] ID112592_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[23] ID112592_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[24] ID112593_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[24] ID112593_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[25] ID112594_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[25] ID112594_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[26] ID112595_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[26] ID112595_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[27] ID112596_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[27] ID112596_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[28] ID112597_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[28] ID112597_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[29] ID112598_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[29] ID112598_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[30] ID112599_01_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[30] ID112599_01_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[31] IDSPQC_02_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[31] IDSPQC_02_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[32] IDSPQC_03_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[32] IDSPQC_03_MA10140C_10364_050724.htrms.PG.Quantity")

sheet2_metadata$"[33] IDSPQC_04_MA10140C_10364_050724.htrms.PG.Quantity" <- as.numeric(sheet2_metadata$"[33] IDSPQC_04_MA10140C_10364_050724.htrms.PG.Quantity")

# Log2 normalization applied to all numeric columns
# Added 1 before log transformation to avoid taking log of zero
sheet2_metadata_normalized <- as.data.frame(lapply(sheet2_metadata, function(x) if(is.numeric(x)) log2(x + 1) else x))
```

#Conducting paired T-tests to assess how an individual's phopho levels at a specific residue change between points A, B, C, and D
```{r}
# Generic paired t-test format
#t_test_result <- t.test(data_set$column_before, data_set$column_after, paired = TRUE)

#Kolohe CA to CC
t_test_kolohe_CA_CC <- t.test(sheet2_metadata_normalized$`X.1..ID112569_01_MA10140C_10364_050724.htrms.PG.Quantity`, sheet2_metadata_normalized$`X.2..ID112570_01_MA10140C_10364_050724.htrms.PG.Quantity`, paired = TRUE)
print(t_test_kolohe_CA_CC)
#This approach cannot be correct because it is bunching all of the residues together to compare between these two scenarios

#Instead, try rearranging the dataset to compare CA and CC phosphorylation for a specific residue among all individuals

#other issue is I can't even use Sheet 2 exclusively bc it is by protein, not residue, bc protein amount is done by protein

#therefore, sheet 2 is more protein-wide phosphorylation rather than residue phosphorylation
```
